### 2023-05-04 paper/model-documentation.qmd
updating model, model documentation

---
title: Modeling costs and benefits of genomic wastewater surveillance
subtitle: Model appendix draft
#author:
#  - name: Pedro Nascimento de Lima
#    email: plima@rand.org
#    affiliations: 
#        - id: rand
#          name: RAND Corporation
#          department: Engineering and Applied Sciences
#          address: 1776 Main Street
#          city: Santa Monica
#          state: CA
#          postal-code: 90401
#    attributes:
#        corresponding: true
#  - name: and colleagues
#    email: others@rand.org
#    affiliations:
#        - ref: rand
abstract: |
  Introduces a model to assess the value provided by wastewater genomic surveillance network. This document is meant to be used as a supplementary appendix of the paper.
keywords: 
  - Genomic Surveillance
  - Pandemic Response
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  elsevier-html:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  docx:
      cite-style: super
---

<!-- Template source: https://github.com/quarto-journals/elsevier download repository and place it within this folder.-->

# Model

We start with a standard disease transmission model describing person-to-person transmission of an infectious pathogen across $n$ jurisdictions. The state-space of the model is comprised of a set of health statuses $\mathbf{S}, \mathbf{E}, \mathbf{P}, \mathbf{I}, \mathbf{R}$, each a $1 \times n$ vector representing the population within each compartment, respectively representing the number of Susceptibles, Exposed, Pre-Symptomatic, Infected and Removed[^1]. @eq-ode1 describes a system of differential equations representing the model, and @tbl-parameters contains parameter definitions, values and their sources.

[^1]: We start with a simplified version including only a minimal set of disease states before expanding the state-space of the model to include asymptomatic, mildy symptomatic and severely symptomatic compartments

$$ 
\begin{align}
\dot{\mathbf{S}} &=\ - \boldsymbol{\lambda}_t \odot \mathbf{S} + p_S \mathbf{m}_S \\
\dot{\mathbf{E}} &=\ \boldsymbol{\lambda}_t \odot \mathbf{S} - \sigma \mathbf{E} + p_E \mathbf{m}_E \\
\dot{\mathbf{P}} &=\ \sigma \mathbf{E} - \delta \mathbf{P} + p_P \mathbf{m}_P \\
\dot{\mathbf{I}} &=\ \delta \mathbf{P} - \gamma \mathbf{I} + p_I \mathbf{m}_I \\
\dot{\mathbf{R}} &=\ \gamma \mathbf{I} + p_R \mathbf{m}_R \\
\end{align}
$$ {#eq-ode1}

where $\dot{\mathbf{S}}, \dot{\mathbf{E}}, \dot{\mathbf{P}}, \dot{\mathbf{I}}, \dot{\mathbf{R}}$ represent the time derivatives of each disease status, greek letters represent progression rates and $p_x\mathbf{m}_x$ terms represent migration flows $\mathbf{m}_x$ modulated by propensity to travel $p_x$ at disease state $x \in \{S, E, P, I, R\}$. Force of infection $\boldsymbol{\lambda}_t$ is computed using a time-varying effective contact-rate vector $\boldsymbol{\beta}_t$ which will depend on contact rates within each strata as well as nonpharmaceutical interventions (discussed later). Infectiousness of each disease status $\mathbf{P}$ and $\mathbf{I}$ are modulated by constants $c_p$ and $c_i$.

$$ 
\boldsymbol{\lambda}_t = \boldsymbol{\beta}_t (c_p\mathbf{P}_t + c_i\mathbf{I}_t)
$$ {#eq-force-inf}

## Nonpharmaceutical interventions {#npi}

Following mitigation plans adopted during the COVID-19 pandemic, nonpharmaceutical interventions (NPIs) are modeled as a set of intervention portfolios organized within intensity levels with a pre-specified marginal effectiveness $\tau$ at each intervention level $l \in L = \{1,2,3,...,l_{max}\}$. In this analysis, $\tau$ is specified as a percent reduction in infectious contacts induced by a set of interventions introduced by public health decision-makers. In principle, $\tau_l$ can be specified so that it is dependent on the intervention level as in [@nascimentodelima2021], but this analysis assumes a constant, additive marginal effectiveness of interventions. For instance, if $\tau = 0.15$, then introducing the first intervention level reduces potential infectious contacts by 15% relative to a no-intervention scenario and introducing the second intervention level reduces infectious contacts by 30%. Under this specification, $l_{max} \tau \in (0,1)$.

$$ 
\boldsymbol{\beta}_t = (1-\mathbf{l}_t\tau) \boldsymbol{\beta}_b
$$ {#eq-betat}

and the intervention level $\mathbf{l}_t$ is a jurisdiction-dependent nonpharmaceutical intervention level. $\mathbf{l}_t$ can be either specified as an exogenous variable or an endogenous variable controlled by local publich health departments.

## Introducing and removing NPIs

In this model, local public health decision-makers follow epidemiological outcomes that are proportional to $\mathbf{I_t}$ (such as case counts) to decide the appropriate level of intervention. Thus, we specify a target NPI level $\mathbf{l}_t^*$ as a function of the disease state.

$$ 
\mathbf{l}_t^* = cI
$$ {#eq-targetnpi}

This formulation approximates phased mitigation plans where policymakers introduce or remove interventions depending on disease prevalence. While public health decision-makers strive to use up-to-date information to introduce and remove NPIs, they do not have perfect and instantaneous information about the disease state. In continuous time, we represent this process with a stock vector $\mathbf{L}$ representing the NPI level in effect

$$ 
\dot{\mathbf{L}} = \frac{1}{a}(\mathbf{l}_t^*-\mathbf{L})
$$ {#eq-actualnpi}

where $a$ represents the adjustment time for NPIs (i.e., one or two weeks) using conventional measures.

## Genomic surveillance

Genomic surveillance might provide an early signal to decision makers by providing information about current disease prevalence, hence effectively reducing $a$. One might also represent genomic surveillance using a mechanistic formulation where

```{=html}
<!-- 
Sections to complete

## Travel

Individuals travel from jurisdiction $i$ to $j$ at a daily rate $r_{ij}$. This baseline rate is multiplied by a factor  by a factor  The daily travel flows $t_{ij} \sim Pois(\lambda_{ij})$. We used data from {blank} to estimate travel flows between

### Information sharing and coordination

Jurisdictions can share information and use each other's epidemiological indicators to decide when to introduce and remove Nonpharmaceutical interventions.

## Effect of genomic surveillance

Genomic surveillance is assumed to provide benefits to public health decision makers by providing an early signal of surges that would be otherwise detected later via testing of symptomatic individuals.

# Model parameters {#parameters}

@tbl-parameters presents the set of model parameters used in this analysis, their meaning, and their baseline values and sources used.

-->
```
# Tables

```{r}
#| label: tbl-parameters
#| tbl-cap: Parameters and values
#| echo: false

params <- data.frame(Parameter = "$\\lambda_{ij}$ Travel rate", `Value` = "Mobility Matrix A", Source = "[@Feynman1963118]")

knitr::kable(params)
```

# References {.unnumbered}

### 2023-05-04 R/scripts/00_metapopulation_no_deaths.R

@@ -12,7 +12,6 @@
 
 library(odin)
 
-
 # Helper functions:
 
 SEIR_cont <- odin::odin({
@@ -23,10 +22,19 @@ SEIR_cont <- odin::odin({
   # Assignments in arrays are translated by odin to for loops in C
   # See https://mrc-ide.github.io/odin/articles/functions.html
 
+
+  # Disease transmission equation
   lambda_prod[ , ] <- beta[i, j] * (I[j] + P[j])
   lambda[] <- sum(lambda_prod[i, ]) # rowSums
 
+
+  # Mobility equations
+  # mob_prod object is re-used for each disease status.
+
+  # Calculate the number of people leaving each state
   mob_prod[ , ] <- S[i] * C[i, j]
+
+  # and compute net flow of mobility to S
   mob_S[] <- sum(mob_prod[, i])     # colSums
 
   mob_prod[ , ] <- E[i] * C[i, j]
@@ -43,10 +51,12 @@ SEIR_cont <- odin::odin({
 
   N[] <- S[i] + E[i] + P[i] + I[i] + R[i]
 
-  # I think this outputs N
+  # output desired variables for debugging
   output(N[]) <- TRUE
   output(lambda_prod[]) <- TRUE
   output(lambda[]) <- TRUE
+  output(S_E[]) <- TRUE
+  output(mob_prod[]) <- TRUE
 
   ## Epidemiological Flows
   S_E[] <- S[i] * lambda[i]
@@ -55,11 +65,11 @@ SEIR_cont <- odin::odin({
   I_R[] <- gamma * I[i]
 
   ## Derivatives
-  deriv(S[]) <- - S_E[i]         + M[1] * mob_S[i]
-  deriv(E[]) <- S_E[i] - E_P[i]  + M[2] * mob_E[i]
-  deriv(P[]) <- E_P[i] - P_I[i]  + M[3] * mob_P[i]
-  deriv(I[]) <- P_I[i] - I_R[i]  + M[4] * mob_I[i]
-  deriv(R[]) <- I_R[i]           + M[5] * mob_R[i]
+  deriv(S[]) <- - S_E[i]         + mp[1] * mob_S[i]
+  deriv(E[]) <- S_E[i] - E_P[i]  + mp[2] * mob_E[i]
+  deriv(P[]) <- E_P[i] - P_I[i]  + mp[3] * mob_P[i]
[…]

### 2023-05-04 paper/model-documentation.qmd
pushing latest version of model documentation

@@
 title: Modeling costs and benefits of genomic wastewater surveillance
-subtitle: Model appendix draft
+subtitle: Supplementary appendix draft
 #author:
@@
 
-## Genomic surveillance
-
-Genomic surveillance might provide an early signal to decision makers by providing information about current disease prevalence, hence effectively reducing $a$.
-One might also represent genomic surveillance using a mechanistic formulation where
-
 ```{=html}
@@
 Sections to complete
+
+## Genomic surveillance
+
+Genomic surveillance might provide an early signal to decision makers by providing an early signal of prevalence, hence effectively reducing $a$.
 

### 2023-05-04 R/functions/model.R

@@ -1,2 +1,3 @@
 
 # Model should be defined here.
+

### 2023-05-04 R/odin_models/logitic_growth.R

@@ -1,0 +1,23 @@
+
+# -------------------------------------------------------------------------
+# This is an Odin script
+# Consult the odin documentation to understand the valid syntax.
+# https://mrc-ide.github.io/odin/
+# Do not assume this is a script
+# -------------------------------------------------------------------------
+
+# Parameters --------------------------------------------------------------
+N0 <- user(1)
+K <- user(100)
+r <- user()
+
+
+# Initial Conditions ------------------------------------------------------
+
+initial(N) <- N0
+
+# Flows and Derivatives ---------------------------------------------------
+
+deriv(N) <- r * N * (1 - N / K)
+
+

### 2023-05-05 paper/model-documentation.qmd
adding model diagram

@@
     keep-tex: true
+    mermaid:
+      theme: neutral
     journal:
@@
       cite-style: super
+  html:
+    keep-tex: true
+    mermaid:
+      theme: neutral
 ---
@@
 where $\dot{\mathbf{S}}, \dot{\mathbf{E}}, \dot{\mathbf{P}}, \dot{\mathbf{I}}, \dot{\mathbf{R}}$ represent the time derivatives of each disease status, greek letters represent progression rates and $p_x\mathbf{m}_x$ terms represent migration flows $\mathbf{m}_x$ modulated by propensity to travel $p_x$ at disease state $x \in \{S, E, P, I, R\}$.
+@fig-model-diagram illustrates the model.
+
+```{mermaid}
+%%| label: fig-model-diagram
+%%| fig-cap: Model diagram
+%%| fig-width: 6.5
+%%| fig-align: center
+flowchart LR
+  S --> E --> P --> I --> R
+  I --> l --> E
+```
+
 Force of infection $\boldsymbol{\lambda}_t$ is computed using a time-varying effective contact-rate vector $\boldsymbol{\beta}_t$ which will depend on contact rates within each strata as well as nonpharmaceutical interventions (discussed later).

### 2023-05-11 paper/model-documentation.qmd
updates model documentation

@@
 title: Modeling costs and benefits of genomic wastewater surveillance
-subtitle: Supplementary appendix draft
+# subtitle: Supplementary appendix draft
 #author:
@@
 <!-- Template source: https://github.com/quarto-journals/elsevier download repository and place it within this folder.-->
+
+# Background
+
+While researchers have demonstrated that wastewater-based surveillance (WBS) systems can provide an early warning system for pandemic response [@zhu2021] and be used to support estimation of the latent number of infections in COVID-19 transmission models [@soller2022], the value of information provided by WBS has yet to be demonstrated.
+
+This paper develops a stochastic metapopulation disease transmission model and uses it to assess the conditions under which WBS systems can provide value to local public health decision makers who respond to a pandemic using surveillance data.
+We evaluate a range of scenarios that represent different potential pathogens and highlight scenarios under which WBS systems may or may not be cost-effective.
 
@@
 
-## Introducing and removing NPIs
+## Controlling NPIs
 
@@
 where $a$ represents the adjustment time for NPIs (i.e., one or two weeks) using conventional measures.
+
+## Cost of interventions
+
+The cost of reducing infectious contacts over a time period $t_0$ to $t_f$ may be defined as a function of the effect of interventions
+
+$$
+C = \int_{t_0}^{t_f}{f(\mathbf{l}_t\tau)}
+$${#eq-cost}
+
+where $f(x)$ is a function that maps a percent reduction in potentially infectious contacts to a monetary value.
+At the time of the intervention, decision-makers do not have precise information about the costs of interventions $\tau$ or its costs represented by function $f(x)$.
 

### 2023-05-12 paper/model-documentation.qmd
updating model documentation

@@
 
-# Background
+# Introduction
 
@@
 We evaluate a range of scenarios that represent different potential pathogens and highlight scenarios under which WBS systems may or may not be cost-effective.
+<!--
+Our results highlight that the value of WBS systems is contingent on a set of biological and societal parameters, including pathogen characteristics (i.e., transmissibility and severity), as well as the geographic scale of
+-->
+# Methods
 
-# Model
+## Disease Transmission Model
 
@@
   S --> E --> P --> I --> R
-  I --> l --> E
+  I --> D
 ```
@@
 
-## Nonpharmaceutical interventions {#npi}
+### Nonpharmaceutical interventions {#npi}
 
@@
 and the intervention level $\mathbf{l}_t$ is a jurisdiction-dependent nonpharmaceutical intervention level.
-$\mathbf{l}_t$ can be either specified as an exogenous variable or an endogenous variable controlled by local publich health departments.
-
-## Controlling NPIs
-
-In this model, local public health decision-makers follow epidemiological outcomes that are proportional to $\mathbf{I_t}$ (such as case counts) to decide the appropriate level of intervention.
-Thus, we specify a target NPI level $\mathbf{l}_t^*$ as a function of the disease state.
+$\mathbf{l}_t$ can be either specified as an exogenous variable or an endogenous variable controlled by local public health departments.
+Public health decision-makers follow epidemiological outcomes that are proportional to $\mathbf{I_t}$ (such as case counts) to decide the appropriate level of intervention.
+Thus, we specify a target NPI level $\mathbf{l}_t^*$ as a linear function of disease prevalence.
 
 $$ 
-\mathbf{l}_t^* = cI
+\mathbf{l}_t^* = \min{(cI, l_{max})}
 $$ {#eq-targetnpi}
 
-This formulation approximates phased mitigation plans where policymakers introduce or remove interventions depending on disease prevalence.
-While public health decision-makers strive to use up-to-date information to introduce and remove NPIs, they do not have perfect and instantaneous information about the disease state.
+This formulation approximates phased mitigation plans adopted during the COVID-19 pandemic, where policymakers introduce or remove interventions depending on disease prevalence or other severity indicators.
+While public health decision-makers strive to use up-to-date information to introduce and remove NPIs, they do not have perfect and instantaneous information on disease prevalence; hence actual policy behaviors lag epidemiological indicators.
 In continuous time, we represent this process with a stock vector $\mathbf{L}$ representing the NPI level in effect
@@
 $$
-C = \int_{t_0}^{t_f}{f(\mathbf{l}_t\tau)}
+C_{npi} = \int_{t_0}^{t_f}{f(\mathbf{l}_t\tau)}
 $${#eq-cost}
 
-where $f(x)$ is a function that maps a percent reduction in potentially infectious contacts to a monetary value.
-At the time of the intervention, decision-makers do not have precise information about the costs of interventions $\tau$ or its costs represented by function $f(x)$.
+where $f(x)$ is a function that maps a percent reduction in potentially infectious contacts to a monetary cost.
+At the time of the intervention, decision-makers do not have precise information on either the effects of interventions (expressed in our model as a percent reduction $100 * \tau$ in infectious contacts) or its costs represented by function $f(x)$.
+Even if public health decision makers could estimate the instantaneous costs of nonpharmaceutical interventions in real-time, they would not be able to observe the full costs of interventions that do not manifest during the time-frame of interventions (i.e., the life-time effect of school closures).
+Nevertheless, any attempt to value the benefits of surveillance systems require the specification of the costs of nonpharmaceutical interventions.
+
+## Net monetary benefit (NMB)
+
+The net monetary benefit of interventions can be expressed as
+
+$$
+E[NMB] = E[QALYs] * WTP - E[C_{npi}] - C_{surv} 
+$${#eq-nbm}
+
+where $QALYs$ represent the quality-adjusted life-years gained from non-pharmaceutical interventions, $WTP$ is the willingness-to-pay for one quality-adjusted life_year, $C_{npi}$ are the total costs of NPIs and $C_{surv}$ is the total cost of surveillance system employed.
+
+# Valuing Wastewater surveillance information
+
+In this analysis, WBS systems are not assumed to resolve uncertainty of an exogenous model parameter $\boldsymbol \theta$.
+If WBS systems did resolve those uncertainties, then it would be possible to estimate the value of information (VOI) provided by wastewater surveillance systems using an expected value of partial perfect information (EVPPI) approach.
+Rather, we assume that WBS systems change what signals are available to policymakers and hence *how* they respond to an endogenous model quantity.
+Hence, we define the value of wastewater surveillance information as simply the difference between the net monetary benefit of with and without surveillance
+
+$$
+V = E[NMB|WBS=1] - E[NMB|WBS=0]
+$$ {#eq-wbsvalue}
 
@@
 ```
-# Tables
+
+# Results
+
+What follows is an outline of the results section.
+
+- Disease Elimination Scenarios
+
+First, we demonstrate the value of genomic surveillance scenarios in conditions under which temporary disease elimination is possible.
+These scenarios represent disease dynamics observed in small geographic regions (cite https://www.nejm.org/doi/full/10.1056/NEJMc2025203 New Zealand).
+
+{Figure with model dynamics with and without genomic surveillance system}
+
+- Disease Control Scenarios
+
+Early-warning systems can also be valuable in jurisdictions that do not achieve disease elimination.
+
+# Appendix
+
+## Tables
 

### 2023-05-12 paper/model-documentation.qmd
updating model documentation, trying out stochastic implementation with odin

@@
 -->
+
 # Methods

### 2023-05-12 R/scripts/00_metapopulation_no_deaths_stochastic.R

@@ -1,0 +1,183 @@
+
+
+# Metapopulation model implemented using the odin package.
+# See https://mrc-ide.github.io/odin/
+# for documentation on using odin
+# And
+# for examples of models implemented using odin
+# This model is inspired in http://epirecip.es/epicookbook/chapters/lloydjansen2004/r_odin
+# Which provides an implementation of a metapopulation model
+# with coupling and/or migration between patches
+# https://www.sciencedirect.com/science/article/abs/pii/S0025556403001809?via%3Dihub
+
+library(odin)
+
+# Helper functions:
+
+SEIR_cont <- odin::odin({
+  nr_patches <- user(2)
+  n <- nr_patches
+
+  ## Params
+  # Assignments in arrays are translated by odin to for loops in C
+  # See https://mrc-ide.github.io/odin/articles/functions.html
+
+
+  # Disease transmission equation
+  lambda_prod[ , ] <- beta[i, j] * (I[j] + P[j])
+  lambda[] <- sum(lambda_prod[i, ]) # rowSums
+
+  # This is the probability of infection | susceptible
+  p_SI[] <- 1 - exp(-lambda[i]/N[i])
+
+
+
+  # Mobility equations
+  # mob_prod object is re-used for each disease status.
+
+  # Calculate the number of people leaving each state
+  mob_prod[ , ] <- S[i] * C[i, j]
+
+  # and compute net flow of mobility to S
+  mob_S[] <- sum(mob_prod[, i])     # colSums
+
+  mob_prod[ , ] <- E[i] * C[i, j]
+  mob_E[] <- sum(mob_prod[, i])
+
+  mob_prod[ , ] <- P[i] * C[i, j]
+  mob_P[] <- sum(mob_prod[, i])
+
+  mob_prod[ , ] <- I[i] * C[i, j]
+  mob_I[] <- sum(mob_prod[, i])
+
+  mob_prod[ , ] <- R[i] * C[i, j]
+  mob_R[] <- sum(mob_prod[, i])
+
+  N[] <- S[i] + E[i] + P[i] + I[i] + R[i]
+
+  # output desired variables for debugging
+  output(N[]) <- TRUE
+  output(lambda_prod[]) <- TRUE
+  output(lambda[]) <- TRUE
+  output(S_E[]) <- TRUE
+  output(mob_prod[]) <- TRUE
+
+  ## Epidemiological Flows
[…]

### 2023-05-12 R/scripts/00_metapopulation_no_deaths_stochastic.R
renaming p_SE

@@ -28,7 +28,7 @@ SEIR_cont <- odin::odin({
   lambda[] <- sum(lambda_prod[i, ]) # rowSums
 
   # This is the probability of infection | susceptible
-  p_SI[] <- 1 - exp(-lambda[i]/N[i])
+  p_SE[] <- 1 - exp(-lambda[i]/N[i])
 
 
 
@@ -63,7 +63,7 @@ SEIR_cont <- odin::odin({
   output(mob_prod[]) <- TRUE
 
   ## Epidemiological Flows
-  S_E[] <- rbinom(S[i], p_SI[i]) # S[i] * lambda[i]
+  S_E[] <- rbinom(S[i], p_SE[i]) # S[i] * lambda[i]
   E_P[] <- rbinom(E[i], 1-exp(-sigma))
   P_I[] <- rbinom(P[i], 1-exp(-delta))
   I_R[] <- rbinom(I[i], 1-exp(-gamma))
@@ -112,7 +112,7 @@ SEIR_cont <- odin::odin({
   dim(E_P)         <- n
   dim(P_I)         <- n
   dim(I_R)         <- n
-  dim(p_SI)         <- n
+  dim(p_SE)        <- n
 })
 
 

### 2023-05-15 paper/manuscript.qmd
updating manuscript

--- 2023-05-12 paper/model-documentation.qmd
@@
       model: 3p
-      cite-style: super
-  docx:
       cite-style: super
@@
 
+<!--
+### Discrete-time stochastic implementation
+
+The model described above is implemented as a discrete-time stochastic model using the following approach.
+We simulate outflows for each disease state vector $\mathbf{X}$ using a stochastic  $\mathbf{X_t} = \mathbf{X_{t-1}} \sim Binomial(\mathbf{X}, 1-exp(\mathbf{p}))$ hence.
+
+-->
+
 ### Nonpharmaceutical interventions {#npi}
@@
 where $f(x)$ is a function that maps a percent reduction in potentially infectious contacts to a monetary cost.
+$f(x)$ is monotonically increasing with $\mathbf{l}_t\tau$, and may exhibity decreasing marginal returns to a reduction in transmissibility.
 At the time of the intervention, decision-makers do not have precise information on either the effects of interventions (expressed in our model as a percent reduction $100 * \tau$ in infectious contacts) or its costs represented by function $f(x)$.
 Even if public health decision makers could estimate the instantaneous costs of nonpharmaceutical interventions in real-time, they would not be able to observe the full costs of interventions that do not manifest during the time-frame of interventions (i.e., the life-time effect of school closures).
-Nevertheless, any attempt to value the benefits of surveillance systems require the specification of the costs of nonpharmaceutical interventions.
+Nevertheless, any attempt to value the benefits of surveillance systems requires the specification of the costs of nonpharmaceutical interventions.
 
@@
 
+<!--
+
 # Valuing Wastewater surveillance information
@@
 $$ {#eq-wbsvalue}
+
+-->
 

### 2023-05-16 paper/manuscript.qmd
updating manuscript

@@
 
-[^1]: We start with a simplified version including only a minimal set of disease states before expanding the state-space of the model to include asymptomatic, mildy symptomatic and severely symptomatic compartments
+[^1]: This model specification includes only a minimal set of disease states before expanding the state-space of the model to include asymptomatic, mildly symptomatic and severely symptomatic compartments.
 
@@
 where $\dot{\mathbf{S}}, \dot{\mathbf{E}}, \dot{\mathbf{P}}, \dot{\mathbf{I}}, \dot{\mathbf{R}}$ represent the time derivatives of each disease status, greek letters represent progression rates and $p_x\mathbf{m}_x$ terms represent migration flows $\mathbf{m}_x$ modulated by propensity to travel $p_x$ at disease state $x \in \{S, E, P, I, R\}$.
-@fig-model-diagram illustrates the model.
-
-```{mermaid}
-%%| label: fig-model-diagram
-%%| fig-cap: Model diagram
-%%| fig-width: 6.5
-%%| fig-align: center
-flowchart LR
-  S --> E --> P --> I --> R
-  I --> D
-```
 

### 2023-05-23 R/odin_models/stochastic_SIR_NPIs.R
implementing NPIs with a lagged outcome

@@ -68,7 +68,7 @@ update(Time) <- Time + 1
 # nonpharmaceutical interventions -----------------------------------------
 
 # lagged epidemiological outcome to use for controlling NPIs.
-Ilag <- delay(I, 5)
+Ilag <- delay(I, 20)
 
 # target intervention level depends on prevalence
 # this might be modified to better represent alternative surveillance methods

### 2023-05-23 R/odin_models/stochastic_metapopulation.R

@@ -34,7 +34,10 @@ delta   <- 1/6      # rate of progression from pre-symptomatic to Infected
 gamma   <- 1/3      # rate of recovery from active disease
 C[,]    <- user()   # origin-destination matrix of proportion of population that travels
 mp[]     <- user()  # relative migration propensity by disease status
-
+tau     <- user(0.1)     # policy effectiveness
+c       <- user(1) # stringency
+obs_lag <- user(5) # days
+days_to_adjust_NPI <- user(7)
 
 # initial conditions ------------------------------------------------------
 
@@ -43,6 +46,7 @@ initial(E[]) <- 0.0
 initial(P[]) <- 0.0
 initial(I[]) <- 1
 initial(R[]) <- 0.0
+initial(L[]) <- 0
 
 
 # dimensions --------------------------------------------------------------
@@ -69,6 +73,9 @@ dim(E_P)         <- n
 dim(P_I)         <- n
 dim(I_R)         <- n
 dim(p_SE)        <- n
+dim(L)           <- n
+dim(L_star)      <- n
+dim(I_lag)       <- n
 
 # additional outputs ------------------------------------------------------
 
@@ -80,8 +87,17 @@ output(mob_prod[]) <- TRUE
 
 # equations ---------------------------------------------------------------
 
+# nonpharmaceutical interventions -----------------------------------------
+
+I_lag[] <- delay(I[i], obs_lag) # i'm not sure this works
+output(I_lag) <- TRUE
+
+L_star[] <- c*I_lag[i] # implies same stringency for everyone
+
+update(L[]) <- L[i] + (L_star[i] - L[i]) / days_to_adjust_NPI
+
 # Disease transmission equation
-lambda_prod[ , ] <- beta[i, j] * (I[j] + P[j])
+lambda_prod[ , ] <- (1-L[i]*tau) * beta[i, j] * (I[j] + P[j])
 lambda[] <- sum(lambda_prod[i, ]) # rowSums
 
 # This is the probability of infection | susceptible
@@ -108,6 +124,10 @@ mob_prod[ , ] <- R[i] * C[i, j]
 mob_R[] <- sum(mob_prod[, i])
 
 N[] <- S[i] + E[i] + P[i] + I[i] + R[i]
+
+
+
+
 
 # difference equations ----------------------------------------------------
 

### 2023-05-23 R/scripts/00_test_models.R

@@ -21,7 +21,7 @@ set.seed(1234)
 stoc_SIR <- odinpbm$new("stochastic_SIR_NPIs.R", stringency = 1)
 
 # run model in first time-step
-stoc_SIR_res <- stoc_SIR$o$run(1:100)
+stoc_SIR_res <- stoc_SIR$o$run(1:700)
 
 
 head(stoc_SIR_res)

### 2023-05-31 paper/manuscript.qmd
implementing NPIs with a lagged outcome
writing up travel and stochastic implementation equations

@@
 \begin{align}
-\dot{\mathbf{S}} &=\ - \boldsymbol{\lambda}_t \odot \mathbf{S} + p_S \mathbf{m}_S \\
-\dot{\mathbf{E}} &=\ \boldsymbol{\lambda}_t \odot \mathbf{S} - \sigma \mathbf{E} + p_E \mathbf{m}_E \\
-\dot{\mathbf{P}} &=\ \sigma \mathbf{E} - \delta \mathbf{P} + p_P \mathbf{m}_P \\
-\dot{\mathbf{I}} &=\ \delta \mathbf{P} - \gamma \mathbf{I} + p_I \mathbf{m}_I \\
-\dot{\mathbf{R}} &=\ \gamma \mathbf{I} + p_R \mathbf{m}_R \\
+\dot{\mathbf{S}} &=\ - \boldsymbol{\lambda}_t \odot \mathbf{S} + \mathbf{m}_S \\
+\dot{\mathbf{E}} &=\ \boldsymbol{\lambda}_t \odot \mathbf{S} - \sigma \mathbf{E} + \mathbf{m}_E \\
+\dot{\mathbf{P}} &=\ \sigma \mathbf{E} - \delta \mathbf{P} + \mathbf{m}_P \\
+\dot{\mathbf{I}} &=\ \delta \mathbf{P} - \gamma \mathbf{I} + \mathbf{m}_I \\
+\dot{\mathbf{R}} &=\ \gamma \mathbf{I} + \mathbf{m}_R \\
 \end{align}
@@
 
-where $\dot{\mathbf{S}}, \dot{\mathbf{E}}, \dot{\mathbf{P}}, \dot{\mathbf{I}}, \dot{\mathbf{R}}$ represent the time derivatives of each disease status, greek letters represent progression rates and $p_x\mathbf{m}_x$ terms represent migration flows $\mathbf{m}_x$ modulated by propensity to travel $p_x$ at disease state $x \in \{S, E, P, I, R\}$.
+where $\dot{\mathbf{S}}, \dot{\mathbf{E}}, \dot{\mathbf{P}}, \dot{\mathbf{I}}, \dot{\mathbf{R}}$ represent the time derivatives of each disease status, greek letters represent progression rates and $\mathbf{m}_x$ terms represent net travel flows $\mathbf{m}_x$ at disease state $x \in \{S, E, P, I, R\}$.
 
@@
 
-<!--
 ### Discrete-time stochastic implementation
 
-The model described above is implemented as a discrete-time stochastic model using the following approach.
-We simulate outflows for each disease state vector $\mathbf{X}$ using a stochastic  $\mathbf{X_t} = \mathbf{X_{t-1}} \sim Binomial(\mathbf{X}, 1-exp(\mathbf{p}))$ hence.
+We implement the model as a discrete-time stochastic model using the following approach.
+We simulate outflows for each disease state vector $\mathbf{X}$ using drawing from a binomial distributions $\mathbf{out_{X_t}} \sim Binomial(\mathbf{X}, 1-exp(\mathbf{p}))$.
+We then carry over outflows from one disease state to the next.
+Hence, our disease state equations can be re-written as
 
--->
+$$ 
+\begin{align}
+\mathbf{S_t} &\sim \mathbf{S_{t-1}} 
+                    -B(\mathbf{S_{t-1}},1-\exp(-\boldsymbol{\lambda}_{t-1}/\mathbf{N}))
+                      + \mathbf{m_S} \\
+\mathbf{E_t} &\sim \mathbf{E_{t-1}} 
+                    +B(\mathbf{S_{t-1}},1-\exp(-\boldsymbol{\lambda}_{t-1}/\mathbf{N}))
+                    -B(E_{t-1},1-\exp(-\sigma)) 
+                    +\mathbf{m_E}\\
+\mathbf{P_t} &\sim \mathbf{P_{t-1}}
+                    +B(E_{t-1},1-\exp(-\sigma))
+                    -B(P_{t-1},1-\exp(-\delta)) +
+                    \mathbf{m_P}\\
+\mathbf{I_t} &\sim \mathbf{I_{t-1}}
+                    +B(P_{t-1},1-\exp(-\delta))
+                    -B(I_{t-1},1-\exp(-\gamma)) +
+                    \mathbf{m_I}\\
+\mathbf{R_t} &\sim \mathbf{R_{t-1}}
+                    -B(I_{t-1},1-\exp(-\gamma)) +
+                    \mathbf{m_R}
+\end{align}
+$$ {#eq-stochastic}
+
+where $B(\mathbf{n},p)$ results in a binomial draw from population vector $n$ with probability $p$.
+The model simulates outflows and inflows such that the outflows from one disease state are the same as the inflows to a different state.
+
+### Travel
+
+Stochastic travel flows are modeled as follows.
+Mobility matrix $\mathbf{A}$ is an $n \times n$ matrix where $a_{ij}$ represent daily travel rates from jurisdiction $i$ to $j$ and $a_{i=j}$ is the probability of no travel.
+Travel between jurisdictions is modeled using multinomial draws for each source jurisdiction $i$:
+
+$$
+\mathbf{X'_{td}} \sim \sum_{j}{Multinom(\mathbf{X_{t-1}}, A_i )} \ \forall \ d \in \{S, E, P, I, R\}
+$$ {#eq-travel}
+where $\mathbf{X'_{td}}$ are disease state vectors after travel.
+Once travel is computed in the model, we back out net travel flows as $\mathbf{m_d} = \mathbf{X'_{td}} - \mathbf{X_{td}}$ and use those in @eq-stochastic.
+This formulation implies that travel is simulated before disease transitions *within* each time step.
 
@@
 $$ 
-\boldsymbol{\beta}_t = (1-\mathbf{l}_t\tau) \boldsymbol{\beta}_b
+\boldsymbol{\beta}_t = (1-\mathbf{L}_t\tau) \boldsymbol{\beta}_b
 $$ {#eq-betat}
@@
 $$ 
-\mathbf{l}_t^* = \min{(cI, l_{max})}
+\mathbf{L}_t^* = \min{(cI_{t-\Delta}, l_{max})}
 $$ {#eq-targetnpi}
-
+where $\Delta$ is the delay imposed by the surveillance method used to monitor the disease.
 This formulation approximates phased mitigation plans adopted during the COVID-19 pandemic, where policymakers introduce or remove interventions depending on disease prevalence or other severity indicators.
@@
 
-where $a$ represents the adjustment time for NPIs (i.e., one or two weeks) using conventional measures.
+where $a$ represents the adjustment time for NPIs for decision making reasons, whereas $\Delta$ represents the lag imposed by surveillance methods.
 

### 2023-05-31 R/dev/odin_stochatic_travel.R

@@ -27,7 +27,7 @@ t_final <- 500
 # Diagonal (proportion who don't travel is set to zero).
 # m_ij: proportion of population that travels from row i to column j
 # everyone has a daily chance of traveling ~ 1%
-M <- matrix(data = 0.001,
+M <- matrix(data = 0.01,
             byrow = T, nrow = length(P), ncol = length(P), dimnames = list(loc_names, loc_names))
 
 diag(M) <- 0
@@ -136,6 +136,4 @@ res %>%
   geom_line(mapping = aes(y = `P[2]`), color = "red") +
   geom_line(mapping = aes(y = `P[3]`), color = "green") +
   geom_hline(yintercept = 0)
-
-
 

### 2023-05-31 paper/manuscript.qmd
updates manuscript

@@
 ---
-title: Modeling costs and benefits of genomic wastewater surveillance
-# subtitle: Supplementary appendix draft
+title: On the value of genomic wastewater surveillance information for pandemic response
+subtitle: A model-based analysis
 #author:
@@
 abstract: |
-  Introduces a model to assess the value provided by wastewater genomic surveillance network. This document is meant to be used as a supplementary appendix of the paper.
+  While wastewater-based genomic surveillance (WBS) systems may provide an early warning system for pandemic response, its adoption is still incomplete and it is unclear whether public health departments should allocate a larger budget to those systems.
+This paper introduces a model to assess the value provided by wastewater genomic surveillance network and clarifies the conditions under which WBS systems provide the most value to public health decision makers.
 keywords: 
@@
   - Pandemic Response
-date: last-modified
+# date: last-modified
 bibliography: bibliography.bib
@@
 This paper develops a stochastic metapopulation disease transmission model and uses it to assess the conditions under which WBS systems can provide value to local public health decision makers who respond to a pandemic using surveillance data.
+In the model, WBS systems help decision-makers by reducing the lag between the latent disease prevalence and the epidemiological time-series (i.e., case counts) decision-makers often use to introduce interventions targeted to reduce disease transmission.
 We evaluate a range of scenarios that represent different potential pathogens and highlight scenarios under which WBS systems may or may not be cost-effective.
@@
 
-We start with a standard disease transmission model describing person-to-person transmission of an infectious pathogen across $n$ jurisdictions.
+Our analysis starts with a standard disease transmission model describing person-to-person transmission of an infectious pathogen across $n$ jurisdictions.
 The state-space of the model is comprised of a set of health statuses $\mathbf{S}, \mathbf{E}, \mathbf{P}, \mathbf{I}, \mathbf{R}$, each a $1 \times n$ vector representing the population within each compartment, respectively representing the number of Susceptibles, Exposed, Pre-Symptomatic, Infected and Removed[^1].
@@
 $$ 
-\mathbf{L}_t^* = \min{(cI_{t-\Delta}, l_{max})}
+\mathbf{L}_t^* = \min{(cI_{t-s_{lag}}, l_{max})}
 $$ {#eq-targetnpi}
-where $\Delta$ is the delay imposed by the surveillance method used to monitor the disease.
+where $s_{lag}$ is a surveillance lag imposed by the surveillance method used to monitor the disease.
 This formulation approximates phased mitigation plans adopted during the COVID-19 pandemic, where policymakers introduce or remove interventions depending on disease prevalence or other severity indicators.
@@
 
-where $a$ represents the adjustment time for NPIs for decision making reasons, whereas $\Delta$ represents the lag imposed by surveillance methods.
+where $a$ represents the adjustment time for NPIs for decision making and implementation reasons, whereas $s_{lag}$ represents the lag imposed by surveillance methods.
 

### 2023-06-13 paper/manuscript.qmd
Fix n in binomial declaration explaination to be a vector rather than a number
updating model, sample post-processing function

@@
 ---
-title: On the value of genomic wastewater surveillance information for pandemic response
+title: On the value of early warning systems for outbreak response
 subtitle: A model-based analysis
@@
       theme: neutral
+  docx:
+      keep-tex: true
 ---
@@
 
-where $B(\mathbf{n},p)$ results in a binomial draw from population vector $n$ with probability $p$.
+where $B(\mathbf{n},p)$ results in a binomial draw from population vector $\mathbf{n}$ with probability $p$.
 The model simulates outflows and inflows such that the outflows from one disease state are the same as the inflows to a different state.
@@
 
-## Net monetary benefit (NMB)
+## Net monetary benefit (NMB) of the surveillance system
 
-The net monetary benefit of interventions can be expressed as
+The total expected cost the outbreak for a given set of parameters $\mathbf{\theta}$ and surveillance system $i$ can be expressed as the sum of health costs $E[CH_{i}|\mathbf{\theta}]$, nonpharmaceutical intervention costs $E[CNPI_{ti}|\mathbf{\theta}]$ and surveillance costs $CSURV_{it}|\mathbf{\theta}$, where the expectation is taken across stochastic model realizations.
+Note that we chose to condition the expected cost of the pandemic on $\mathbf{\theta}$ rather than assuming a probability distribution over it.
 
 $$
-E[NMB] = E[QALYs] * WTP - E[C_{npi}] - C_{surv} 
-$${#eq-nbm}
+E[C_{i}|\mathbf{\theta}] = E[CH_{i}|\mathbf{\theta}] + \sum_t{E[CNPI_{ti}|\mathbf{\theta}] + CSURV_{it}|\mathbf{\theta} } 
+$${#eq-costs}
 
-where $QALYs$ represent the quality-adjusted life-years gained from non-pharmaceutical interventions, $WTP$ is the willingness-to-pay for one quality-adjusted life_year, $C_{npi}$ are the total costs of NPIs and $C_{surv}$ is the total cost of surveillance system employed.
+Health costs $CH_{i}$ are computed as a function of number of removed individuals $R$ in the simulation
+$$
+E[CH_{i}|\mathbf{\theta}] = E[R(rw+o) |\mathbf{\theta}]
+$${#eq-costs}
+where $r$ is the infection fatality rate, $w$ is the average cost per death (assuming a set willingness to pay), and $o$ are other costs per infection, including days absent from work and any potential long-term disability.
 
-<!--
-
-# Valuing Wastewater surveillance information
-
-In this analysis, WBS systems are not assumed to resolve uncertainty of an exogenous model parameter $\boldsymbol \theta$.
-If WBS systems did resolve those uncertainties, then it would be possible to estimate the value of information (VOI) provided by wastewater surveillance systems using an expected value of partial perfect information (EVPPI) approach.
-Rather, we assume that WBS systems change what signals are available to policymakers and hence *how* they respond to an endogenous model quantity.
-Hence, we define the value of wastewater surveillance information as simply the difference between the net monetary benefit of with and without surveillance
+In this analysis, WBS systems are not assumed not resolve uncertainty of any exogenous model parameter in $\boldsymbol \theta$.
+If WBS systems did resolve at least one of those uncertainties,  it would be possible to estimate the value of information (VOI) provided by wastewater surveillance systems using an expected value of partial perfect information (EVPPI) method.
+Rather, we assume that WBS systems change what signals are available to policymakers and hence *how* they respond to the pandemic.
+Hence, we define the net monetary benefit of wastewater surveillance information as the difference between the total cost of the pandemic with and without WBS surveillance.
 
 $$
-V = E[NMB|WBS=1] - E[NMB|WBS=0]
+NMB_{surv} |\mathbf{\theta} = E[C_{conv surv}|\mathbf{\theta}]  - E[C_{meta surv}|\mathbf{\theta}]
 $$ {#eq-wbsvalue}
-
--->
 

### 2023-06-13 R/functions/odinmetapopulation.R

@@ -28,6 +28,32 @@ odinmetapop <- R6::R6Class(
       inputs$nr_patches <- 2
       #inputs$mp <- c(1, 1, 0.5, 1, 1)
       return(inputs)
+    },
+
+    #' Computes replication-level summaries based on model results
+    #'
+    #' @export
+    post_process = function() {
+
+      # hard-coded for first jurisdiction
+      # Sarah to generalize
+      # Compute time-varying costs:
+      self$res <- self$res %>%
+        # This is a linear function from now, it can be non-linear
+        mutate(CNPI = `L[1]` * self$oi$tau * self$inputs$jurisdiction$cost.npi[1])
+
+      # This is where we summarize costs:
+      self$summary <- self$res %>%
+        group_by(rep) %>%
+        summarise(CNPI = sum(CNPI),
+                  R = max(`R[1]`)) %>%
+        # Compute other costs:
+        mutate(CH = R * (self$oi$r*self$oi$w + self$oi$o),
+               CSURV = self$oi$C_surv,
+               C = CH + CSURV + CNPI)
+
     }
+
+
   )
 )

### 2023-06-13 R/functions/odinpbm.R

@@ -22,6 +22,15 @@ odinpbm <- R6::R6Class(
     #' @field o odin model object
     o = NULL,
 
+    #' @field oi odin model inputs
+    oi = NULL,
+
+    #' @field res day-level results
+    res = NULL,
+
+    #' @field res rep-level results
+    summary = NULL,
+
     #' @description
     #' Create a new `odinpbm` object.
     #' @param odin_file odin model file under R/odin_models. Should have an .R extension.
@@ -46,7 +55,7 @@ odinpbm <- R6::R6Class(
       inputs <- self$collect_default_inputs()
 
       # add custom inputs and override original inputs by name
-      inputs <- modifyList(inputs, list(...))
+      self$oi <- modifyList(inputs, list(...))
 
       # create the odin constructor
       # We don't use the constructor after this step, so I don't save it
@@ -55,7 +64,7 @@ odinpbm <- R6::R6Class(
       # initialize the model with default parameters, and additional parameters
       # we may want to provide additional wrappers around other odin functions
       # in the future
-      self$o <- do.call(odin_constructor$new, inputs)
+      self$o <- do.call(odin_constructor$new, self$oi)
 
     },
 
@@ -69,7 +78,7 @@ odinpbm <- R6::R6Class(
       res_list <- lapply(seq(dim(res)[3]), function(x) res[ , , x] %>% as.data.frame(.) %>% mutate(rep = x))
 
       # return as a data.frame
-      do.call(rbind, res_list)
+      self$res <- do.call(rbind, res_list)
 
     },
 

### 2023-06-13 R/scripts/00_test_models.R

@@ -14,13 +14,28 @@ source("./R/library.R")
 
 # stochastic metapopulation model -----------------------------------------
 
-meta_SIR_stoc <- odinmetapop$new("stochastic_metapopulation.R", s$data_file)
+meta_SIR_stoc <- odinmetapop$new("stochastic_metapopulation.R", s$data_file,obs_lag = 5)
 
-meta_SIR_stoc_res <- meta_SIR_stoc$run(0:100, reps = 100)
+meta_SIR_stoc$run(0:100, reps = 100)
+
+meta_SIR_stoc$post_process()
+
+
+summary(meta_SIR_stoc$summary)
 
 View(meta_SIR_stoc_res)
 
+
+
+
 head(meta_SIR_stoc_res)
+
+
+
+
+
+
+
 
 # deterministic metapopulation ode ----------------------------------------
 
@@ -60,9 +75,9 @@ rbind(res_2_delay, res_5_delay, res_10_delay) %>%
   mutate(TotalCases = cumsum(I),
          TotalNPICost = cumsum(NPI) * 1) %>% # Where 1 is the NPI cost
   tidyr::pivot_longer(cols = c(S,I,R,NPI,TotalCases,TotalNPICost)) %>%
-  filter(name %in% c("I", "NPI")) %>%
+  filter(name %in% c("I")) %>%
   ggplot(mapping = aes(x = step, y = value, color = Scenario, group = rep)) +
   geom_line() +
   xlab("Days") +
-  facet_wrap(facets = ~name + Scenario, scales = "free")
+  facet_wrap(facets = ~name + Scenario)
 

### 2023-09-14 paper/manuscript.qmd
WIP: Update to manuscript

@@
 This paper develops a stochastic metapopulation disease transmission model and uses it to assess the conditions under which WBS systems can provide value to local public health decision makers who respond to a pandemic using surveillance data.
-In the model, WBS systems help decision-makers by reducing the lag between the latent disease prevalence and the epidemiological time-series (i.e., case counts) decision-makers often use to introduce interventions targeted to reduce disease transmission.
+In the model, WBS systems help decision-makers by reducing the lag between latent disease prevalence and the epidemiological time-series (i.e., case counts) decision-makers often use to introduce interventions targeted to reduce disease transmission.
 We evaluate a range of scenarios that represent different potential pathogens and highlight scenarios under which WBS systems may or may not be cost-effective.
@@
 Our analysis starts with a standard disease transmission model describing person-to-person transmission of an infectious pathogen across $n$ jurisdictions.
-The state-space of the model is comprised of a set of health statuses $\mathbf{S}, \mathbf{E}, \mathbf{P}, \mathbf{I}, \mathbf{R}$, each a $1 \times n$ vector representing the population within each compartment, respectively representing the number of Susceptibles, Exposed, Pre-Symptomatic, Infected and Removed[^1].
+The state-space of the model is comprised of a set of health statuses $\mathbf{S}, \mathbf{E}, \mathbf{P}, \mathbf{I}, \mathbf{R}$, each a $1 \times n$ vector representing the population in each of the following compartments: Susceptible, Exposed, Pre-Symptomatic, Infected and Removed, respectively[^1].
 @eq-ode1 describes a system of differential equations representing the model, and @tbl-parameters contains parameter definitions, values and their sources.
@@
 
+### Information sharing and coordination
+
+During pandemics in the past, some jurisdictions have adjusted their NPI levels based on not only on local disease indicators, but based on the behavior and disease indicators of neighboring jurisdictions as well.
+For example, countries have restricted travel or required traveler quarantines for those arriving from areas with high disease prevalence.
+Additionally, they may coordinate efforts with other nearby cities, adding and lifting mandates in unison to reduce confusion, anticipate travel-based spread, and present unified messaging.
+
+To simulate behavior, we have allowed for jurisdictions to coordinate and synchronize NPIs levels with other jurisdictions. They may choose to:
+1. Pick their NPI intervention level based exclusively on their own wastewater or other local indicators (their own target NPI level absent any coordination)
+2. Set their NPI intervention level based on the average target NPI level across their own and their neighboring jurisdictions.
+3. Set their NPI intervention level to the maximum indicated intervention level across all jurisdictions with which they coordinate.
+
+# CITATION HERE
+### Time varying infection fatality rate
+
+During the course of a pandemic, the fatality rate of infection may vary for several reasons. We examine two of these:
+1. Health system capacity limitations
+2. Improved treatment (evolving standards of care)
+
+#### Health System Capacity Limitations: Fluctuating Infection Fatality Rates
+
+Health system capacity limitations were heavily discussed during the initial months of the COVID-19 pandemic.
+These calls to "flatten the curve" concentrated not just on limiting overall infection counts, but on ensuring that the need for medical resources, did not exceed supply.
+https://www.nytimes.com/2020/03/18/business/coronavirus-ventilator-shortage.html
+https://www.healthaffairs.org/doi/full/10.1377/hlthaff.2022.00643
+
+We have thus added the following limitation to our model; when the number of patients in need of hospital care exceeds the available supply (e.g.
+if there is an insufficient supply of supplemental oxygen, ventilators, extracorporeal membrane oxygenation machines, and medical staff to meet demand), the case fatality rate among these patients increases.
+
+#### Evolving Standards of Care: Decreasing Infection Fatality Rates
+
+During the course of a pandemic, providers may learn more about the pathogen and illness, developing improved standards of care, even absent pharmaceudical intervention.
+This has been the case for treatment of hypoxemia and acute respiratory distress syndrome (ARDS) during the COVID-19 pandemic.
+https://pubmed.ncbi.nlm.nih.gov/33156382/ For example, initually, patients suffering from hypoxemia were usually laid in a supine position with "aggressive airway management and intubation".
+However, in the Spring of 2020, evidence mounted that initially simply placing patients in a prone, rather than supine position, as well as less aggressive intubation, led to improved outcomes.
+https://nap.nationalacademies.org/cart/download.cgi?record_id=26573 pg 7 https://pubmed.ncbi.nlm.nih.gov/32569585/.
+Implementation of the new protocols, likely resulted in lower infection fatality rates and health system burden.
+https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7253105/
+https://pubmed.ncbi.nlm.nih.gov/37016834/
+
+# CITATION HERE
+
 ## Cost of interventions

### 2024-02-16 paper/results.qmd
cleaning up and documenting summary functions

---
title: "results"
format: html
---

# Reproducibility notes

This quarto document reproduces the results section in Nascimento de Lima et al. (2024). *The value of environmental surveillance for pandemic response*.

# Results

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2024-02-16 reproduce-results.qmd
cleaning up and documenting summary functions

--- 2024-02-16 paper/results.qmd
@@
 format: html
+author: "Pedro Nascimento de Lima"
 ---
@@
 
-This quarto document reproduces the results section in Nascimento de Lima et al. (2024). *The value of environmental surveillance for pandemic response*.
+This quarto document reproduces the results section in Nascimento de Lima et al.
+(2024).
+*The value of environmental surveillance for pandemic response*.
+For details on how to install dependencies, and reproduce these results, see the `README.md` file.
+
+# Run Results
+
+First, we instantiate the model and run it over the experimental designs used in the paper.
+
+```{r}
+# Source all dependencies and model scripts
+source("./R/library.R")
+
+# Single run examples -----------------------------------------------------
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
+
+# Verify that model runs with base-case parameters, no NPIs
+
+# Set NPI case threshold to 1000 (i.e., do not introduce NPIs)
+model$set_input(name = "c", value = Inf)
+test_results <- model$simulate()
+
+# The res_long data.frame contains jurisdiction-day-replication-level results
+
+head(model$res_long)
+```
 

### 2024-02-16 R/dev/edge_cases_analysis.R

@@ -24,7 +24,7 @@ model$set_input("c", 10000)$
   set_input("R0", 2.5)$
   set_input("p", 0.5)$
   set_input("tau", 0.18)$
-  set_input("surv_lag", 1)$
+  set_input("total_surv_lag", 1)$
   set_input("cost_max_npi", 0.25)$
   set_input("L_c", 1)$
   set_input("a_up", 1)$

### 2024-02-16 R/dev/jurisdiction-mixing.R

@@ -7,12 +7,11 @@
 
 # First, we have a
 
-n.jurisdictions <- 2
-names.jurisdictions <- c("A", "B")
+names.jurisdictions <- c("A", "B", "C", "D", "E")
+n.jurisdictions <- length(names.jurisdictions)
 
-n.modes <- 3
 names.modes <- c("Home", "Work", "Other")
-
+n.modes <- length(names.modes)
 # Following the same notation, k is the normalized contact rates across modes, by jurisdictions
 # k can be thought of the intensity of contacts across these mixing modes for each jurisdiction
 

### 2024-02-16 R/dev/policy_profile_analysis.R

@@ -91,7 +91,7 @@ fig_prevalence_fan <- all_long_df %>%
 
 fig_deaths_fan <- all_long_df %>%
   filter(jurisdiction.name == "NY") %>%
-  ggplot(mapping = aes(x = step, y = Deaths.per.100k)) +
+  ggplot(mapping = aes(x = step, y = deaths_per_100k)) +
   ggfan::geom_fan() +
   facet_wrap(~policy_profile, nrow = 1) +
   ylab("Deaths per 100,000 people") +

### 2024-02-16 R/functions/OdinMetapop.R

@@ -84,10 +84,8 @@ OdinMetapop <- R6::R6Class(
 
       }
 
-      # Concatenate all mixing mdoes into one array (order matters!):
+      # Concatenate all mixing modes into one array (order matters!):
       M <- array(c(home_mixing, work_travel_mixing, other_mixing),dim = c(n.jurisdictions, n.jurisdictions, n.modes), dimnames = list(names.jurisdictions, names.jurisdictions, names.modes))
-
-      # This is the same as Mij.
 
       # Overall mixing matrix is the sum of element-wise multiplications:
       mixing_matrix <- matrix(data = 0, nrow = n.jurisdictions, ncol = n.jurisdictions,dimnames = list(names.jurisdictions, names.jurisdictions))
@@ -105,10 +103,6 @@ OdinMetapop <- R6::R6Class(
       # Now that we have Mij, we need to align it with a specified R0.
 
       # This approach does so by assuming that cbeta is uniform across juridictions.
-
-      # Simplistic approach: Results match, but R0 may not match when population sizes
-
-      # are different.
 
       tau.eff <- (1/self$inputs$delta + 1/self$inputs$gamma)
 
@@ -230,8 +224,8 @@ OdinMetapop <- R6::R6Class(
 
       # Based on time (i.e., due to improved standard of care)
       if(as.logical(self$inputs$time_varying_IFR)) {
-        # Scale parameter of the RR risk function
-        self$set_input("r_scale_factor", calib_logistic_fn(y_max = self$inputs$r_terminal_RR, x_mid_point = self$inputs$t_mid_IFR, x_trans = self$inputs$t_trans_IFR, x_vector = seq.default(from = 0, to = 365, by = 0.01)))
+        # Scale parameter of the logistic RR risk function
+        self$set_input("r_scale_factor", calib_logistic_fn(y_max = self$inputs$p_r_star, x_mid_point = self$inputs$t_mid_IFR, x_trans = self$inputs$t_r_star, x_vector = seq.default(from = 0, to = 365, by = 0.01)))
       }
 
       # Based on prevalence (i.e., hospital utilization)
@@ -251,7 +245,38 @@ OdinMetapop <- R6::R6Class(
     post_process = function() {
 
       # select only the variables we want for the summary variable
-      required_jurisdiction_variables = c("rep", "step", "L", "NPI", "S", "E", "P", "I", "R")
[…]

### 2024-02-16 R/functions/logistic_fn.R

@@ -110,14 +110,14 @@ calib_logistic_fn <- function(y_max, x_mid_point, x_trans, x_vector = seq.defaul
 
   solution <- optim(par = scale_guess, method = "Brent", # "L-BFGS-B",
                     fn = log_calib_obj_fn,
-                    lower = scale_guess * 0.2,
+                    lower = scale_guess * 0.1,
                     upper = scale_guess * 3,
                     y_max = y_max,
                     x_mid_point = x_mid_point,
                     x_trans = x_trans,
                     x_vector = x_vector)
 
-  print(solution)
+  #print(solution)
 
   solution$par
 

### 2024-02-16 R/odin_models/stochastic_metapopulation.R

@@ -32,11 +32,10 @@ rho   <- user()       # prop asymptomatic
 tau     <- user()     # policy marginal effectiveness
 c       <- user()     # policy stringency
 t_o <- user()         # Time of "opening" (i.e., where NPIs are no longer used)
-surv_lag <- user()    # case Observation lag (in days)
+total_surv_lag <- user()    # case Observation lag (in days)
 a_up <- user()           # NPI *a*djustment time
 a_down <- user()           # a decrease
 L_max <- user()       # max intervention level
-beta_mult <- user()   # transmissibility multiplier (used for scenario analysis)
 C[,] <- user()        # *C*: NPI coordination matrix. 1 if jurisdiction i follows NPI of jurisdiction j if jurisdiction's j NPI is more stringent.
 L_c <- user() # 1 if NPIs are coordinated across jurisdictions, 0 otherwise.
 p <- user() # case ascertainment proportion.
@@ -95,7 +94,7 @@ dim(change_NPI_now) <- n
 # NPIS & NPI Coordination
 
 # lagged_incidence is the epidemiological signal used to introduce interventions.
-lagged_incidence[] <- delay(rbinom(S_E[i], p), surv_lag)
+lagged_incidence[] <- delay(rbinom(S_E[i], p), total_surv_lag)
 
 # Effective NPI strigency: only active temporarily
 eff_c <- if(step <= t_o) c else 100000
@@ -131,7 +130,7 @@ update(NPI[]) <- if (change_NPI_now[i]) floor(L[i]) else NPI[i]
 
 # Disease transmission
 # Disease transmission equation
-lambda_prod[ , ] <- beta_mult * (1-NPI[i]*tau) * beta[i, j] * ((P[j] + I[j] + A[j])/N[j])
+lambda_prod[ , ] <- (1-NPI[i]*tau) * beta[i, j] * ((P[j] + I[j] + A[j])/N[j])
 lambda[] <- sum(lambda_prod[i, ]) # rowSums
 
 # This is the probability of infection | susceptible
@@ -150,7 +149,7 @@ E_P[] <- rbinom(E[i], 1-exp(-sigma))
 P_IA[] <- rbinom(P[i], 1-exp(-delta))
 # draw symptomatic:
 P_I[] <- rbinom(P_IA[i], 1-rho)
-# Asymptomatic are those who remain:
[…]

### 2024-02-16 R/scripts/01_single_run_test.R

@@ -14,6 +14,14 @@ source("./R/library.R")
 model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
 #model2 <- OdinMetapop$new("stochastic_metapopulation2.R", "./data/metapopulation-inputs-heterogeneous.xlsx")
 
+model$set_input("c", 10)$
+  set_input("tau", 0.05)$
+  simulate()
+
+View(model$summary)
+
+
+
 # all default model inputs are in the inputs object
 model$inputs$beta
 # IT looks like the mixing is not being partitioned correctly!
@@ -21,7 +29,7 @@ model$inputs$beta
 
 # We can run the model with default inputs by using the simulate function
 # i.e., run for 100 days, for 100 replications
-model$simulate(step = 0:300, reps = 100)
+View(model$simulate(step = 0:300, reps = 100))
 
 
 

### 2024-02-16 R/scripts/02_sample_analysis.R

@@ -35,7 +35,7 @@ experiment <- R6Experiment$new(model)
 
 # Set experimental parameters
 experiment$
-  set_parameter("surv_lag", "grid", seq.default(from = 0, to = 21, by = 3))$ # Surveillance lag
+  set_parameter("total_surv_lag", "grid", seq.default(from = 0, to = 21, by = 3))$ # Surveillance lag
   set_parameter("p", "grid", c(0.5,1))$ # case ascertainment proportion
   set_parameter("tau", "grid", c(0.12,0.15,0.18))$ # intevention effectiveness per intervention level
   set_parameter("c", "grid", c(0.01,1,5,10,100))$ # case intervention threshold per 100,000 people

### 2024-02-16 R/scripts/03_run_larger_exp_design.R

@@ -42,15 +42,15 @@ policy_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx",sheet = "policy_pr
 
 
 # Other variables
-surv_lag <- seq.default(from = 2, to = 20, by = 2)
+total_surv_lag <- seq.default(from = 2, to = 20, by = 2)
 R0 <- c(1.5,2,3)
 cost_max_npi <- c(0.1,0.25,0.87, 1.25)
 L_c <- c(0,1)
 r <- c(0.01, 0.001, 0.02)
 
 # Surveillance design
-surv_design <- data.frame(surv_lag = surv_lag) %>%
-  mutate(p = ifelse(surv_lag < 10, 1, 0.5)) %>%
+surv_design <- data.frame(total_surv_lag = total_surv_lag) %>%
+  mutate(p = ifelse(total_surv_lag < 10, 1, 0.5)) %>%
   mutate(surv.id = row_number())
 
 # Other variables design:
@@ -69,7 +69,7 @@ full_design <- expand_grid(policy_scenarios, surv_design, other_vars_design)
 
 # Set experimental parameters
 # experiment$
-#   set_parameter("surv_lag", "grid", seq.default(from = 0, to = 21, by = 3))$ # Surveillance lag
+#   set_parameter("total_surv_lag", "grid", seq.default(from = 0, to = 21, by = 3))$ # Surveillance lag
 #   set_parameter("p", "grid", c(0.5,1))$ # case ascertainment proportion
 #   set_parameter("tau", "grid", c(0.12,0.15,0.18))$ # intevention effectiveness per intervention level
 #   set_parameter("c", "grid", c(0.01,1,5,10,100))$ # case intervention threshold per 100,000 people
@@ -87,7 +87,7 @@ exp_results <- experiment$run(parallel = T,cluster_eval_script = "./R/scripts/cl
 
 # Compute NMB counterfactuals:
 ref_costs <- exp_results %>%
-  filter(surv_lag == 10) %>%
+  filter(total_surv_lag == 10) %>%
   rename(ref_C = C) %>%
   select(policy.id, other.vars.id, ref_C)
 

### 2024-02-16 archive/set_ram.R

@@ -1,0 +1,20 @@
+
+
+
+# This script demonstrates how to run the model for a single run:
+
+source("./R/library.R")
+
+# Single run examples -----------------------------------------------------
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
+
+# all default model inputs are in the inputs object
+model$inputs
+
+# We can run the model with default inputs by using the simulate function
+# i.e., run for 100 days, for 100 replications
+model$simulate(step = 0:100, reps = 100)
+
+model$set_input("coordination_mode", 2)$
+  simulate()
+

### 2024-02-19 reproduce-results.qmd
supporting variant strains

--- 2024-02-16 paper/results.qmd
@@
 
-This quarto document reproduces the results section in Nascimento de Lima et al. (2024). *The value of environmental surveillance for pandemic response*.
+This quarto document reproduces the results section in Nascimento de Lima et al.
+(2024).
+*The value of environmental surveillance for pandemic response*.
+For details on how to install dependencies, and reproduce these results, see the `README.md` file.
+
+# Run Results
+
+First, we instantiate the model and run it over the experimental designs used in the paper.
+
+```{r}
+
+```
+
+# Results
+
+## Value of EWS under base-case results
+{Describe dynamics and results uder base-case assumptions}
+
+D. EWS provide meaningful value under base-case assumptions
+
+## Conditions for high-value early warning 
+{Describe conditions under which value is lower or higher}
+D. Value of EWS hinges on timeliness of and adherence to NPIs. 
+There was a
+
+## EWS and policy coordination
+D. EWS alleviate lack of policy coordination. 
+There is a concern that EWS data will need to be integrated to be useful. However, we find that policy coordination and EWS act as substitutes. Policy coordination h
+
+## Marginal Value of EWS
+D. EWS deployment also face decreasing marginal returns
+The value of EWS decreases as coverage increases.
 

### 2024-02-19 R/functions/OdinMetapop.R

@@ -113,7 +113,6 @@ OdinMetapop <- R6::R6Class(
       # C beta risk ratios are inputs
       c_rr <- self$inputs$jurisdiction$mixing.risk.ratio
 
-
       # This assumes that the first jurisdiction is the reference jurisdiction
       # From c = sum_i(c_i*pop_i), we can derive c1
       c_1 = cbeta / (pop[1] + sum(c_rr[2:self$inputs$nr_patches] * pop[2:self$inputs$nr_patches]))
@@ -140,6 +139,10 @@ OdinMetapop <- R6::R6Class(
       # R0 verification:
       # R0 should be this:
       R0_verif <- sum(self$inputs$beta %*% pop * tau.eff)
+
+      # Sampling new variant strain introdiction.
+
+      self$set_input("variant_t", round(runif(1, min = self$inputs$variant_min_t, max = self$inputs$variant_max_t)))
 
       stopifnot(abs(R0_verif - self$inputs$R0)/self$inputs$R0 < 1e-2)
 
@@ -275,7 +278,8 @@ OdinMetapop <- R6::R6Class(
         CH = "health costs per person",
         CSURV = "surveillance cost per person",
         C = "total pandemic cost per person",
-        epi_size = "final epidemic size"
+        epi_size = "final epidemic size",
+        L5_days = "days under level 5 NPI"
       )
 
       # save results into long format
@@ -324,16 +328,19 @@ OdinMetapop <- R6::R6Class(
         mutate(CH_deaths = deaths_per_100k * self$inputs$VSL / 10^5) %>%
         mutate(CH = CH_deaths + CH_illness,
                CSURV = self$inputs$C_surv,
-               C = CSURV + CH + CNPI)
+               C = CSURV + CH + CNPI) %>%
+        mutate(L5_days = as.integer(NPI == 5)) %>%
+        mutate(L1plus_days = as.integer(NPI >= 1))
 
       # Summarize costs by jurisdiction over time:
       self$summary_jurisdiction <- self$res_long %>%
-        select(rep, jurisdiction.id, population, new_removed, deaths_per_100k, CH_illness, CH_deaths, CH, CSURV, CNPI, C) %>%
+        select(rep, jurisdiction.id, population, new_removed, deaths_per_100k, CH_illness, CH_deaths, CH, CSURV, CNPI, C, L5_days) %>%
         mutate(epi_size = new_removed / population) %>%
         select(-c(new_removed, population)) %>%
         group_by(rep, jurisdiction.id) %>%
[…]

### 2024-02-19 R/functions/OdinSim.R

@@ -44,16 +44,17 @@ OdinSim <- R6::R6Class(
     #' Create a new `OdinSim` object.
     #' @param odin_file odin model file under R/odin_models. Should have an .R extension.
     #' @param inputs_file spreadsheet with model inputs
+    #' @param model_name optional model name (useful when using the same odin model file)
     #' @param ... additional inputs to te odin model.
     #' @return s new `OdinSim` object.
-    initialize = function(odin_file, inputs_file, ...) {
+    initialize = function(odin_file, inputs_file, model_name = "", ...) {
 
       super$initialize(name = odin_file)
 
       model_path <- paste0("./R/odin_models/", odin_file)
 
-      odin_workdir <- paste0("./cpp/", substr(odin_file,start = 1,
-                                              stop = nchar(odin_file)-2), "/")
+      odin_workdir <- paste0("./cpp/", paste0(substr(odin_file,start = 1,
+                                                     stop = nchar(odin_file)-2), "_", model_name) , "/")
 
       # Get (and set) inputs from spreadsheet:
       self$get_inputs(inputs_file)
@@ -64,13 +65,8 @@ OdinSim <- R6::R6Class(
       # Pre-process inputs (which will also assign any new parameters to the odin model)
       self$pre_process_inputs()
 
-      #browser()
-
       # Create odin constructor (this is an R6 class):
       odin_constructor <- odin::odin(model_path, workdir = odin_workdir, debug_enable = F)
-
-      # trying to use odin.dust for parallel execution:
-      # odin_constructor <- odin.dust::odin_dust(model_path, workdir = odin_workdir, debug_enable = F)
 
       # Create list of odin inputs
       self$odin_parms <- odin_constructor$private_fields$user

### 2024-02-19 R/functions/summary_funs.R

@@ -1,0 +1,39 @@
+
+
+# Create summary functions list:
+# One can add more percentiles to the calculation by changing the vector below.
+p <- c(0.025, 0.975)
+p_names <- map_chr(p, ~paste0(".q.",.x*100))
+summary_functions = map(p, ~partial(quantile, probs = .x, na.rm = TRUE)) %>%
+  set_names(nm = p_names)
+summary_functions$.mean = mean
+
+# Get estimate in Text Form:
+
+get_estimate = function(summary_tbl, strategy, col.sens, model_name, outcome, sig_numbers = 3) {
+
+  r = summary_tbl %>%
+    filter(Strategy == strategy, model == model_name, sensitivity == col.sens) %>%
+    mutate(across(where(is.numeric), .fns = ~signif(x = .x, digits = sig_numbers)))
+
+  vars = paste0(outcome, "_.", c("mean", "q.2.5", "q.97.5"))
+
+  paste0(r[,vars[1]], " (95% PI [", r[,vars[2]], ", ", r[,vars[3]],"])")
+
+}
+
+get_outcome <- function(df, outcome_var, sig_digits = 3, ...) {
+  r <- df %>%
+    dplyr::filter(...) %>%
+    as.data.frame() %>%
+    mutate(across(where(is.numeric), .fns = ~signif(x = .x, digits = sig_digits)))
+
+  vars = paste0(outcome_var, "_.", c("mean", "q.2.5", "q.97.5"))
+
+  paste0(r[,vars[1]], " (", r[,vars[2]], "-", r[,vars[3]],")")
+
+}
+
+# get_outcome(df, "outcome", a == 2, b == 2)
+
+

### 2024-02-19 R/library.R

@@ -22,6 +22,7 @@ library(dde)
 library(officer)
 library(R6Sim)
 library(patchwork)
+library(purrr)
 
 # Load settings:
 s <- yaml::read_yaml("settings.yml")

### 2024-02-19 R/odin_models/stochastic_metapopulation.R

@@ -41,6 +41,8 @@ L_c <- user() # 1 if NPIs are coordinated across jurisdictions, 0 otherwise.
 p <- user() # case ascertainment proportion.
 S0[] <- user()
 I0[] <- user()
+variant_beta_rr <- user()
+variant_t <- user()
 
 # initial conditions ------------------------------------------------------
 
@@ -84,6 +86,7 @@ dim(L_star_ind) <- n # Jurisdiction's own L_star without considering other's
 dim(L_star_max) <- c(n,n) # Target NPI considering maximum NPI level of coordinating jurisdictions.
 dim(lagged_incidence) <- n
 dim(change_NPI_now) <- n
+#dim(variant_introduced) <- 1
 
 # additional outputs
 
@@ -128,9 +131,13 @@ change_NPI_now[] <- if(L[i] > NPI[i]) (step %% a_up) == 0 else (step %% a_down) 
 
 update(NPI[]) <- if (change_NPI_now[i]) floor(L[i]) else NPI[i]
 
+# Variant strain modified transmissibility
+
+variant_rr <- if(step >= variant_t) variant_beta_rr else 1
+
 # Disease transmission
 # Disease transmission equation
-lambda_prod[ , ] <- (1-NPI[i]*tau) * beta[i, j] * ((P[j] + I[j] + A[j])/N[j])
+lambda_prod[ , ] <- variant_rr * (1-NPI[i]*tau) * beta[i, j] * ((P[j] + I[j] + A[j])/N[j])
 lambda[] <- sum(lambda_prod[i, ]) # rowSums
 
 # This is the probability of infection | susceptible

### 2024-02-19 R/scripts/01_single_run_test.R

@@ -11,8 +11,11 @@
 source("./R/library.R")
 
 # Single run examples -----------------------------------------------------
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file, "model1")
 #model2 <- OdinMetapop$new("stochastic_metapopulation2.R", "./data/metapopulation-inputs-heterogeneous.xlsx")
+
+model$simulate()
+
 
 model$set_input("c", 10)$
   set_input("tau", 0.05)$
@@ -20,6 +23,8 @@ model$set_input("c", 10)$
 
 View(model$summary)
 
+
+get_outcome(model$summary, outcome_var = "CH")
 
 
 # all default model inputs are in the inputs object

### 2024-02-19 R/scripts/run_results.R

@@ -1,0 +1,32 @@
+
+# Source all dependencies and model scripts
+source("./R/library.R")
+
+
+# Base case runs
+
+# Single run examples -----------------------------------------------------
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
+
+model$set_input("c", 10^5)$set_input("tau", 0.15)
+
+model$simulate(reps = 500)
+
+get_outcome(model$summary, outcome_var = "L5_days", sig_digits = 3)
+
+get_outcome(model$summary, outcome_var = "epi_size", sig_digits = 3)
+
+get_outcome(model$summary, outcome_var = "deaths_per_100k", sig_digits = 3)
+
+
+library(ggplot2)
+
+model$res_long %>%
+  filter(jurisdiction.id == 1) %>%
+  ggplot() +
+  geom_line(mapping = aes(x = step, y = NPI, group = interaction(rep, jurisdiction.id) , color = as.factor(jurisdiction.id)))
+
+model$res_long %>%
+  filter(jurisdiction.id == 1) %>%
+  ggplot() +
+  geom_line(mapping = aes(x = step, y = I, group = interaction(rep, jurisdiction.id), color = as.factor(jurisdiction.id)))

### 2024-02-21 reproduce-results.qmd
beginning of results section

--- 2024-02-16 paper/results.qmd
@@
 ---
-title: "results"
+title: "The value of early warning systems for pandemic response"
 format: html
+execute: 
+  echo: false
+  warning: false
+  error: false
+author: "Pedro Nascimento de Lima"
 ---
@@
 
-This quarto document reproduces the results section in Nascimento de Lima et al. (2024). *The value of environmental surveillance for pandemic response*.
+This `quarto` document can be used to reproduce the results discussed in in *Nascimento de Lima et al.
+(2024).
+The value of environmental surveillance for pandemic response*.
+For details on how to install dependencies, see the `README.md` file.
+Before attempting to run this document, install all dependencies and run the `run_analysis.R` script.
+You should be able to reproduce all figures and tables from our paper using this document on a modern laptop within 30 minutes.
+
+# Running the model
+
+The following code block uses results stored in the `./output/r.rds` file or runs the script that generates this file.
+
+```{r}
+# The r object contains the results objects used in this file
+if(!file.exists("./output/r.rds")) {
+  source("./R/scripts/run_analysis.R")
+} else {
+  source("./R/library.R")
+  r <- readRDS("./output/r.rds")
+}
+
+```
 
@@
 
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+## Value of EWS under baseline assumptions
 
-## Running Code
+EWS systems provide meaningful value under base-case assumptions representing a hypothetical new pandemic with parameters mirroring those of SARS-Cov-2 (Supplementary Table 1).
 
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
+Assuming no NPIs interventions are put in place (No NPIs scenario), a new COVID-19-like pandemic would produce significant damage, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` deaths per 100,000 people (mean (95% prediction interval) over a one-year time-frame (Table 1).
+The unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to ilness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
+The total health cost would be projected at `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH")` dollars per person.
+
+If NPIs are in place and policy makers follow our base-case NPI policy, the death toll of the pandemic would be substantially reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` deaths per 100,000 population even if no enhanced early-warning system (EWS) is adopted.
+As it was the case during the COVID-19 pandemic, such an effort would involve a substantial collective effort.
+Under base-case assumptions, this effort would require `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L1plus_days")` days of public health interventions, of which `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L5_days")` days are spent under the most-restrictive policy intervention level.
+We project this effort to cost on the order of `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "CNPI")` per person.
+Based on those assumptions, the total cost of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "C")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "C")` per person, making the non-pharmaceutical intervention policy effective.
+
+Under this framework, EWS provide net-positive value to society if they enhance the effectiveness of NPIs more than they increase the burden of additional interventions introduced due to the early warning.
+We investigate this question investigating the value of EWS assuming that they provide 2, 5 or 10-day leading indicator relative to the disease incidence indicator used to inform COVID-19 response policies, and investigate how the key economic outcomes respond.
 
 ```{r}
-1 + 1
+gt::gt(r$table_1)
 ```
 
-You can add options to executable code like this
+* Notes: All costs are expressed as 2020 dollars per person. All outcomes were computed over a one-year time-frame.
 
-```{r}
-#| echo: false
-2 * 2
-```
+Points from this section for discussion:
 
-The `echo: false` option disables the printing of code (only output is displayed).
+* **Base-case results align with observed data in the pandemic** (by constrution with reality observed during COVID-19).
+US obseved on the order of 160 deaths per 100,000 people during the first year of the pandemic, and many states.
+* **EWS provide meaningful value under base-case assumptions.** Net-monetary benefit from surveillance sytem is on the order of 1,000 dollars per person over a one-year time-frame.
+This outweighs any available estimate of costs.
+
+## Conditions for high-value pandemic response EW systems
+{Describe conditions under which value is lower or higher}
+
+Points from this section for discussion:
+
+* Value of EWS hinges on:
+  + timeliness of decision-making
+  + adherence to NPIs
+  + relative costs/severity of disease
+
+# If we have time to look at those other cases:
+
+## EWS and policy coordination
+D. EWS alleviate lack of policy coordination. 
+There is a concern that EWS data will need to be integrated to be useful. However, we find that policy coordination and EWS act as substitutes. Policy coordination h
+
+## If we have time: Marginal Value of EWS
+D. EWS deployment also face decreasing marginal returns
+The value of EWS decreases as coverage increases.

### 2024-02-21 R/functions/OdinMetapop.R

@@ -310,14 +310,14 @@ OdinMetapop <- R6::R6Class(
 
 
       # Summarize across all jurisdictions
-      self$summary_all <- self$summary_jurisdiction %>%
+      self$summary <- self$summary_jurisdiction %>%
         group_by(rep) %>%
         select(-jurisdiction.id) %>%
         # Mean assumes equal population sizes, otherwise we would need a weighted mean.
         summarise(across(everything(),.fns = ~mean(.x)), .groups = "drop")
 
       # summarize across replications:
-      self$summary <- self$summary_all %>%
+      self$summary_all <- self$summary %>%
         select(-rep) %>%
         group_by() %>%
         summarise_all(summary_functions)

### 2024-02-21 R/functions/clear.labels.R

@@ -1,0 +1,13 @@
+
+# From: https://stackoverflow.com/questions/2394902/remove-variable-labels-attached-with-foreign-hmisc-spss-import-functions
+clear.labels <- function(x) {
+  if(is.list(x)) {
+    for(i in 1 : length(x)) class(x[[i]]) <- setdiff(class(x[[i]]), 'labelled')
+    for(i in 1 : length(x)) attr(x[[i]],"label") <- NULL
+  }
+  else {
+    class(x) <- setdiff(class(x), "labelled")
+    attr(x, "label") <- NULL
+  }
+  return(x)
+}

### 2024-02-21 R/functions/summary_funs.R

@@ -10,16 +10,10 @@ summary_functions$.mean = mean
 
 # Get estimate in Text Form:
 
-get_estimate = function(summary_tbl, strategy, col.sens, model_name, outcome, sig_numbers = 3) {
-
-  r = summary_tbl %>%
-    filter(Strategy == strategy, model == model_name, sensitivity == col.sens) %>%
-    mutate(across(where(is.numeric), .fns = ~signif(x = .x, digits = sig_numbers)))
-
-  vars = paste0(outcome, "_.", c("mean", "q.2.5", "q.97.5"))
-
-  paste0(r[,vars[1]], " (95% PI [", r[,vars[2]], ", ", r[,vars[3]],"])")
-
+pull_estimate <- function(df, ...) {
+  df %>%
+    dplyr::filter(...) %>%
+  .$estimate
 }
 
 get_outcome <- function(df, outcome_var, sig_digits = 3, ...) {
@@ -31,8 +25,10 @@ get_outcome <- function(df, outcome_var, sig_digits = 3, ...) {
   vars = paste0(outcome_var, "_.", c("mean", "q.2.5", "q.97.5"))
 
   paste0(r[,vars[1]], " (", r[,vars[2]], "-", r[,vars[3]],")")
-
 }
+
+
+
 
 # get_outcome(df, "outcome", a == 2, b == 2)
 

### 2024-02-26 reproduce-results.qmd
sensitivity analysis results with epi and societal inputs are sensible

--- 2024-02-19 reproduce-results.qmd
@@
 ---
-title: "results"
-format: html
+title: "The value of early warning systems for pandemic response"
+format:
+  html:
+    self-contained: true
+execute: 
+  echo: true
+  warning: false
+  error: false
 author: "Pedro Nascimento de Lima"
@@
 
-This quarto document reproduces the results section in Nascimento de Lima et al.
+This `quarto` document can be used to reproduce the results presented in *Nascimento de Lima et al.
 (2024).
-*The value of environmental surveillance for pandemic response*.
-For details on how to install dependencies, and reproduce these results, see the `README.md` file.
+The value of environmental surveillance for pandemic response*.
+For details on how to install dependencies, see the `README.md` file.
+Before attempting to run this document, install all dependencies and run the `run_analysis.R` script.
+You should be able to reproduce all figures and tables from our paper using this document on a modern laptop within 30 minutes.
 
-# Run Results
+# Running the model
 
-First, we instantiate the model and run it over the experimental designs used in the paper.
+The following code block uses results stored in the `./output/r.rds` file or runs the script that generates this file.
 
 ```{r}
+# The r object contains the results objects used in this file
+if(!file.exists("./output/r.rds")) {
+  source("./R/scripts/run_analysis.R")
+} else {
+  source("./R/library.R")
+  r <- readRDS("./output/r.rds")
+}
 
@@
 
-## Value of EWS under base-case results
-{Describe dynamics and results uder base-case assumptions}
+## Value of EWS under baseline assumptions
 
-D. EWS provide meaningful value under base-case assumptions
+<!--EWS systems provide meaningful value under base-case assumptions representing a hypothetical new pandemic with parameters mirroring those of SARS-Cov-2. -->
+Assuming no NPIs interventions are put in place (No NPIs scenario), a new COVID-19-like pandemic would produce significant damage, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` deaths per 100,000 people [mean (95% prediction interval of stochastic replications)] over a one-year time-frame (Table 1).
+The unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to ilness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
+The total health cost would be projected at `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH")` dollars per person.
 
-## Conditions for high-value early warning 
+If policymakers follow the base-case NPI policy simulated in this paper, the death toll of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` deaths per 100,000 population if no enhanced early-warning system (EWS) is adopted (NPIs w/o EWS scenario).
+As it was the case during the COVID-19 pandemic, such an achievement would involve a substantial public health effort.
+Under baseline assumptions, this effort would require `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L1plus_days")` days of public health interventions, of which `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L5_days")` days are spent under the most-restrictive policy intervention level.
+We project this effort to cost on the order of `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "CNPI")` per person.
+Based on those assumptions, the total cost of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "C")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "C")` per person.
+
+```{r}
+#| label: tbl-1
+#| tbl-cap: Base-case results
+gt::gt(r$table_1)
+```
+
+Notes: All costs are expressed as 2020 dollars per person. All outcomes were computed over a one-year time-frame.
+
+Under base-case assumptions, EWS provide net-positive value to society because they enhance the effectiveness of NPIs (i.e., preventing illness and death) more than they increase the burden of additional interventions introduced due to the early warning.
+Even an early warning system that provides only a 2-day early warning relative to a baseline pandemic response system (NPI + 2-day EWS scenario) is projected to reduce the pandemic death toll from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 2-day EWS", variable == "deaths_per_100k")` deaths per 100,000 population.
+Health benefits increase if the system can provide earlier  warning.
+A system that provides a 5-day early warning (NPIs + 5-day EWS scenario) would be projected to result in `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "deaths_per_100k")` deaths per 100,000 population, and a 10-day early warning would result in `r pull_estimate(r$table_1_long, Scenario == "NPIs + 10-day EWS", variable == "deaths_per_100k")` deaths per 100,000 people.
+
+The use of early warning would however result in earlier public health intervention, increasing the number of days of under NPIs from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L1plus_days")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "L1plus_days")` days for a system providing a 5-day early warning signal.
+That said, the number of days at the *maximum* intervention level (i.e., a lockdown) was projected to *decrease* with the use of the early warning system, as earlier action prevented surges from reaching alarming levels that prompted blunt interventions.
+For example, a system that provided a 5-day early warning reduced the number of days under the maximum intervention level from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L5_days")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "L5_days")` days.
+Still, the net effect of early warning system was to marginally increase NPI intervention costs, from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "CNPI")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "CNPI")` dollars per person, assuming a 5-day early warning.
+Health and economic outcomes weighted, the early warning system reduced the total pandemic costs from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "C")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "C")` dollars per person.
+The difference between the two costs can be interpreted as the net-monetary benefit of the system, which is estimated at `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "NMB")` dollars per person in the first year of the pandemic.
+
+<!--
+Points from this section for discussion:
+* **EWS provide meaningful value under base-case assumptions.** Net-monetary benefit from surveillance system is on the order of 1,000 dollars per person over a one-year time-frame.
+This outweighs any available estimate of costs.
+* **Base-case results align with observed COVID-19 data.** US obseved on the order of 160 deaths per 100,000 people during the first year of the pandemic, and many states.
+-->
+
+## Value of EWS system depends on epidemiological and societal parameters
+
+### Effect of epidemiological parameters
+
+A pathogen that is 50% more transmissible than wild-type SARS-Cov-2 would pose a greater challenge for pandemic response, causing `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "deaths_per_100k", NMB_comparator == T)` deaths per 100,000 people even if mitigated with the same NPI policy considered in our base-case scenario (Supplementary Table 2).
+Such a pandemic would cause a substantial societal cost of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "C", NMB_comparator == T)` dollars per person.
+In those conditions, the EWS would be particularly valuable, reducing the death toll of the pandemic to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "deaths_per_100k", NMB_comparator == F)` per 100,000 people, and would be projected to provide a NMB of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
+Conversely, a pathogen that is 50% less transmissible than SARS-Cov-2 would result in a lower death toll of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "deaths_per_100k", NMB_comparator == T)` without the EWS system, reducing the net-monetary benefit of the system to `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
+
+The value of the EWS system also depends on the infection fatality rate of the pathogen, but the net-benefit of the EWS would be positive even for a pathogen that is half as deadly as SARS-Cov-2.
+
+```{r}
+#| label: fig-1
+#| fig-cap: Base-case results
+#| column: screen
+#| fig-format: svg
+#| out-width: 100%
+plot(r$fig_A_epi)
+```
+
+## Effect of economic parameters
+
+## Conditions for high-value pandemic response EW systems
+
 {Describe conditions under which value is lower or higher}
-D. Value of EWS hinges on timeliness of and adherence to NPIs. 
-There was a
+
+Points from this section for discussion:
+
+* Value of EWS hinges on:
+  + timeliness of decision-making
+  + adherence to NPIs
+  + relative costs/severity of disease
+
+# If we have time to look at those other cases:
 
 ## EWS and policy coordination
-D. EWS alleviate lack of policy coordination. 
-There is a concern that EWS data will need to be integrated to be useful. However, we find that policy coordination and EWS act as substitutes. Policy coordination h
+EWS alleviate lack of policy coordination. 
+There is a concern that EWS data will need to be integrated to be useful. However, we find that policy coordination and EWS act as substitutes.
 
-## Marginal Value of EWS
-D. EWS deployment also face decreasing marginal returns
-The value of EWS decreases as coverage increases.
-
-# Results
-
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
-
-## Running Code
-
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
-
-```{r}
-1 + 1
-```
-
-You can add options to executable code like this
-
-```{r}
-#| echo: false
-2 * 2
-```
-
-The `echo: false` option disables the printing of code (only output is displayed).
+## If we have time: Marginal Value of EWS
+EWS deployment also face decreasing marginal returns. The value of EWS decreases as coverage increases.

### 2024-02-26 R/functions/OdinMetapop.R

@@ -333,7 +333,7 @@ OdinMetapop <- R6::R6Class(
     },
 
     # Simulate function
-    simulate = function(step = 0:365, y = NULL, use_names = TRUE, reps = 1000){
+    simulate = function(step = 0:365, y = NULL, use_names = TRUE, reps = 500){
 
       self$pre_process_inputs()
 

### 2024-02-26 R/functions/OdinSim.R

@@ -102,7 +102,7 @@ OdinSim <- R6::R6Class(
     #' @param type optional character string defining the type of input. Useful when one wants to only write inputs of a certain type to json.
     set_input = function(name, value, type = NA_character_) {
 
-      set.seed(1234)
+      # set.seed(1234)
 
       # If the model has been instantiated, and the input type is odin, set odin inptut
       if(!is.null(self$o)) {

### 2024-02-26 R/functions/summary_funs.R

@@ -2,8 +2,14 @@
 
 # Create summary functions list:
 # One can add more percentiles to the calculation by changing the vector below.
-p <- c(0.025, 0.975)
-p_names <- map_chr(p, ~paste0(".q.",.x*100))
+alpha <- 0.99
+p <- c((1-alpha)/2, 1-(1-alpha)/2)
+p_names <- c(".lower", ".upper")
+
+summary_functions <- map(p, ~partial(quantile, probs = .x, na.rm = TRUE)) %>%
+  set_names(nm = p_names)
+
+#p_names <- map_chr(p, ~paste0(".q.",.x*100))
 summary_functions = map(p, ~partial(quantile, probs = .x, na.rm = TRUE)) %>%
   set_names(nm = p_names)
 summary_functions$.mean = mean
@@ -22,7 +28,7 @@ get_outcome <- function(df, outcome_var, sig_digits = 3, ...) {
     as.data.frame() %>%
     mutate(across(where(is.numeric), .fns = ~signif(x = .x, digits = sig_digits)))
 
-  vars = paste0(outcome_var, "_.", c("mean", "q.2.5", "q.97.5"))
+  vars = paste0(outcome_var, "_.", c("mean", "lower", "upper"))
 
   paste0(r[,vars[1]], " (", r[,vars[2]], "-", r[,vars[3]],")")
 }

### 2024-02-26 R/functions/var.labels.R

@@ -25,16 +25,19 @@ var.labels <- c(
   IFR_time_mult = "time-varying ifr risk ratio",
   new_removed = "individuals removed from the simulation",
   deaths_per_100k = "Deaths per 100,000 people",
+  deaths_per_100k_diff = "Deaths averted per 100,000 people",
   total_cost_of_illness = "total cost of ilness per person, exc. deaths",
   CNPI = "NPI costs",
   CH_illness = "Cost of illness",
   CH_deaths = "Cost of deaths",
-  CH = "Total health costs",
+  CH = "Health costs",
+  CH_diff = "value of life saved",
   CSURV = "Surveillance cost",
-  C = "Total pandemic cost",
+  C = "Total costs",
   epi_size = "Epidemic size",
   L5_days = "Days of max NPI",
-  L1plus_days = "Days of any NPI"
+  L1plus_days = "Days of any NPI",
+  NMB = "Net monetary benefit"
 )
 
 var.labels.df <- data.frame(variable = names(var.labels), labels = var.labels)

### 2024-02-26 R/scripts/cluster_eval.R

@@ -19,7 +19,7 @@
 # Source necessary functions:
 source("./R/library.R")
 
-set.seed(1234)
+#set.seed(1234)
 
 # Instantiate model -------------------------------------------------------
 model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)

### 2024-02-26 R/scripts/run_analysis.R

@@ -1,0 +1,203 @@
+
+# Source all dependencies and model scripts
+source("./R/library.R")
+
+# results list
+r <- list()
+
+
+# create model and experiment ---------------------------------------------
+
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
+
+model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
+
+base_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx", sheet = "one-way-scenarios")
+
+# counterfactual scenarios without enhanced EWS
+count_scenarios <- base_scenarios %>%
+  filter(Section == "Scenarios") %>%
+  mutate(total_surv_lag = 13, p = 0.3, NMB_comparator = T)
+
+all_scenarios <- rbind(base_scenarios, count_scenarios)
+
+# create experiment
+base_experiment <- R6Experiment$new(model)
+
+base_experiment$set_design(grid_design_df = all_scenarios)
+
+
+
+# run experiment ----------------------------------------------------------
+
+base_results <- base_experiment$run(parallel = T,
+                                    cluster_eval_script = "./R/scripts/cluster_eval.R",
+                                    parallel::detectCores() - 2,
+                                    model_from_cluster_eval = T)
+
+
+
+# compute outcomes --------------------------------------------------------
+
+outcomes <- c("deaths_per_100k", "CH_illness", "CH_deaths", "CH", "L5_days", "L1plus_days", "CNPI", "C", "epi_size")
+
+# Compute differences at the replication level:
+comp_results <- base_results %>%
+  filter(NMB_comparator) %>%
+  select(rep, counterfactual.id, any_of(outcomes)) %>%
+  clear.labels() %>%
+  rename_at(vars(any_of(outcomes)), function(x) paste0(x, "_no_ews"))
+
+
+base_res_diff <- base_results %>%
+  clear.labels() %>%
+  left_join(comp_results, by = join_by(rep, counterfactual.id))
+
+base_res_diff <- base_res_diff %>%
[…]

### 2024-02-26 R/scripts/run_results.R

@@ -1,119 +1,0 @@
-
-# Source all dependencies and model scripts
-source("./R/library.R")
-
-r <- list()
-
-# Base case runs
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
-
-base_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx", sheet = "base-case")
-
-base_experiment <- R6Experiment$new(model)
-
-base_experiment$set_design(grid_design_df = base_scenarios)
-
-base_results <- base_experiment$run(parallel = T,
-                                    cluster_eval_script = "./R/scripts/cluster_eval.R",
-                                    parallel::detectCores() - 3,
-                                    model_from_cluster_eval = T)
-
-
-# Create posterior summaries for the parameters:
-r$base_results_long <- base_results %>%
-  select(-c(grid.id, lhs.id, params_design.id, param.id, model.id, all.params.id, c, p, total_surv_lag, a_up, a_down, policy.exp.id)) %>%
-  group_by(Scenario) %>%
-  pivot_longer(cols = -Scenario, names_to = "statistic", values_to = "value") %>%
-  as.data.frame() %>%
-  separate(col = statistic,into = c("variable", "stat"), sep = "_\\.")
-
-# table_param_summaries
-r$base_results_summary_long <- r$base_results_long %>%
-  #mutate(value = signif(value, digits = 3)) %>%
-  mutate(value = gt::vec_fmt_number(value, n_sigfig = 3, use_seps = T)) %>%
-  #mutate(value = formatC(value, format="f", big.mark=",")) %>%
-  #mutate(value = scales::label_comma()(value)) %>%
-  pivot_wider(id_cols = c(Scenario,variable), names_from = stat, values_from = value) %>%
-  mutate(estimate = paste0(mean, " (", q.2.5, "-", q.97.5,")")) %>%
-  select(Scenario, variable, estimate)
-
-r$table_1 <- r$base_results_summary_long %>%
-  #pivot_wider(id_cols = Scenario, names_from = variable, values_from = estimate)
[…]

### 2024-02-28 reproduce-results.qmd
effectiveness sensitivity analyses

@@
 
-### Effect of epidemiological parameters
+### Epidemiological parameters
 
@@
 Conversely, a pathogen that is 50% less transmissible than SARS-Cov-2 would result in a lower death toll of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "deaths_per_100k", NMB_comparator == T)` without the EWS system, reducing the net-monetary benefit of the system to `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
-
 The value of the EWS system also depends on the infection fatality rate of the pathogen, but the net-benefit of the EWS would be positive even for a pathogen that is half as deadly as SARS-Cov-2.
+All else equal, the EWS would still yield an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in a pandemic caused by a pathogen that is 50% less deadly than SARS-Cov-2, and would yield `r pull_estimate(r$table_1_long_all, Scenario == "1.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in responding to a pathogen that is 50% more deadly than COVID-19.
 
@@
 
-## Effect of economic parameters
+Other socio-economic and contextual variables also have a meaningful effect on the value of EWS for pandemic response, including NPI costs, NPI effectiveness, decision lags and the stringency with which decision-makers introduce nonpharmaceutical interventions (Supplementary Figure 1).
+
+### Policy lags and stringency
+
+Longer decision and implementation lags have an overall negative effect on the total costs of the pandemic, longer decision lags also would *increase* the net benefit of EWS systems for pandemic response.
+Our base-case scenario assume that decision-makers take one week to decide and implement non-pharmaceutical interventions.
+If instead policymakers took two weeks to introduce and implement interventions, the total pandemic cost increases from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "C", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == T)` dollars per person, assuming conventional surveillance.
+The early warning system effectively helps to roll-back decision lags, reducing total costs to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == F)` dollars per person, yielding a NMB of `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "NMB", NMB_comparator == F)` relative to conventional surveillance.
+Conversely, in contexts where NPI implementation is swift (i.e, decisions are made within 3 days), the NMB of EWS would be lower but still positive at `r pull_estimate(r$table_1_long_all, Scenario == "3-day decision lag", variable == "NMB", NMB_comparator == F)` dollars per person.
+
+Policy stringency (i.e., how stringent the case thresholds used to introduce or remove interventions) has a similar effect on the value of EWS relative to decision lags - the EWS provides the most value under the conditions that are least desirable for prevention of disease.
+In a context where decision makers are 50% more stringent than in our base-case assumptions (i.e., their case tresholds are half of its base-case), society is able to avert a substantial number of deaths, reducing deaths from `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "deaths_per_100k", NMB_comparator == T)`
+
+### Policy stringency
+
+the total cost of the pandemic is  baseline  o have a detrimental effect If decision-makers take two weeks to introduce and implement NPIs
+
+### Decision lags and policy stringency
+
+In our base-line assumptions, we assumed decision-makers take one week to update their risk classification levels based on the epidemiological data.
+That decision lag includes both the time it takes for public health decision-makers to
 

### 2024-02-28 R/functions/OdinMetapop.R

@@ -289,9 +289,9 @@ OdinMetapop <- R6::R6Class(
         ungroup() %>%
         # Compute expected deaths - since IFR is low, we compute expecte deaths
         # Expected deaths:
-        mutate(deaths = IFR * new_removed) %>%
+        #mutate(deaths = IFR * new_removed) %>%
         # Stochastic deaths:
-        #mutate(deaths = rbinom(n = nrow(.), size = new_removed,prob = IFR)) %>%
+        mutate(deaths = rbinom(n = nrow(.), size = new_removed,prob = IFR)) %>%
         mutate(deaths_per_100k = (deaths/population) * 10^5) %>%
         mutate(CH_illness = (new_removed * self$inputs$average_health_cost_per_infection)/population) %>%
         mutate(CH_deaths = deaths_per_100k * self$inputs$VSL / 10^5) %>%
@@ -333,7 +333,7 @@ OdinMetapop <- R6::R6Class(
     },
 
     # Simulate function
-    simulate = function(step = 0:365, y = NULL, use_names = TRUE, reps = 500){
+    simulate = function(step = 0:365, y = NULL, use_names = TRUE, reps = 1000){
 
       self$pre_process_inputs()
 

### 2024-02-28 R/functions/summary_funs.R

@@ -2,7 +2,7 @@
 
 # Create summary functions list:
 # One can add more percentiles to the calculation by changing the vector below.
-alpha <- 0.99
+alpha <- 0.95
 p <- c((1-alpha)/2, 1-(1-alpha)/2)
 p_names <- c(".lower", ".upper")
 
@@ -37,5 +37,3 @@ get_outcome <- function(df, outcome_var, sig_digits = 3, ...) {
 
 
 # get_outcome(df, "outcome", a == 2, b == 2)
-
-

### 2024-02-28 R/scripts/03_run_larger_exp_design.R

@@ -42,10 +42,10 @@ policy_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx",sheet = "policy_pr
 
 
 # Other variables
-total_surv_lag <- seq.default(from = 2, to = 20, by = 2)
-R0 <- c(1.5,2,3)
+total_surv_lag <- c(13,11,8,3)
+R0 <- c(1.5,2.5,3)
 cost_max_npi <- c(0.1,0.25,0.87, 1.25)
-L_c <- c(0,1)
+#L_c <- c(0,1)
 r <- c(0.01, 0.001, 0.02)
 
 # Surveillance design
@@ -57,7 +57,7 @@ surv_design <- data.frame(total_surv_lag = total_surv_lag) %>%
 other_vars_design <- expand_grid(
   R0
   ,cost_max_npi
-  ,L_c
+#  ,L_c
   ,r
 ) %>%
   mutate(other.vars.id = row_number())

### 2024-02-28 R/scripts/run_analysis.R

@@ -109,21 +109,48 @@ r$table_1 <- r$table_1_long %>%
   relocate(Outcome, `No NPIs`, `NPIs w/o EWS`, `NPIs + 2-day EWS`, `NPIs + 5-day EWS`, `NPIs + 10-day EWS`)
 
 
-
 # dot plot with differences -----------------------------------------------
 
 colors = c("#8856a7", "#9ebcda")
 
+
+# Deaths figure:
+
 r$fig_A_data <- r$table_1_long_numeric %>%
   mutate(EWS = ifelse(!NMB_comparator, "NPIs w/ EWS", "NPIs w/o EWS")) %>%
-  filter(variable %in% c("CH", "CNPI", "C", "NMB")) %>%
+  filter(variable %in% c("CH", "CNPI", "C", "NMB", "deaths_per_100k")) %>%
   left_join(var.labels.df, by = join_by(variable)) %>%
   filter(Section == "Scenarios") %>%
-  mutate(variable = factor(variable,levels = c("CH", "CNPI", "C", "NMB"), ordered = T)) %>%
-  mutate(labels = factor(labels,levels = c("Health costs", "NPI costs", "Total costs", "Net monetary benefit"), ordered = T)) %>%
+  mutate(variable = factor(variable,levels = c("CH", "CNPI", "C", "NMB", "deaths_per_100k"), ordered = T)) %>%
+  mutate(labels = factor(labels,levels = c("Health costs", "NPI costs", "Total costs", "Net monetary benefit", "Deaths per 100,000 people"), ordered = T)) %>%
   mutate(Scenario = factor(Scenario, levels = rev(unique(base_scenarios$Scenario)), ordered = T)) %>%
   arrange(variable, Scenario) %>%
   filter(!(variable == "NMB" & NMB_comparator))
+
+# Simplified figure for presentation
+
+r$deaths_figure <- r$fig_A_data %>%
+  filter(variable %in% c("deaths_per_100k", "NMB")) %>%
+  filter(Scenario %in% c("Base-case", "0.5x transmissible", "1.5x transmissible")) %>%
+  ggplot() +
+  geom_segment(aes(x = Scenario, xend = Scenario, y = lower, yend = upper, color = EWS), linewidth = 1.5) +
+  geom_point(aes(x = Scenario, y = mean, color = EWS), size = 3) +
+  #geom_hline(yintercept = 1500, color = "grey") +
+  coord_flip() +
[…]

### 2024-02-28 reproduce-results.qmd
edits from grammarly

@@
 
-This `quarto` document can be used to reproduce the results presented in *Nascimento de Lima et al.
+This `quarto` document produces the results presented in *Nascimento de Lima et al.
 (2024).
@@
 
+# Scenarios
+Supplementary Table 1 provides details on parameter values used under our baseline scenario.
+The baseline scenario with conventional surveillance assumes a syndromic surveillance system based upon confirmed case counts, and there is a thirteen-day time span from exposure to the pathogen to detection by the public health authority.
+This thirteen-day lag covers the disease incubation phase (6.6 days)(cite) and testing and reporting delays (6 days).(cite) Consistent with estimates from the COVID-19 pandemic, the conventional surveillance system has a 30% case ascertainment rate.(cite) We assume the environmental sampling system can provide absolute, unbiased disease incidence estimates that lead the case counts by five days in the base-case scenario.
+We explore longer and shorter lead times, as this lag depends on  two to ten days.
+We use 5 days as the base-case scenario, as wastewater surveillance systems have been able to demonstrate.
+(cite) Moreover, this analysis assumed that all the populations within the jurisdictions are covered by the EWS (or equivalently, that the sampled population is a representative sample of all the population).
+
+# Population
+The model simulates disease percolating through six interconnected jurisdictions that independently make nonpharmaceutical decisions.
+Individuals can travel and work across jurisdiction lines, and the six jurisdictions each have a population of 10^6 individuals.
+Mortality parameters were chosen to reflect mortality rates observed in the United States during the first year of the COVID-19 pandemic, hence our results can be interpreted as expected outcomes for a cluster of US counties.
+
+# Outcomes
+The average mortality per 100,000 population across the jurisdiction is the primary health outcome in this study.
+We also compute the cost of illness due to disease, using SARS-Cov-2 as the base-case scenario.
+We also account for economic costs caused by the nonpharmaceutical interventions introduced (discussed in the Supplementary Appendix).
+All outcomes are computed over a one-year time frame from the pandemic onset and are averages computed across the jurisdictions in our model.
+Since the model is stochastic, we simulate the model 1,000 times and present means and 95% prediction intervals (computed with the 2.5% and the 97.5% percentiles) of each outcome.
+
+# Inter-jurisdiction mixing
+We use the following approach to align a
+
+# Calibration: Add as a footnote to tau.
+We use parameter estimates from the literature whenever possible (Supplementary Table 1).
+Parameters that are unobserved in the data, notably tau and.
+R0 is computed based on disease progression rates.
+We calibrate only the $tau$ parameter, which is highly heterogeneous and context-dependent (cite Laura), to match the number of deaths observed during the first year in the United States.
+We find that $tau = 0.137$.
+We explore the implication of a wider range of policy effectiveness in sensitivity analyses, and explore this particular parameter in more details.
+Surprisingly, the policy effectiveness
+
+## Sensitivity Analyses
+We explore the performance of the EWS under conditions that depart from COVID-19 by varying model parameters one at a time, aiming to represent a new emerging pathogen against which the population has no immunity.
+Since a probabilistic characterization of uncertainty to parameters is elusive, we refrain from making probabilistic statements
+
 # Results
@@
 Assuming no NPIs interventions are put in place (No NPIs scenario), a new COVID-19-like pandemic would produce significant damage, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` deaths per 100,000 people [mean (95% prediction interval of stochastic replications)] over a one-year time-frame (Table 1).
-The unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to ilness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
+The unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to illness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
 The total health cost would be projected at `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH")` dollars per person.
@@
 
-Notes: All costs are expressed as 2020 dollars per person. All outcomes were computed over a one-year time-frame.
+Notes: All costs are expressed as 2020 dollars per person. All outcomes were computed over a one-year time frame.
 
-Under base-case assumptions, EWS provide net-positive value to society because they enhance the effectiveness of NPIs (i.e., preventing illness and death) more than they increase the burden of additional interventions introduced due to the early warning.
-Even an early warning system that provides only a 2-day early warning relative to a baseline pandemic response system (NPI + 2-day EWS scenario) is projected to reduce the pandemic death toll from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 2-day EWS", variable == "deaths_per_100k")` deaths per 100,000 population.
+Under baseline assumptions, the EWS provides net-positive value to society because they enhance the effectiveness of NPIs (i.e., preventing illness and death) more than they increase the costs of additional interventions introduced due to the early warning.
+Even an early warning system that provides only a 2-day early warning relative to a conventional pandemic response system (NPI + 2-day EWS scenario) is projected to reduce the pandemic death toll from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 2-day EWS", variable == "deaths_per_100k")` deaths per 100,000 population.
 Health benefits increase if the system can provide earlier warning.
@@
 
-The use of early warning would however result in earlier public health intervention, increasing the number of days of under NPIs from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L1plus_days")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "L1plus_days")` days for a system providing a 5-day early warning signal.
-That said, the number of days at the *maximum* intervention level (i.e., a lockdown) was projected to *decrease* with the use of the early warning system, as earlier action prevented surges from reaching alarming levels that prompted blunt interventions.
+The use of early warning would result in earlier public health intervention, increasing the number of days of under NPIs from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L1plus_days")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "L1plus_days")` days for a system providing a 5-day early warning signal.
+That said, the number of days at the *maximum* intervention level (i.e., a lockdown) was projected to *decrease* with the early warning system, as earlier action prevented surges from reaching alarming levels that prompted blunt interventions.
 For example, a system that provided a 5-day early warning reduced the number of days under the maximum intervention level from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L5_days")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "L5_days")` days.
-Still, the net effect of early warning system was to marginally increase NPI intervention costs, from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "CNPI")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "CNPI")` dollars per person, assuming a 5-day early warning.
+Still, the net effect of the EWS was to marginally increase NPI intervention costs, from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "CNPI")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "CNPI")` dollars per person, assuming a 5-day early warning.
 Health and economic outcomes weighted, the early warning system reduced the total pandemic costs from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "C")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "C")` dollars per person.
-The difference between the two costs can be interpreted as the net-monetary benefit of the system, which is estimated at `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "NMB")` dollars per person in the first year of the pandemic.
+The difference between the two costs is the net-monetary benefit of the system, which is estimated at `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "NMB")` dollars per person in the first year of the pandemic.
 
@@
 
-Policy stringency (i.e., how stringent the case thresholds used to introduce or remove interventions) has a similar effect on the value of EWS relative to decision lags - the EWS provides the most value under the conditions that are least desirable for prevention of disease.
-In a context where decision makers are 50% more stringent than in our base-case assumptions (i.e., their case tresholds are half of its base-case), society is able to avert a substantial number of deaths, reducing deaths from `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "deaths_per_100k", NMB_comparator == T)`
+Policy stringency (i.e., how stringent the case thresholds used to introduce or remove interventions) has a similar effect on the value of EWS relative to decision lags - the EWS provides the most value under the least desirable conditions for disease prevention.
+In a context where decision-makers are 50% more stringent than in our base-case assumptions (i.e., their case thresholds are half of its base-case), society is able to avert a substantial number of deaths, reducing mortality from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "deaths_per_100k", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "deaths_per_100k", NMB_comparator == T)` deaths per 100,000 population without the EWS.
+In that scenario, the EWS yeilds `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "NMB", NMB_comparator == F)` dollars per person of net benefit since most of the deaths would already be prevented through more stringent NPIs.
+Conversely, lenient policy choises worsen overall outcomes but also imply a higher value of the enhanced surveillance system.
+
+### NPI effectiveness
+The effectiveness of NPIs to reduce transmission is a key unkonwn at the outset of a new pandemic.
+At our baseline parameterization, NPI effectiveness has a non-monotonic effect on the value of surveillance systems to pandemic response (Supplementary Figure X).
+If NPI effectiveness is 0%, an improved surveillance system has no value, as expected.
+The value of the enhanced surveillance system non-linearly increases with NPI effectiveness slowly for maximum NPI effectiveness below 50%, and peaked when the maximum NPI effectiveness reaches 60%, enough to reduce the baseline R0 of 2.5 below 1 at pandemic outset.
+Beyond that point, increases in NPI effectiveness produce lower total costs, but reduce the value of the surveillance system since such betterFor a baseline R0 of 2.5, a maximum NPI effectiveness of 50% implies that even the maximum intervention level (i.e., a lockdown) is not able to reduce Rt below 1 at the outset of the pandemic.
 
@@
 
-In our base-line assumptions, we assumed decision-makers take one week to update their risk classification levels based on the epidemiological data.
+In our baseline assumptions, we assumed decision-makers take one week to update their risk classification levels based on the epidemiological data.
 That decision lag includes both the time it takes for public health decision-makers to
@@
 
+<!--
+
 # If we have time to look at those other cases:
@@
 EWS deployment also face decreasing marginal returns. The value of EWS decreases as coverage increases.
+
+-->

### 2024-02-28 reproduce-results.qmd
creating calibration script

@@
 
-# Inter-jurisdiction mixing
-We use the following approach to align a
-
 # Calibration: Add as a footnote to tau.
@@
 
+## Early warning systems can shape disease and policy dynamics
+
+**MAybe this is useful for a briefing or slides, but maybe not the paper**:
+An early warning system that shifts the epidemiological time-series used to inform non-pharmaceutical decisions can change the course of disease when used with effective interventions (Figure 1).
+We consider a network of jurisdictions that are independently managed by their own public health departments.
+For simplicity, we consider two cases: One in which the jurisdictions base their decisions upon syndromic surveillance data, and another (counterfactual) scenario where they are able to use data from environmental sampling sources (i.e., without loss of generality, wastewater surveillance), which lead the conventional surveillance data by a few days.
+The two sets of jurisdictions are otherwise equal, and face the exact same pandemic.
+
 ## Value of EWS under baseline assumptions
@@
 <!--EWS systems provide meaningful value under base-case assumptions representing a hypothetical new pandemic with parameters mirroring those of SARS-Cov-2. -->
+
 Assuming no NPIs interventions are put in place (No NPIs scenario), a new COVID-19-like pandemic would produce significant damage, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` deaths per 100,000 people [mean (95% prediction interval of stochastic replications)] over a one-year time-frame (Table 1).
-The unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to illness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
+In the absence of any behavioral change, the unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to illness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
 The total health cost would be projected at `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH")` dollars per person.
 
-If policymakers follow the base-case NPI policy simulated in this paper, the death toll of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` deaths per 100,000 population if no enhanced early-warning system (EWS) is adopted (NPIs w/o EWS scenario).
+If policymakers followed the base-case NPI policy simulated in this paper, the death toll of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` deaths per 100,000 population if no enhanced early-warning system (EWS) is adopted (NPIs w/o EWS scenario).
 As it was the case during the COVID-19 pandemic, such an achievement would involve a substantial public health effort.
@@
 Based on those assumptions, the total cost of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "C")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "C")` per person.
-
-```{r}
-#| label: tbl-1
-#| tbl-cap: Base-case results
-gt::gt(r$table_1)
-```
-
-Notes: All costs are expressed as 2020 dollars per person. All outcomes were computed over a one-year time frame.
 
@@
 
-```{r}
-#| label: fig-1
-#| fig-cap: Base-case results
-#| column: screen
-#| fig-format: svg
-#| out-width: 100%
-plot(r$fig_A_epi)
-```
+### NPI Costs
 
-Other socio-economic and contextual variables also have a meaningful effect on the value of EWS for pandemic response, including NPI costs, NPI effectiveness, decision lags and the stringency with which decision-makers introduce nonpharmaceutical interventions (Supplementary Figure 1).
+NPI costs had a moderate effect on the value of surveillance systems.
+If NPI costs were 50% higher than assumed in our base-case scenario (i.e., if achieving the same NPI effectiveness were to cost 50% more than assumed under our base-case scenario), then the total cost of the pandemic would increase from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "C", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x NPI costs", variable == "C", NMB_comparator == T)` dollars per person with the conventional surveillance system.
+(Supplementary Figure 1, **Supplementary Table X**).
+Despite that increase, the NMB of the system would be only marginally reduced from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "NMB", NMB_comparator == F)` to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x NPI costs", variable == "NMB", NMB_comparator == F)` dollars per person, because the EWS system only marginally increased the number of days under interventions.
 
@@
 In a context where decision-makers are 50% more stringent than in our base-case assumptions (i.e., their case thresholds are half of its base-case), society is able to avert a substantial number of deaths, reducing mortality from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "deaths_per_100k", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "deaths_per_100k", NMB_comparator == T)` deaths per 100,000 population without the EWS.
-In that scenario, the EWS yeilds `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "NMB", NMB_comparator == F)` dollars per person of net benefit since most of the deaths would already be prevented through more stringent NPIs.
-Conversely, lenient policy choises worsen overall outcomes but also imply a higher value of the enhanced surveillance system.
+In that scenario, the EWS yields `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "NMB", NMB_comparator == F)` dollars per person of net benefit since most of the deaths would already be prevented through more stringent NPIs.
+Conversely, if policy makers were more lenient policy choices would change the syst overall outcomes but also imply a higher value of the enhanced surveillance system.
 
 ### NPI effectiveness
-The effectiveness of NPIs to reduce transmission is a key unkonwn at the outset of a new pandemic.
-At our baseline parameterization, NPI effectiveness has a non-monotonic effect on the value of surveillance systems to pandemic response (Supplementary Figure X).
+
+NPI effectiveness (i.e., the % reduction in infectious contacts caused by NPIs) had a non-monotonic effect on the value of surveillance systems to pandemic response (**Supplementary Figure X**).
 If NPI effectiveness is 0%, an improved surveillance system has no value, as expected.
-The value of the enhanced surveillance system non-linearly increases with NPI effectiveness slowly for maximum NPI effectiveness below 50%, and peaked when the maximum NPI effectiveness reaches 60%, enough to reduce the baseline R0 of 2.5 below 1 at pandemic outset.
-Beyond that point, increases in NPI effectiveness produce lower total costs, but reduce the value of the surveillance system since such betterFor a baseline R0 of 2.5, a maximum NPI effectiveness of 50% implies that even the maximum intervention level (i.e., a lockdown) is not able to reduce Rt below 1 at the outset of the pandemic.
-
-### Policy stringency
-
-the total cost of the pandemic is baseline o have a detrimental effect If decision-makers take two weeks to introduce and implement NPIs
-
-### Decision lags and policy stringency
-
-In our baseline assumptions, we assumed decision-makers take one week to update their risk classification levels based on the epidemiological data.
-That decision lag includes both the time it takes for public health decision-makers to
-
-## Conditions for high-value pandemic response EW systems
-
-{Describe conditions under which value is lower or higher}
-
-Points from this section for discussion:
-
-* Value of EWS hinges on:
-  + timeliness of decision-making
-  + adherence to NPIs
-  + relative costs/severity of disease
+The value of the enhanced surveillance system increases with NPI effectiveness slowly for maximum NPI effectiveness below around 50%, and peaked when the maximum NPI effectiveness reaches 60%, enough to reduce the baseline R0 of 2.5 below 1 at pandemic outset.
+In general, we find that the value of the EWS peaks at a maximum NPI effectiveness value at $1-1/R0$, for different values of R0.
+Beyond that point, increases in NPI effectiveness result in lower total pandemic costs, but the value of the surveillance system starts to decrease.
 

### 2024-02-28 R/functions/OdinMetapop.R

@@ -333,7 +333,7 @@ OdinMetapop <- R6::R6Class(
     },
 
     # Simulate function
-    simulate = function(step = 0:365, y = NULL, use_names = TRUE, reps = 1000){
+    simulate = function(step = 0:365, y = NULL, use_names = TRUE, reps = 200){
 
       self$pre_process_inputs()
 

### 2024-02-28 R/scripts/04_calibration.R

@@ -1,0 +1,52 @@
+
+# Source all dependencies and model scripts
+source("./R/library.R")
+
+# results list
+
+# Policy effectiveness and stringency
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
+
+
+calib_tau_c_obj_fn <- function(x = c(7.6, 0.125), target_deaths = 160, target_l5_days = 240, deaths_weight = 0.5) {
+
+  set.seed(1234)
+
+  model$set_input("c", x[1])$
+        set_input("tau", x[2])
+
+  model$simulate(reps = 100)
+
+  deaths_sim <- as.numeric(model$summary_all$deaths_per_100k_.mean[1])
+  l5days_sim <- as.numeric(model$summary_all$L5_days_.mean[1])
+
+
+  sq_distance <- deaths_weight * (target_deaths - deaths_sim)^2 + (1-deaths_weight) * (target_l5_days - l5days_sim)^2
+  deaths_ape <- round(100*(deaths_sim - target_deaths)/target_deaths, 2)
+  l5days_ape <- round(100*(l5days_sim - target_l5_days)/target_l5_days, 2)
+
+  print(paste0("Params: tau: ", signif(x[2], 3), ". c: ", signif(x[1], 3)))
+  print(paste0("Abs Perc Error: Deaths: ", deaths_ape, ". L5 days: ", l5days_ape))
+
+  return(sq_distance)
+
+  }
+
+# Test function:
+calib_tau_c_obj_fn(x = c(c = 5, tau = 0.115))
+
+# Calibrate:
+solution <- optim(par = c(c = 10, tau = 0.137),
+                  control = list(maxit = 50, abstol = 5^2),
+                  #method = "Brent", # "L-BFGS-B",
+                  fn = calib_tau_c_obj_fn,
+                  lower = c(5, 0.1),
+                  upper = c(15, 0.18),
+                  target_deaths = 160,
+                  target_l5_days = 240,
+                  deaths_weight = 0.5)
+
+
+
+solution$par
+

### 2024-02-28 R/scripts/run_analysis.R

@@ -3,7 +3,12 @@
 source("./R/library.R")
 
 # results list
-r <- list()
+
+if(!file.exists("./output/r.rds")) {
+  r <- list()
+} else {
+  r <- readRDS("./output/r.rds")
+}
 
 
 # create model and experiment ---------------------------------------------
@@ -77,7 +82,7 @@ r$base_results_long <- base_results_rep_summaries %>%
 # table_param_summaries
 r$table_1_long_all <- r$base_results_long %>%
   #mutate(value = signif(value, digits = 3)) %>%
-  mutate(value = gt::vec_fmt_number(value, n_sigfig = 4, use_seps = T)) %>%
+  mutate(value = gt::vec_fmt_number(value, n_sigfig = 3, use_seps = T)) %>%
   #mutate(value = formatC(value, format="f", big.mark=",")) %>%
   #mutate(value = scales::label_comma()(value)) %>%
   pivot_wider(id_cols = c(Scenario,Section, Class, NMB_comparator, counterfactual.id, variable), names_from = stat, values_from = value) %>%
@@ -153,6 +158,7 @@ ggsave(plot = r$deaths_figure, filename = "./output/deaths_plot.svg",units = "in
 
 
 r$fig_A <- r$fig_A_data %>%
+  filter(variable != "deaths_per_100k") %>%
   ggplot() +
   geom_segment(aes(x = Scenario, xend = Scenario, y = lower, yend = upper, color = EWS)) +
   geom_point(aes(x = Scenario, y = mean, color = EWS), size = 3) +

### 2024-03-01 archive/paper/archive/manuscript.qmd
cleaning up old files

--- 2023-09-14 paper/manuscript.qmd
@@
 
-We have thus added the following limitation to our model; when the number of patients in need of hospital care exceeds the available supply (e.g.
-if there is an insufficient supply of supplemental oxygen, ventilators, extracorporeal membrane oxygenation machines, and medical staff to meet demand), the case fatality rate among these patients increases.
+We have thus added a capacity limitation for care provisioning to our model: when the number of patients in need of hospital care exceeds the available supply (e.g.
+if there is an insufficient supply of supplemental oxygen, ventilators, extracorporeal membrane oxygenation machines, and medical staff to meet demand), the infection fatality rate, $r$, increases.
 
@@
 
-During the course of a pandemic, providers may learn more about the pathogen and illness, developing improved standards of care, even absent pharmaceudical intervention.
+During the course of a pandemic, providers may learn more about the pathogen and illness, developing improved standards of care, even absent pharmaceutical intervention.
 This has been the case for treatment of hypoxemia and acute respiratory distress syndrome (ARDS) during the COVID-19 pandemic.
-https://pubmed.ncbi.nlm.nih.gov/33156382/ For example, initually, patients suffering from hypoxemia were usually laid in a supine position with "aggressive airway management and intubation".
+https://pubmed.ncbi.nlm.nih.gov/33156382/ For example, initially, patients suffering from hypoxemia were usually laid in a supine position with "aggressive airway management and intubation".
 However, in the Spring of 2020, evidence mounted that initially simply placing patients in a prone, rather than supine position, as well as less aggressive intubation, led to improved outcomes.
 https://nap.nationalacademies.org/cart/download.cgi?record_id=26573 pg 7 https://pubmed.ncbi.nlm.nih.gov/32569585/.
-Implementation of the new protocols, likely resulted in lower infection fatality rates and health system burden.
+Some extant pharmaceutical treatments, such as the provisioning of dexamethasone to patients with severe illness, proved to be useful while others were shown not to improve infection fatality rates ($r$).
+https://www.nature.com/articles/d41586-020-03132-4
+
+Implementation of these new protocols likely contributed to the decreasing fatality rates and health system burden in the months preceeding the introduction of the COVID-19 vaccine, as shown by [Covid forecasting team 2022 DOI:  10.1016/S0140-6736(21)02867-1], with the infection fatality rate dropping by approximately 33% over the first eight months, based on a meta-analysis of seroprevalence studies and reported fatalities.
+
+https://pubmed.ncbi.nlm.nih.gov/35219376/"During the pre-vaccine period, the estimated 33% decrease in median IFR over 8 months suggests that treatment for COVID-19 has improved over time.
+Estimating IFR for the pre-vaccine era provides an important baseline for describing the progression of COVID-19 mortality patterns.
+"
+
+Lancet https://pubmed.ncbi.nlm.nih.gov/35219376/
+
+.  
 https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7253105/
 https://pubmed.ncbi.nlm.nih.gov/37016834/
+Similarly, the improvement in
 

### 2024-03-01 archive/paper/archive/template.qmd
cleaning up old files

---
title: Short Paper
subtitle: A Short Subtitle
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliations: 
        - id: some-tech
          name: Some Institute of Technology
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Bob Security
    email: bob@example.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    note: |
      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  - name: Cat Memes
    email: cat@example.com
    affiliations:
        - ref: another-u
    note: Yet another author footnote.
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  elsevier-html:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
---
<!-- Source: https://github.com/quarto-journals/elsevier download repository -->
Please make sure that your manuscript follows the guidelines in the 
Guide for Authors of the relevant journal. It is not necessary to 
typeset your manuscript in exactly the same way as an article, 
unless you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see   <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

Here are two sample references:  @Feynman1963118 @Dirac1953888.

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML. 
You can sets extra options with `natbiboptions` variable in YAML header. Example

```
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at
<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. 
To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL

If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded
and stored locally, or the url can be used as in the example header.

# Equations

Here is an equation:
$$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {-}

### 2024-03-01 R/scripts/01_single_run_test.R

@@ -1,117 +1,0 @@
-
-#------------------------------------------------------------------------------#
-# Code repository for Analysis of Genomic Sequencing information
-#
-# Author: Pedro Nascimento de Lima
-# See README.md for information on usage and licensing
-#------------------------------------------------------------------------------#
-
-# This script demonstrates how to run the model for a single run:
-
-source("./R/library.R")
-
-# Single run examples -----------------------------------------------------
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file, "model1")
-#model2 <- OdinMetapop$new("stochastic_metapopulation2.R", "./data/metapopulation-inputs-heterogeneous.xlsx")
-
-model$simulate()
-
-
-model$set_input("c", 10)$
-  set_input("tau", 0.05)$
-  simulate()
-
-View(model$summary)
-
-
-get_outcome(model$summary, outcome_var = "CH")
-
-
-# all default model inputs are in the inputs object
-model$inputs$beta
-# IT looks like the mixing is not being partitioned correctly!
-#. it should be 0.1, 0.1
-
-# We can run the model with default inputs by using the simulate function
-# i.e., run for 100 days, for 100 replications
-View(model$simulate(step = 0:300, reps = 100))
-
-
-
-
-model2$simulate(step = 0:300, reps = 100)
-
-# Compare both models:
-res1 <- model$res_long %>%
-  group_by(rep, step) %>%
-  summarise(I = sum(E + P + I)) %>%
-  mutate(model = "homogeneous")
-
-res2 <- model2$res_long %>%
-  group_by(rep, step) %>%
-  summarise(I = sum(E + P + I)) %>%
-  mutate(model = "heterogeneous")
-
-all_res <-rbind(res1, res2)
-
-rep_model <- all_res %>%
-  select(rep, model) %>%
-  distinct() %>%
-  mutate(rep_model_id = row_number())
-
-all_res %>%
-  #left_join(rep_model) %>%
-  ggplot() +
-  geom_line(mapping = aes(x = step, y = I, group = interaction(model, rep), color = model)) #+
-  #facet_wrap(~model)
-
-
-
-
-
-
-
-model$inputs$R
-model$inputs$beta
-
-View(model$res_long %>% mutate(E_perc = E/(E+P+I), P_perc = P/(E+P+I), I_perc = I/(E+P+I)))
[…]

### 2024-03-01 R/scripts/02_sample_analysis.R

@@ -1,59 +1,0 @@
-
-
-
-#------------------------------------------------------------------------------#
-# Code repository for Analysis of Genomic Sequencing information
-#
-# Author: Pedro Nascimento de Lima
-# See README.md for information on usage and licensing
-#------------------------------------------------------------------------------#
-
-# This script demonstrates how to run an experimental design using the model
-
-# Set up an experimental design and run the model:
-source("./R/library.R")
-
-# Instantiate model -------------------------------------------------------
-
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-# Run full-factorial experiment -------------------------------------------
-
-# Create sample parameter set
-# This is here to demonstrate that data.frame with parameter sets (i.e., a
-# sample from the posterior distribution of model parameters)
-# can be assigned to the model. This is required for R6Sim models that
-# are used within the R6Experiment class, but not necessary otherwise.
-model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
-
-# Instantiate an experiment
-
-# Note that if the experiment is to be run in parallel, there is an underlying
-# assumption that the model will be instantiated in the cluster eval
-# script
-experiment <- R6Experiment$new(model)
-
-# Set experimental parameters
-experiment$
-  set_parameter("total_surv_lag", "grid", seq.default(from = 0, to = 21, by = 3))$ # Surveillance lag
-  set_parameter("p", "grid", c(0.5,1))$ # case ascertainment proportion
-  set_parameter("tau", "grid", c(0.12,0.15,0.18))$ # intevention effectiveness per intervention level
-  set_parameter("c", "grid", c(0.01,1,5,10,100))$ # case intervention threshold per 100,000 people
-  set_parameter("R0", "grid", c(1.5,2,2.5))$ # Initial reproduction number
-  set_parameter("cost_max_npi", "grid", c(0.1,0.25,0.87))$ # Initial reproduction number
-  set_parameter("L_c", "grid", c(0,1)) # Turn intervention coordination on and off
-
-
-# Set designs creates the experimental design data.frame
-experiment$set_design()
-
-# Run experiment (parallel option not available with odin at this time):
[…]

### 2024-03-01 R/scripts/03_run_larger_exp_design.R

@@ -1,110 +1,0 @@
-
-
-
-#------------------------------------------------------------------------------#
-# Code repository for Analysis of Genomic Sequencing information
-#
-# Author: Pedro Nascimento de Lima
-# See README.md for information on usage and licensing
-#------------------------------------------------------------------------------#
-
-# This script demonstrates how to run an experiment using the model
-
-# Set up an experimental design and run the model:
-source("./R/library.R")
-
-# Instantiate model -------------------------------------------------------
-
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-# Run full-factorial experiment -------------------------------------------
-
-# Create sample parameter set
-# This is here to demonstrate that data.frame with parameter sets (i.e., a
-# sample from the posterior distribution of model parameters)
-# can be assigned to the model. This is required for R6Sim models that
-# are used within the R6Experiment class, but not necessary otherwise.
-model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
-
-# Instantiate an experiment
-
-# Note that if the experiment is to be run in parallel, there is an underlying
-# assumption that the model will be instantiated in the cluster eval
-# script
-experiment <- R6Experiment$new(model)
-
-
-# Create an experimental design combining decision-maker characteristics and epi/social characteristics
-
-# Policy scenarios represent potential decision makers
-policy_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx",sheet = "policy_profile") %>%
-  mutate(policy.id = row_number())
-
-
-# Other variables
-total_surv_lag <- c(13,11,8,3)
-R0 <- c(1.5,2.5,3)
-cost_max_npi <- c(0.1,0.25,0.87, 1.25)
-#L_c <- c(0,1)
-r <- c(0.01, 0.001, 0.02)
-
-# Surveillance design
-surv_design <- data.frame(total_surv_lag = total_surv_lag) %>%
-  mutate(p = ifelse(total_surv_lag < 10, 1, 0.5)) %>%
-  mutate(surv.id = row_number())
-
-# Other variables design:
-other_vars_design <- expand_grid(
-  R0
-  ,cost_max_npi
-#  ,L_c
-  ,r
-) %>%
[…]

### 2024-03-01 R/scripts/04_calibration.R

@@ -1,99 +1,0 @@
-
-# Source all dependencies and model scripts
-source("./R/library.R")
-
-# This script is used to calibrate parameters tau and c,
-# tau: marginal effectiveness of NPIs (in reducing transmission)
-# c: disease incidence threshold used to introduce intervention
-
-# Here, the expectation is that the number of days under maximum intervention level
-# i.e., near a "lockdown" helps identify "c" and the number of deaths helps identify tau.
-# For this study, I use a simple Nelder-mead algorithm.
-# I find tau ~ 0.14, right in the middle of our expected range of 0.12 to 0.18.
-# And c is ~ 17, which is necessary to reproduce the number of days under the maximum intervention level.
-
-# Set targets:
-
-# And median deaths per 100k during first year of pandemic
-# Across US states
-augm_inputs <- readRDS("./data/archive/augm_inputs.rds")
-
-# Deaths target:
-# Average deaths per 100,000 population across US states:
-deaths_target <- augm_inputs$locationtimeseries %>%
-  filter(Date <= as.Date("2021-03-01")) %>%
-  group_by(LocationID) %>%
-  summarise(DeathsPer100K = sum(DeathsPer100K)) %>%
-  .$DeathsPer100K %>%
-  mean()
-
-# Number of days at maximum intervention level is computed
-# From Oxford's COVID-19 policy tracking project.
-
-# Data from:
-# https://github.com/OxCGRT/covid-policy-tracker/tree/master
-# Citation:  https://doi.org/10.1038/s41562-021-01079-8
-
-median_days_near_max_intervention <- read.csv("./data/OxCGRT_USA_latest.txt") %>%
-  dplyr::filter(Jurisdiction == "STATE_TOTAL") %>%
-  mutate(day = lubridate::as_date( as.character(Date))) %>%
-  filter(day <= as.Date("2021-03-01"), day >= as.Date("2020-03-01")) %>%
-  select(RegionName, Date, StringencyIndex_Average) %>%
-  group_by(RegionName) %>%
-  mutate(max_stringency = max(StringencyIndex_Average)) %>%
-  mutate(max_intervention = StringencyIndex_Average >= (max_stringency * (1-(1/5)))) %>%
-  summarise(days_max_intervention = sum(max_intervention), max_stringency = mean(max_stringency)) %>%
-  .$days_max_intervention %>%
-  median()
-
-
[…]

### 2024-03-01 R/scripts/calibration.R

@@ -1,0 +1,99 @@
+
+# Source all dependencies and model scripts
+source("./R/library.R")
+
+# This script is used to calibrate parameters tau and c,
+# tau: marginal effectiveness of NPIs (in reducing transmission)
+# c: disease incidence threshold used to introduce intervention
+
+# Here, the expectation is that the number of days under maximum intervention level
+# i.e., near a "lockdown" helps identify "c" and the number of deaths helps identify tau.
+# For this study, I use a simple Nelder-mead algorithm.
+# I find tau ~ 0.14, right in the middle of our expected range of 0.12 to 0.18.
+# And c is ~ 17, which is necessary to reproduce the number of days under the maximum intervention level.
+
+# Set targets:
+
+# And median deaths per 100k during first year of pandemic
+# Across US states
+augm_inputs <- readRDS("./data/archive/augm_inputs.rds")
+
+# Deaths target:
+# Average deaths per 100,000 population across US states:
+deaths_target <- augm_inputs$locationtimeseries %>%
+  filter(Date <= as.Date("2021-03-01")) %>%
+  group_by(LocationID) %>%
+  summarise(DeathsPer100K = sum(DeathsPer100K)) %>%
+  .$DeathsPer100K %>%
+  mean()
+
+# Number of days at maximum intervention level is computed
+# From Oxford's COVID-19 policy tracking project.
+
+# Data from:
+# https://github.com/OxCGRT/covid-policy-tracker/tree/master
+# Citation:  https://doi.org/10.1038/s41562-021-01079-8
+
+median_days_near_max_intervention <- read.csv("./data/OxCGRT_USA_latest.txt") %>%
+  dplyr::filter(Jurisdiction == "STATE_TOTAL") %>%
+  mutate(day = lubridate::as_date( as.character(Date))) %>%
+  filter(day <= as.Date("2021-03-01"), day >= as.Date("2020-03-01")) %>%
+  select(RegionName, Date, StringencyIndex_Average) %>%
+  group_by(RegionName) %>%
+  mutate(max_stringency = max(StringencyIndex_Average)) %>%
+  mutate(max_intervention = StringencyIndex_Average >= (max_stringency * (1-(1/5)))) %>%
+  summarise(days_max_intervention = sum(max_intervention), max_stringency = mean(max_stringency)) %>%
+  .$days_max_intervention %>%
+  median()
+
+
[…]

### 2024-03-01 R/scripts/run_analysis.R

@@ -1,308 +1,0 @@
-
-# Source all dependencies and model scripts
-source("./R/library.R")
-
-# results list
-
-if(!file.exists("./output/r.rds")) {
-  r <- list()
-} else {
-  r <- readRDS("./output/r.rds")
-}
-
-
-# 1. Set up experiments and constants ----------------------------------------
-
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
-
-base_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx", sheet = "one-way-scenarios")
-
-outcomes <- c("deaths_per_100k", "CH_illness", "CH_deaths", "CH", "L5_days", "L1plus_days", "CNPI", "C", "epi_size")
-
-# counterfactual scenarios without enhanced EWS
-count_scenarios <- base_scenarios %>%
-  filter(Section == "Scenarios") %>%
-  mutate(total_surv_lag = 13, p = 0.3, NMB_comparator = T)
-
-all_scenarios <- rbind(base_scenarios, count_scenarios)
-
-# create experiment
-base_experiment <- R6Experiment$new(model)
-
-base_experiment$set_design(grid_design_df = all_scenarios)
-
-
-# tau vs R0 experimental design
-
-tau <- c(seq.default(from = 0, to = 1, by = 0.025)/5)
-R0 <- 2.5 * c(0.5, 1, 1.5)
-
-EWS <- c(T,F)
-tau_design <- expand_grid(tau, EWS, R0) %>%
-  mutate(p = ifelse(EWS, 1, 0.3)) %>%
-  mutate(total_surv_lag = ifelse(EWS, 8,13))
-
-tau_experiment <- R6Experiment$new(model)
-
-tau_experiment$set_design(grid_design_df = tau_design)
-
-
-
-# 2. Run experiments ------------------------------------------------------
-
-## 2.1 Run main experiment -----------------------------------------------------
-
-base_results <- base_experiment$run(parallel = T,
-                                    cluster_eval_script = "./R/scripts/cluster_eval.R",
-                                    parallel::detectCores() - 4,
-                                    model_from_cluster_eval = T)
-
-
[…]

### 2024-03-01 R/scripts/tests.R

@@ -1,29 +1,0 @@
-# test of coordination
-
-test_coordination <- function(model){
-  # TEST FOR WHEN JURISDICTIONS 1 and 2 coordinate and jurisdiction 3 doesn't
-  g <- select(model$res_long, rep, step, jurisdiction.id, L)
-  h <- g %>% pivot_wider(names_from = jurisdiction.id, id_cols=c(step, rep), values_from= L)
-
-  # function to check if only has digits
-  numbers_only <- function(x) !grepl("\\D", x)
-
-
-  numerical_col_names = sort(as.numeric(colnames(h)[numbers_only(colnames(h))]))
-  for (row in 1:nrow(model$inputs$npi_coord)){
-    for (col in 1:ncol(model$inputs$npi_coord)){
-      if (model$inputs$npi_coord[row,col]==1){
-        # make sure coordinating jurisdictions have same NPI level
-        if (any(h[as.character(row)] != h[as.character(col)])){
-          warning("Jurisdictions not coordinating when they should!")
-        }
-      }
-      # Make sure non-coordinating ones have different NPI levels at some point
-      # NOTE: This may not always need to be true (maybe not if they have full mixing
-      # and the same responses to the same levels of COVID?)
-      else if (all(h[as.character(row)] == h[as.character(col)])){
-        warning("Jurisdictions MAY be coordinationg when they shouldn't")
-      }
-    }
-  }
-}

### 2024-03-01 archive/R/00_get_populations.R

@@ -1,0 +1,19 @@
+install.packages("usdata")
+
+library(usdata)
+library(dplyr)
+
+
+counties <- county_2019 %>%
+  select(state, name, fips, pop) %>%
+  filter(state %in% c("New York", "Connecticut", "New Jersey"))
+  #filter(state %in% c("California"))
+  #filter(state %in% c("Washington", "Oregon"))
+  group_by(state) %>%
+  summarise(n = n(),
+            pop = sum(pop))
+
+
+# I might be able to get data from tidycensus or from another dataset.
+library(tidycensus)
+

### 2024-03-01 archive/R/01_single_run_test.R

@@ -1,0 +1,117 @@
+
+#------------------------------------------------------------------------------#
+# Code repository for Analysis of Genomic Sequencing information
+#
+# Author: Pedro Nascimento de Lima
+# See README.md for information on usage and licensing
+#------------------------------------------------------------------------------#
+
+# This script demonstrates how to run the model for a single run:
+
+source("./R/library.R")
+
+# Single run examples -----------------------------------------------------
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file, "model1")
+#model2 <- OdinMetapop$new("stochastic_metapopulation2.R", "./data/metapopulation-inputs-heterogeneous.xlsx")
+
+model$simulate()
+
+
+model$set_input("c", 10)$
+  set_input("tau", 0.05)$
+  simulate()
+
+View(model$summary)
+
+
+get_outcome(model$summary, outcome_var = "CH")
+
+
+# all default model inputs are in the inputs object
+model$inputs$beta
+# IT looks like the mixing is not being partitioned correctly!
+#. it should be 0.1, 0.1
+
+# We can run the model with default inputs by using the simulate function
+# i.e., run for 100 days, for 100 replications
+View(model$simulate(step = 0:300, reps = 100))
+
+
+
+
+model2$simulate(step = 0:300, reps = 100)
+
+# Compare both models:
+res1 <- model$res_long %>%
+  group_by(rep, step) %>%
+  summarise(I = sum(E + P + I)) %>%
+  mutate(model = "homogeneous")
+
+res2 <- model2$res_long %>%
+  group_by(rep, step) %>%
+  summarise(I = sum(E + P + I)) %>%
+  mutate(model = "heterogeneous")
+
+all_res <-rbind(res1, res2)
+
+rep_model <- all_res %>%
+  select(rep, model) %>%
+  distinct() %>%
+  mutate(rep_model_id = row_number())
+
+all_res %>%
+  #left_join(rep_model) %>%
+  ggplot() +
+  geom_line(mapping = aes(x = step, y = I, group = interaction(model, rep), color = model)) #+
+  #facet_wrap(~model)
+
+
+
+
+
+
+
+model$inputs$R
+model$inputs$beta
+
+View(model$res_long %>% mutate(E_perc = E/(E+P+I), P_perc = P/(E+P+I), I_perc = I/(E+P+I)))
[…]

### 2024-03-01 archive/R/02_sample_analysis.R

@@ -1,0 +1,59 @@
+
+
+
+#------------------------------------------------------------------------------#
+# Code repository for Analysis of Genomic Sequencing information
+#
+# Author: Pedro Nascimento de Lima
+# See README.md for information on usage and licensing
+#------------------------------------------------------------------------------#
+
+# This script demonstrates how to run an experimental design using the model
+
+# Set up an experimental design and run the model:
+source("./R/library.R")
+
+# Instantiate model -------------------------------------------------------
+
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
+
+# Run full-factorial experiment -------------------------------------------
+
+# Create sample parameter set
+# This is here to demonstrate that data.frame with parameter sets (i.e., a
+# sample from the posterior distribution of model parameters)
+# can be assigned to the model. This is required for R6Sim models that
+# are used within the R6Experiment class, but not necessary otherwise.
+model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
+
+# Instantiate an experiment
+
+# Note that if the experiment is to be run in parallel, there is an underlying
+# assumption that the model will be instantiated in the cluster eval
+# script
+experiment <- R6Experiment$new(model)
+
+# Set experimental parameters
+experiment$
+  set_parameter("total_surv_lag", "grid", seq.default(from = 0, to = 21, by = 3))$ # Surveillance lag
+  set_parameter("p", "grid", c(0.5,1))$ # case ascertainment proportion
+  set_parameter("tau", "grid", c(0.12,0.15,0.18))$ # intevention effectiveness per intervention level
+  set_parameter("c", "grid", c(0.01,1,5,10,100))$ # case intervention threshold per 100,000 people
+  set_parameter("R0", "grid", c(1.5,2,2.5))$ # Initial reproduction number
+  set_parameter("cost_max_npi", "grid", c(0.1,0.25,0.87))$ # Initial reproduction number
+  set_parameter("L_c", "grid", c(0,1)) # Turn intervention coordination on and off
+
+
+# Set designs creates the experimental design data.frame
+experiment$set_design()
+
+# Run experiment (parallel option not available with odin at this time):
[…]

### 2024-03-01 archive/R/03_run_larger_exp_design.R

@@ -1,0 +1,110 @@
+
+
+
+#------------------------------------------------------------------------------#
+# Code repository for Analysis of Genomic Sequencing information
+#
+# Author: Pedro Nascimento de Lima
+# See README.md for information on usage and licensing
+#------------------------------------------------------------------------------#
+
+# This script demonstrates how to run an experiment using the model
+
+# Set up an experimental design and run the model:
+source("./R/library.R")
+
+# Instantiate model -------------------------------------------------------
+
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
+
+# Run full-factorial experiment -------------------------------------------
+
+# Create sample parameter set
+# This is here to demonstrate that data.frame with parameter sets (i.e., a
+# sample from the posterior distribution of model parameters)
+# can be assigned to the model. This is required for R6Sim models that
+# are used within the R6Experiment class, but not necessary otherwise.
+model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
+
+# Instantiate an experiment
+
+# Note that if the experiment is to be run in parallel, there is an underlying
+# assumption that the model will be instantiated in the cluster eval
+# script
+experiment <- R6Experiment$new(model)
+
+
+# Create an experimental design combining decision-maker characteristics and epi/social characteristics
+
+# Policy scenarios represent potential decision makers
+policy_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx",sheet = "policy_profile") %>%
+  mutate(policy.id = row_number())
+
+
+# Other variables
+total_surv_lag <- c(13,11,8,3)
+R0 <- c(1.5,2.5,3)
+cost_max_npi <- c(0.1,0.25,0.87, 1.25)
+#L_c <- c(0,1)
+r <- c(0.01, 0.001, 0.02)
+
+# Surveillance design
+surv_design <- data.frame(total_surv_lag = total_surv_lag) %>%
+  mutate(p = ifelse(total_surv_lag < 10, 1, 0.5)) %>%
+  mutate(surv.id = row_number())
+
+# Other variables design:
+other_vars_design <- expand_grid(
+  R0
+  ,cost_max_npi
+#  ,L_c
+  ,r
+) %>%
[…]

### 2024-03-01 archive/R/tests.R

@@ -1,0 +1,29 @@
+# test of coordination
+
+test_coordination <- function(model){
+  # TEST FOR WHEN JURISDICTIONS 1 and 2 coordinate and jurisdiction 3 doesn't
+  g <- select(model$res_long, rep, step, jurisdiction.id, L)
+  h <- g %>% pivot_wider(names_from = jurisdiction.id, id_cols=c(step, rep), values_from= L)
+
+  # function to check if only has digits
+  numbers_only <- function(x) !grepl("\\D", x)
+
+
+  numerical_col_names = sort(as.numeric(colnames(h)[numbers_only(colnames(h))]))
+  for (row in 1:nrow(model$inputs$npi_coord)){
+    for (col in 1:ncol(model$inputs$npi_coord)){
+      if (model$inputs$npi_coord[row,col]==1){
+        # make sure coordinating jurisdictions have same NPI level
+        if (any(h[as.character(row)] != h[as.character(col)])){
+          warning("Jurisdictions not coordinating when they should!")
+        }
+      }
+      # Make sure non-coordinating ones have different NPI levels at some point
+      # NOTE: This may not always need to be true (maybe not if they have full mixing
+      # and the same responses to the same levels of COVID?)
+      else if (all(h[as.character(row)] == h[as.character(col)])){
+        warning("Jurisdictions MAY be coordinationg when they shouldn't")
+      }
+    }
+  }
+}

### 2024-03-01 archive/set_ram.R

@@ -1,20 +1,0 @@
-
-
-
-# This script demonstrates how to run the model for a single run:
-
-source("./R/library.R")
-
-# Single run examples -----------------------------------------------------
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-# all default model inputs are in the inputs object
-model$inputs
-
-# We can run the model with default inputs by using the simulate function
-# i.e., run for 100 days, for 100 replications
-model$simulate(step = 0:100, reps = 100)
-
-model$set_input("coordination_mode", 2)$
-  simulate()
-

### 2024-03-01 paper/archive/manuscript_files/libs/quarto-diagram/mermaid-init.js

@@ -1,266 +1,0 @@
-// mermaid-init.js
-// Initializes the quarto-mermaid JS runtime
-//
-// Copyright (C) 2022 Posit Software, PBC
-
-/**
- * String.prototype.replaceAll() polyfill
- * https://gomakethings.com/how-to-replace-a-section-of-a-string-with-another-one-with-vanilla-js/
- * @author Chris Ferdinandi
- * @license MIT
- */
-if (!String.prototype.replaceAll) {
-  String.prototype.replaceAll = function (str, newStr) {
-    // If a regex pattern
-    if (
-      Object.prototype.toString.call(str).toLowerCase() === "[object regexp]"
-    ) {
-      return this.replace(str, newStr);
-    }
-
-    // If a string
-    return this.replace(new RegExp(str, "g"), newStr);
-  };
-}
-
-const mermaidOpts = {
-  startOnLoad: false,
-};
-// this CSS is adapted from
-// mkdocs-material
-// Copyright (c) 2016-2022 Martin Donath <martin.donath@squidfunk.com>
-
-const defaultCSS =
-  '.label text {fill: var(--mermaid-fg-color);}.node circle, .node ellipse, .node path, .node polygon, .node rect {fill: var(--mermaid-node-bg-color);stroke: var(--mermaid-node-fg-color);}marker {fill: var(--mermaid-edge-color) !important;}.edgeLabel .label rect {fill: #0000;}.label {color: var(--mermaid-label-fg-color);font-family: var(--mermaid-font-family);}.label foreignObject {line-height: normal;overflow: visible;}.label div .edgeLabel {color: var(--mermaid-label-fg-color);}.edgeLabel, .edgeLabel rect, .label div .edgeLabel {background-color: var(--mermaid-label-bg-color);}.edgeLabel, .edgeLabel rect {fill: var(--mermaid-label-bg-color);color: var(--mermaid-edge-color);}.edgePath .path, .flowchart-link {stroke: var(--mermaid-edge-color);}.edgePath .arrowheadPath {fill: var(--mermaid-edge-color);stroke: none;}.cluster rect {fill: var(--mermaid-fg-color--lightest);stroke: var(--mermaid-fg-color--lighter);}.cluster span {color: var(--mermaid-label-fg-color);font-family: var(--mermaid-font-family);}defs #flowchart-circleEnd, defs #flowchart-circleStart, defs #flowchart-crossEnd, defs #flowchart-crossStart, defs #
[…]

### 2024-03-01 run_analysis.R

@@ -1,0 +1,308 @@
+
+# Source all dependencies and model scripts
+source("./R/library.R")
+
+# results list
+
+if(!file.exists("./output/r.rds")) {
+  r <- list()
+} else {
+  r <- readRDS("./output/r.rds")
+}
+
+
+# 1. Set up experiments and constants ----------------------------------------
+
+model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
+
+model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
+
+base_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx", sheet = "one-way-scenarios")
+
+outcomes <- c("deaths_per_100k", "CH_illness", "CH_deaths", "CH", "L5_days", "L1plus_days", "CNPI", "C", "epi_size")
+
+# counterfactual scenarios without enhanced EWS
+count_scenarios <- base_scenarios %>%
+  filter(Section == "Scenarios") %>%
+  mutate(total_surv_lag = 13, p = 0.3, NMB_comparator = T)
+
+all_scenarios <- rbind(base_scenarios, count_scenarios)
+
+# create experiment
+base_experiment <- R6Experiment$new(model)
+
+base_experiment$set_design(grid_design_df = all_scenarios)
+
+
+# tau vs R0 experimental design
+
+tau <- c(seq.default(from = 0, to = 1, by = 0.025)/5)
+R0 <- 2.5 * c(0.5, 1, 1.5)
+
+EWS <- c(T,F)
+tau_design <- expand_grid(tau, EWS, R0) %>%
+  mutate(p = ifelse(EWS, 1, 0.3)) %>%
+  mutate(total_surv_lag = ifelse(EWS, 8,13))
+
+tau_experiment <- R6Experiment$new(model)
+
+tau_experiment$set_design(grid_design_df = tau_design)
+
+
+
+# 2. Run experiments ------------------------------------------------------
+
+## 2.1 Run main experiment -----------------------------------------------------
+
+base_results <- base_experiment$run(parallel = T,
+                                    cluster_eval_script = "./R/scripts/cluster_eval.R",
+                                    parallel::detectCores() - 4,
+                                    model_from_cluster_eval = T)
+
+
[…]

### 2024-03-02 R/dev/edge_cases_analysis.R
commenting out dev scripts

@@ -1,64 +1,0 @@
-
-
-#------------------------------------------------------------------------------#
-# Code for "The value of environmental surveillance for pandemic response"
-#
-# Author: Pedro Nascimento de Lima
-# See README.md for information on usage and licensing
-#------------------------------------------------------------------------------#
-
-# This script demonstrates how to run an experiment using the model
-
-# Set up an experimental design and run the model:
-source("./R/library.R")
-
-# Instantiate model -------------------------------------------------------
-
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-# base case:
-
-# Elimination:
-# At the near-elimination regime, a lower surveillance lag doesn't mean higher costs!
-model$set_input("c", 10000)$
-  set_input("R0", 2.5)$
-  set_input("p", 0.5)$
-  set_input("tau", 0.18)$
-  set_input("total_surv_lag", 1)$
-  set_input("cost_max_npi", 0.25)$
-  set_input("L_c", 1)$
-  set_input("a_up", 1)$
-  set_input("a_down", 45)$
-  simulate()
-
-
-#model$res_long %>% View()
-
-# Strict and Stringent scenario
-# In this scenario disease elimination is achieved
-
-reps_data <- model$res_long %>%
-  mutate(L_eff = floor(L)) %>%
-  filter(rep %in% 1:3)
-
-
-fig_intlevel <- reps_data %>%
-  ggplot(mapping = aes(x = step, group = rep, color = as.factor(rep))) +
-  geom_line(mapping = aes(y = NPI)) +
-  geom_line(mapping = aes(y = L), alpha = 0.2) +
-  facet_wrap(~jurisdiction.name) +
-  ylab("Intervention level") +
-  xlab("Days from pandemic onset")
-
-
-fig_prevalence <- reps_data %>%
-  ggplot(mapping = aes(x = step, group = rep, color = as.factor(rep))) +
-  geom_line(mapping = aes(y = (E + I + P) / population)) +
-  facet_wrap(~jurisdiction.name) +
-  scale_y_continuous(labels = scales::percent_format()) +
-  ylab("Prevalence") +
-  xlab("Days from pandemic onset")
-
-
-fig_intlevel / fig_prevalence
-

### 2024-03-02 R/dev/policy_profile_analysis.R

@@ -1,111 +1,0 @@
-
-
-#------------------------------------------------------------------------------#
-# Code for "The value of environmental surveillance for pandemic response"
-#
-# Author: Pedro Nascimento de Lima
-# See README.md for information on usage and licensing
-#------------------------------------------------------------------------------#
-
-# This script demonstrates how to run an experiment using the model
-
-# Set up an experimental design and run the model:
-source("./R/library.R")
-
-
-scenarios <- readxl::read_xlsx("./data/scenarios.xlsx",sheet = "policy_profile")
-
-
-# Instantiate model -------------------------------------------------------
-
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-
-all_long_res <- list()
-
-# Run the couple of scenarios and recover long results, under baseline assumptions
-
-for(i in 1:nrow(scenarios)) {
-
-  # Set inputs:
-  for(var in colnames(scenarios)) {
-
-    model$set_input(name = var, value = scenarios[[i,var]])
-
-  }
-
-  model$simulate(reps = 200)
-
-  all_long_res[[i]] <- model$res_long
-
-  all_long_res[[i]]$policy_profile <- scenarios$policy_profile[i]
-
-}
-
-
-all_long_df <- do.call(rbind, all_long_res)
-
-
-# Strict and Stringent scenario
-# In this scenario disease elimination is achieved
-
-reps_data <- all_long_df %>%
-  mutate(L_eff = floor(L)) %>%
-  filter(rep %in% 1:3) %>%
-  filter(jurisdiction.name == "NY")
-
-
-fig_intlevel <- reps_data %>%
-  ggplot(mapping = aes(x = step, group = rep, color = as.factor(rep))) +
-  geom_line(mapping = aes(y = NPI)) +
-  geom_line(mapping = aes(y = L), alpha = 0.2) +
-  facet_wrap(~policy_profile, nrow = 1) +
-  ylab("Intervention level") +
-  xlab("Days from pandemic onset")
-
-
-fig_prevalence <- reps_data %>%
-  ggplot(mapping = aes(x = step, group = rep, color = as.factor(rep))) +
-  geom_line(mapping = aes(y = I / population)) +
-  facet_wrap(~policy_profile, nrow = 1, scales = "free") +
[…]

### 2024-03-02 archive/R/00_get_populations.R

@@ -1,19 +1,19 @@
-install.packages("usdata")
-
-library(usdata)
-library(dplyr)
-
-
-counties <- county_2019 %>%
-  select(state, name, fips, pop) %>%
-  filter(state %in% c("New York", "Connecticut", "New Jersey"))
-  #filter(state %in% c("California"))
-  #filter(state %in% c("Washington", "Oregon"))
-  group_by(state) %>%
-  summarise(n = n(),
-            pop = sum(pop))
-
-
-# I might be able to get data from tidycensus or from another dataset.
-library(tidycensus)
-
+# install.packages("usdata")
+#
+# library(usdata)
+# library(dplyr)
+#
+#
+# counties <- county_2019 %>%
+#   select(state, name, fips, pop) %>%
+#   filter(state %in% c("New York", "Connecticut", "New Jersey"))
+#   #filter(state %in% c("California"))
+#   #filter(state %in% c("Washington", "Oregon"))
+#   group_by(state) %>%
+#   summarise(n = n(),
+#             pop = sum(pop))
+#
+#
+# # I might be able to get data from tidycensus or from another dataset.
+# library(tidycensus)
+#

### 2024-03-02 archive/R/02_sample_analysis.R

@@ -1,59 +1,59 @@
-
-
-
-#------------------------------------------------------------------------------#
-# Code for "The value of environmental surveillance for pandemic response"
 #
-# Author: Pedro Nascimento de Lima
-# See README.md for information on usage and licensing
-#------------------------------------------------------------------------------#
-
-# This script demonstrates how to run an experimental design using the model
-
-# Set up an experimental design and run the model:
-source("./R/library.R")
-
-# Instantiate model -------------------------------------------------------
-
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-# Run full-factorial experiment -------------------------------------------
-
-# Create sample parameter set
-# This is here to demonstrate that data.frame with parameter sets (i.e., a
-# sample from the posterior distribution of model parameters)
-# can be assigned to the model. This is required for R6Sim models that
-# are used within the R6Experiment class, but not necessary otherwise.
-model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
-
-# Instantiate an experiment
-
-# Note that if the experiment is to be run in parallel, there is an underlying
-# assumption that the model will be instantiated in the cluster eval
-# script
-experiment <- R6Experiment$new(model)
-
-# Set experimental parameters
-experiment$
-  set_parameter("total_surv_lag", "grid", seq.default(from = 0, to = 21, by = 3))$ # Surveillance lag
-  set_parameter("p", "grid", c(0.5,1))$ # case ascertainment proportion
-  set_parameter("tau", "grid", c(0.12,0.15,0.18))$ # intevention effectiveness per intervention level
-  set_parameter("c", "grid", c(0.01,1,5,10,100))$ # case intervention threshold per 100,000 people
-  set_parameter("R0", "grid", c(1.5,2,2.5))$ # Initial reproduction number
-  set_parameter("cost_max_npi", "grid", c(0.1,0.25,0.87))$ # Initial reproduction number
-  set_parameter("L_c", "grid", c(0,1)) # Turn intervention coordination on and off
-
-
-# Set designs creates the experimental design data.frame
-experiment$set_design()
-
-# Run experiment (parallel option not available with odin at this time):
[…]

### 2024-03-02 archive/R/03_run_larger_exp_design.R

@@ -1,110 +1,110 @@
-
-
-
-#------------------------------------------------------------------------------#
-# Code for "The value of environmental surveillance for pandemic response"
 #
-# Author: Pedro Nascimento de Lima
-# See README.md for information on usage and licensing
-#------------------------------------------------------------------------------#
-
-# This script demonstrates how to run an experiment using the model
-
-# Set up an experimental design and run the model:
-source("./R/library.R")
-
-# Instantiate model -------------------------------------------------------
-
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-# Run full-factorial experiment -------------------------------------------
-
-# Create sample parameter set
-# This is here to demonstrate that data.frame with parameter sets (i.e., a
-# sample from the posterior distribution of model parameters)
-# can be assigned to the model. This is required for R6Sim models that
-# are used within the R6Experiment class, but not necessary otherwise.
-model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")
-
-# Instantiate an experiment
-
-# Note that if the experiment is to be run in parallel, there is an underlying
-# assumption that the model will be instantiated in the cluster eval
-# script
-experiment <- R6Experiment$new(model)
-
-
-# Create an experimental design combining decision-maker characteristics and epi/social characteristics
-
-# Policy scenarios represent potential decision makers
-policy_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx",sheet = "policy_profile") %>%
-  mutate(policy.id = row_number())
-
-
-# Other variables
-total_surv_lag <- c(13,11,8,3)
-R0 <- c(1.5,2.5,3)
-cost_max_npi <- c(0.1,0.25,0.87, 1.25)
-#L_c <- c(0,1)
-r <- c(0.01, 0.001, 0.02)
-
-# Surveillance design
-surv_design <- data.frame(total_surv_lag = total_surv_lag) %>%
-  mutate(p = ifelse(total_surv_lag < 10, 1, 0.5)) %>%
-  mutate(surv.id = row_number())
-
-# Other variables design:
-other_vars_design <- expand_grid(
-  R0
-  ,cost_max_npi
-#  ,L_c
-  ,r
-) %>%
[…]

### 2024-03-02 archive/R/edge_cases_analysis.R

@@ -1,64 +1,65 @@
-
-
-#------------------------------------------------------------------------------#
-# Code for "The value of environmental surveillance for pandemic response"
 #
-# Author: Pedro Nascimento de Lima
-# See README.md for information on usage and licensing
-#------------------------------------------------------------------------------#
-
-# This script demonstrates how to run an experiment using the model
-
-# Set up an experimental design and run the model:
-source("./R/library.R")
-
-# Instantiate model -------------------------------------------------------
-
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-# base case:
-
-# Elimination:
-# At the near-elimination regime, a lower surveillance lag doesn't mean higher costs!
-model$set_input("c", 10000)$
-  set_input("R0", 2.5)$
-  set_input("p", 0.5)$
-  set_input("tau", 0.18)$
-  set_input("total_surv_lag", 1)$
-  set_input("cost_max_npi", 0.25)$
-  set_input("L_c", 1)$
-  set_input("a_up", 1)$
-  set_input("a_down", 45)$
-  simulate()
-
-
-#model$res_long %>% View()
-
-# Strict and Stringent scenario
-# In this scenario disease elimination is achieved
-
-reps_data <- model$res_long %>%
-  mutate(L_eff = floor(L)) %>%
-  filter(rep %in% 1:3)
-
-
-fig_intlevel <- reps_data %>%
-  ggplot(mapping = aes(x = step, group = rep, color = as.factor(rep))) +
-  geom_line(mapping = aes(y = NPI)) +
-  geom_line(mapping = aes(y = L), alpha = 0.2) +
-  facet_wrap(~jurisdiction.name) +
-  ylab("Intervention level") +
-  xlab("Days from pandemic onset")
-
-
-fig_prevalence <- reps_data %>%
-  ggplot(mapping = aes(x = step, group = rep, color = as.factor(rep))) +
-  geom_line(mapping = aes(y = (E + I + P) / population)) +
-  facet_wrap(~jurisdiction.name) +
-  scale_y_continuous(labels = scales::percent_format()) +
-  ylab("Prevalence") +
-  xlab("Days from pandemic onset")
-
-
-fig_intlevel / fig_prevalence
-
+#
+#
+# #------------------------------------------------------------------------------#
+# # Code for "The value of environmental surveillance for pandemic response"
+# #
[…]

### 2024-03-02 archive/R/jurisdiction-mixing.R

@@ -1,60 +1,60 @@
-
-
-
-# Approaches for inter-jurisdiction mixing matrices
-
-# 1 - Decomposed mixing matrix (similar to current PBM's approach, but without NPI weights)
-
-# First, we have a
-
-names.jurisdictions <- c("A", "B", "C", "D", "E")
-n.jurisdictions <- length(names.jurisdictions)
-
-names.modes <- c("Home", "Work", "Other")
-n.modes <- length(names.modes)
-# Following the same notation, k is the normalized contact rates across modes, by jurisdictions
-# k can be thought of the intensity of contacts across these mixing modes for each jurisdiction
-
-# For generality,
-k.mat <- matrix(data = c(0,0.5,0.5, 0.7,0.2,0.1), byrow = T, nrow = n.jurisdictions, ncol = n.modes, dimnames = list(names.jurisdictions, names.modes))
-
-# stop if rows are not normalized to 1
-stopifnot(!all((rowSums(k.mat) - 1)^2 > 1e-10))
-
-# now, for each mode, we have a mode-specific mixing matrix:
-home_mixing <-  diag(nrow = n.jurisdictions, ncol = n.jurisdictions, names = names.jurisdictions)
-
-# Work mixing needs to be an input from the spreadsheet:
-perc_work_home_county <- 0.9
-
-work_mixing <- diag(x = perc_work_home_county, nrow = n.jurisdictions, ncol = n.jurisdictions)
-
-work_mixing[work_mixing != perc_work_home_county] <- (1-perc_work_home_county)/(n.jurisdictions-1)
-
-rownames(work_mixing) <- names.jurisdictions
-colnames(work_mixing) <- names.jurisdictions
-
-# Other mixing:
-
-perc_other_home_county <- 0.7
-
-other_mixing <- diag(x = perc_other_home_county, nrow = n.jurisdictions, ncol = n.jurisdictions)
-
-other_mixing[other_mixing != perc_other_home_county] <- (1-perc_other_home_county)/(n.jurisdictions-1)
-
-rownames(other_mixing) <- names.jurisdictions
-colnames(other_mixing) <- names.jurisdictions
-
-# Concatenate all mixing mdoes into one array (order matters!):
-M <- array(c(home_mixing, work_mixing, other_mixing),dim = c(n.jurisdictions, n.jurisdictions, n.modes), dimnames = list(names.jurisdictions, names.jurisdictions, names.modes))
-
[…]

### 2024-03-02 archive/R/policy_profile_analysis.R

@@ -1,111 +1,111 @@
-
-
-#------------------------------------------------------------------------------#
-# Code for "The value of environmental surveillance for pandemic response"
 #
-# Author: Pedro Nascimento de Lima
-# See README.md for information on usage and licensing
-#------------------------------------------------------------------------------#
-
-# This script demonstrates how to run an experiment using the model
-
-# Set up an experimental design and run the model:
-source("./R/library.R")
-
-
-scenarios <- readxl::read_xlsx("./data/scenarios.xlsx",sheet = "policy_profile")
-
-
-# Instantiate model -------------------------------------------------------
-
-model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
-
-
-all_long_res <- list()
-
-# Run the couple of scenarios and recover long results, under baseline assumptions
-
-for(i in 1:nrow(scenarios)) {
-
-  # Set inputs:
-  for(var in colnames(scenarios)) {
-
-    model$set_input(name = var, value = scenarios[[i,var]])
-
-  }
-
-  model$simulate(reps = 200)
-
-  all_long_res[[i]] <- model$res_long
-
-  all_long_res[[i]]$policy_profile <- scenarios$policy_profile[i]
-
-}
-
-
-all_long_df <- do.call(rbind, all_long_res)
-
-
-# Strict and Stringent scenario
-# In this scenario disease elimination is achieved
-
-reps_data <- all_long_df %>%
-  mutate(L_eff = floor(L)) %>%
-  filter(rep %in% 1:3) %>%
-  filter(jurisdiction.name == "NY")
-
-
-fig_intlevel <- reps_data %>%
-  ggplot(mapping = aes(x = step, group = rep, color = as.factor(rep))) +
-  geom_line(mapping = aes(y = NPI)) +
-  geom_line(mapping = aes(y = L), alpha = 0.2) +
-  facet_wrap(~policy_profile, nrow = 1) +
-  ylab("Intervention level") +
-  xlab("Days from pandemic onset")
-
-
-fig_prevalence <- reps_data %>%
-  ggplot(mapping = aes(x = step, group = rep, color = as.factor(rep))) +
-  geom_line(mapping = aes(y = I / population)) +
-  facet_wrap(~policy_profile, nrow = 1, scales = "free") +
[…]

### 2024-03-02 archive/R/tests.R

@@ -1,29 +1,29 @@
-# test of coordination
-
-test_coordination <- function(model){
-  # TEST FOR WHEN JURISDICTIONS 1 and 2 coordinate and jurisdiction 3 doesn't
-  g <- select(model$res_long, rep, step, jurisdiction.id, L)
-  h <- g %>% pivot_wider(names_from = jurisdiction.id, id_cols=c(step, rep), values_from= L)
-
-  # function to check if only has digits
-  numbers_only <- function(x) !grepl("\\D", x)
-
-
-  numerical_col_names = sort(as.numeric(colnames(h)[numbers_only(colnames(h))]))
-  for (row in 1:nrow(model$inputs$npi_coord)){
-    for (col in 1:ncol(model$inputs$npi_coord)){
-      if (model$inputs$npi_coord[row,col]==1){
-        # make sure coordinating jurisdictions have same NPI level
-        if (any(h[as.character(row)] != h[as.character(col)])){
-          warning("Jurisdictions not coordinating when they should!")
-        }
-      }
-      # Make sure non-coordinating ones have different NPI levels at some point
-      # NOTE: This may not always need to be true (maybe not if they have full mixing
-      # and the same responses to the same levels of COVID?)
-      else if (all(h[as.character(row)] == h[as.character(col)])){
-        warning("Jurisdictions MAY be coordinationg when they shouldn't")
-      }
-    }
-  }
-}
+# # test of coordination
+#
+# test_coordination <- function(model){
+#   # TEST FOR WHEN JURISDICTIONS 1 and 2 coordinate and jurisdiction 3 doesn't
+#   g <- select(model$res_long, rep, step, jurisdiction.id, L)
+#   h <- g %>% pivot_wider(names_from = jurisdiction.id, id_cols=c(step, rep), values_from= L)
+#
+#   # function to check if only has digits
+#   numbers_only <- function(x) !grepl("\\D", x)
+#
+#
+#   numerical_col_names = sort(as.numeric(colnames(h)[numbers_only(colnames(h))]))
+#   for (row in 1:nrow(model$inputs$npi_coord)){
+#     for (col in 1:ncol(model$inputs$npi_coord)){
+#       if (model$inputs$npi_coord[row,col]==1){
+#         # make sure coordinating jurisdictions have same NPI level
+#         if (any(h[as.character(row)] != h[as.character(col)])){
+#           warning("Jurisdictions not coordinating when they should!")
+#         }
+#       }
[…]

### 2024-03-02 archive/reproduce-results.qmd
running styler

--- 2024-02-28 reproduce-results.qmd
@@
 # The r object contains the results objects used in this file
-if(!file.exists("./output/r.rds")) {
+if (!file.exists("./output/r.rds")) {
   source("./R/scripts/run_analysis.R")
@@
 }
-
 ```

### 2024-03-02 R/dev/jurisdiction-mixing.R

@@ -1,60 +1,0 @@
-
-
-
-# Approaches for inter-jurisdiction mixing matrices
-
-# 1 - Decomposed mixing matrix (similar to current PBM's approach, but without NPI weights)
-
-# First, we have a
-
-names.jurisdictions <- c("A", "B", "C", "D", "E")
-n.jurisdictions <- length(names.jurisdictions)
-
-names.modes <- c("Home", "Work", "Other")
-n.modes <- length(names.modes)
-# Following the same notation, k is the normalized contact rates across modes, by jurisdictions
-# k can be thought of the intensity of contacts across these mixing modes for each jurisdiction
-
-# For generality,
-k.mat <- matrix(data = c(0,0.5,0.5, 0.7,0.2,0.1), byrow = T, nrow = n.jurisdictions, ncol = n.modes, dimnames = list(names.jurisdictions, names.modes))
-
-# stop if rows are not normalized to 1
-stopifnot(!all((rowSums(k.mat) - 1)^2 > 1e-10))
-
-# now, for each mode, we have a mode-specific mixing matrix:
-home_mixing <-  diag(nrow = n.jurisdictions, ncol = n.jurisdictions, names = names.jurisdictions)
-
-# Work mixing needs to be an input from the spreadsheet:
-perc_work_home_county <- 0.9
-
-work_mixing <- diag(x = perc_work_home_county, nrow = n.jurisdictions, ncol = n.jurisdictions)
-
-work_mixing[work_mixing != perc_work_home_county] <- (1-perc_work_home_county)/(n.jurisdictions-1)
-
-rownames(work_mixing) <- names.jurisdictions
-colnames(work_mixing) <- names.jurisdictions
-
-# Other mixing:
-
-perc_other_home_county <- 0.7
-
-other_mixing <- diag(x = perc_other_home_county, nrow = n.jurisdictions, ncol = n.jurisdictions)
-
-other_mixing[other_mixing != perc_other_home_county] <- (1-perc_other_home_county)/(n.jurisdictions-1)
-
-rownames(other_mixing) <- names.jurisdictions
-colnames(other_mixing) <- names.jurisdictions
-
-# Concatenate all mixing mdoes into one array (order matters!):
-M <- array(c(home_mixing, work_mixing, other_mixing),dim = c(n.jurisdictions, n.jurisdictions, n.modes), dimnames = list(names.jurisdictions, names.jurisdictions, names.modes))
-
[…]

### 2024-03-02 R/functions/OdinMetapop.R

@@ -1,6 +1,3 @@
-
-
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -21,7 +18,6 @@ OdinMetapop <- R6::R6Class(
 
     # Pre-processes model inputs before running the simulation.
     pre_process_inputs = function() {
-
       # Create Mixing matrix and scales it as to align wit the desired R0.
 
       # Population density vector:
@@ -42,15 +38,15 @@ OdinMetapop <- R6::R6Class(
 
       # For this version, the jurisdictions spend the same time across mixing modes.
       # This can be changed by making k.mat an input itself
-      k.mat <- matrix(data = c(self$inputs$k_home,self$inputs$k_work_travel,self$inputs$k_other), byrow = T, nrow = n.jurisdictions, ncol = n.modes, dimnames = list(names.jurisdictions, names.modes))
+      k.mat <- matrix(data = c(self$inputs$k_home, self$inputs$k_work_travel, self$inputs$k_other), byrow = T, nrow = n.jurisdictions, ncol = n.modes, dimnames = list(names.jurisdictions, names.modes))
 
       # stop if rows are not normalized to 1
       stopifnot(!all((rowSums(k.mat) - 1)^2 > 1e-10))
 
       # now, for each mode, we have a mode-specific mixing matrix:
-      home_mixing <-  diag(nrow = n.jurisdictions, ncol = n.jurisdictions, names = names.jurisdictions)
+      home_mixing <- diag(nrow = n.jurisdictions, ncol = n.jurisdictions, names = names.jurisdictions)
 
-      other_mixing <-  diag(nrow = n.jurisdictions, ncol = n.jurisdictions, names = names.jurisdictions)
+      other_mixing <- diag(nrow = n.jurisdictions, ncol = n.jurisdictions, names = names.jurisdictions)
 
       # Construct work commuting based on "intra" and "inter" commuting rates
       # "Intra commuting" rate is the rate of commuting within commuting areas, as defined in the jurisdiction table.
@@ -65,35 +61,29 @@ OdinMetapop <- R6::R6Class(
       work_travel_mixing <- matrix(data = 0, nrow = self$inputs$nr_patches, ncol = self$inputs$nr_patches, dimnames = list(self$inputs$jurisdiction$jurisdiction.name, self$inputs$jurisdiction$jurisdiction.name))
 
       for (i in names.jurisdictions) {
-
[…]

### 2024-03-02 R/functions/OdinSim.R

@@ -1,6 +1,3 @@
-
-
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -48,13 +45,14 @@ OdinSim <- R6::R6Class(
     #' @param ... additional inputs to te odin model.
     #' @return s new `OdinSim` object.
     initialize = function(odin_file, inputs_file, model_name = "", ...) {
-
       super$initialize(name = odin_file)
 
       model_path <- paste0("./R/odin_models/", odin_file)
 
-      odin_workdir <- paste0("./cpp/", paste0(substr(odin_file,start = 1,
-                                                     stop = nchar(odin_file)-2), "_", model_name) , "/")
+      odin_workdir <- paste0("./cpp/", paste0(substr(odin_file,
+        start = 1,
+        stop = nchar(odin_file) - 2
+      ), "_", model_name), "/")
 
       # Get (and set) inputs from spreadsheet:
       self$get_inputs(inputs_file)
@@ -75,7 +73,7 @@ OdinSim <- R6::R6Class(
 
       odin_inputs_missing <- self$odin_parms[!self$odin_parms %in% names(self$inputs)]
 
-      if(!length(odin_inputs_missing)==0) {
+      if (!length(odin_inputs_missing) == 0) {
         stop(paste0("odin inputs missing: ", paste0(odin_inputs_missing, collapse = ", ")))
       }
 
@@ -84,9 +82,7 @@ OdinSim <- R6::R6Class(
 
       # The model now should be ready to be simulated.
       return(invisible(self))
-
     },
-
     pre_process_inputs = function() {
       stop("Function must be implemented in class that inherits this model")
     },
@@ -101,59 +97,53 @@ OdinSim <- R6::R6Class(
     #' @param value input value. Can be a single value, a list or a vector.
     #' @param type optional character string defining the type of input. Useful when one wants to only write inputs of a certain type to json.
     set_input = function(name, value, type = NA_character_) {
-
       # set.seed(1234)
 
       # If the model has been instantiated, and the input type is odin, set odin inptut
-      if(!is.null(self$o)) {
-
+      if (!is.null(self$o)) {
         # check that this is an input the odin model needs:
 
-        if(name %in% self$odin_parms) {
-
+        if (name %in% self$odin_parms) {
           input <- list()
           input[[name]] <- value
 
           do.call(self$o$set_user, input)
         }
-
       }
 
[…]

### 2024-03-02 R/functions/clear.labels.R

@@ -1,11 +1,9 @@
-
 # From: https://stackoverflow.com/questions/2394902/remove-variable-labels-attached-with-foreign-hmisc-spss-import-functions
 clear.labels <- function(x) {
-  if(is.list(x)) {
-    for(i in 1 : length(x)) class(x[[i]]) <- setdiff(class(x[[i]]), 'labelled')
-    for(i in 1 : length(x)) attr(x[[i]],"label") <- NULL
-  }
-  else {
+  if (is.list(x)) {
+    for (i in 1:length(x)) class(x[[i]]) <- setdiff(class(x[[i]]), "labelled")
+    for (i in 1:length(x)) attr(x[[i]], "label") <- NULL
+  } else {
     class(x) <- setdiff(class(x), "labelled")
     attr(x, "label") <- NULL
   }

### 2024-03-02 R/functions/logistic_fn.R

@@ -1,6 +1,3 @@
-
-
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -42,9 +39,9 @@
 #' @param scale_factor scaling_factor so that the x_transition time can be achieved
 #'
 #' @return a vector or a number for y
-logistic_fn <- function(y_max, x_mid_point, x_trans, x, scale_factor = 1/6) {
+logistic_fn <- function(y_max, x_mid_point, x_trans, x, scale_factor = 1 / 6) {
   # note scale_factor*x_trans *is* the scale parameter in the logistic growth function.
-  y_max / (1+exp((x_mid_point - x)*(scale_factor*x_trans)))
+  y_max / (1 + exp((x_mid_point - x) * (scale_factor * x_trans)))
 }
 
 # x is the scale_factor
@@ -57,7 +54,6 @@ logistic_fn <- function(y_max, x_mid_point, x_trans, x, scale_factor = 1/6) {
 #' @param x_trans transition duration in terms of x values
 #' @param x_vector a vector of x's over which the scale will be evaluated
 log_calib_obj_fn <- function(x, y_max, x_mid_point, x_trans, x_vector) {
-
   # compute ys for this scale factor
   ys <- logistic_fn(y_max, x_mid_point, x_trans, x = x_vector, scale_factor = x)
 
@@ -67,7 +63,6 @@ log_calib_obj_fn <- function(x, y_max, x_mid_point, x_trans, x_vector) {
   # target is to get the transition range right
   # Hence, we minimize objective function:
   (x_trans - (trans_range[2] - trans_range[1]))^2
-
 }
 
 
@@ -79,8 +74,6 @@ log_calib_obj_fn <- function(x, y_max, x_mid_point, x_trans, x_vector) {
 #' @param x_vector a vector of x's over which the scale will be evaluated
 #'
 calib_logistic_fn <- function(y_max, x_mid_point, x_trans, x_vector = seq.default(from = 0, to = x_mid_point * 3, by = 0.01)) {
-
-
   # The optimization routine does not work very well without a good
   # starting point and ranges for the search. Hence, I provide a
   # good starting point for the function, and verified that it worked.
@@ -103,22 +96,23 @@ calib_logistic_fn <- function(y_max, x_mid_point, x_trans, x_vector = seq.defaul
 
   # Then, solving for scale, we have that a good guess for the scale factor is:
 
[…]

### 2024-03-02 R/functions/summary_funs.R

@@ -1,36 +1,34 @@
-
-
 # Create summary functions list:
 # One can add more percentiles to the calculation by changing the vector below.
 alpha <- 0.95
-p <- c((1-alpha)/2, 1-(1-alpha)/2)
+p <- c((1 - alpha) / 2, 1 - (1 - alpha) / 2)
 p_names <- c(".lower", ".upper")
 
-summary_functions <- map(p, ~partial(quantile, probs = .x, na.rm = TRUE)) %>%
+summary_functions <- map(p, ~ partial(quantile, probs = .x, na.rm = TRUE)) %>%
   set_names(nm = p_names)
 
-#p_names <- map_chr(p, ~paste0(".q.",.x*100))
-summary_functions = map(p, ~partial(quantile, probs = .x, na.rm = TRUE)) %>%
+# p_names <- map_chr(p, ~paste0(".q.",.x*100))
+summary_functions <- map(p, ~ partial(quantile, probs = .x, na.rm = TRUE)) %>%
   set_names(nm = p_names)
-summary_functions$.mean = mean
+summary_functions$.mean <- mean
 
 # Get estimate in Text Form:
 
 pull_estimate <- function(df, ...) {
   df %>%
     dplyr::filter(...) %>%
-  .$estimate
+    .$estimate
 }
 
 get_outcome <- function(df, outcome_var, sig_digits = 3, ...) {
   r <- df %>%
     dplyr::filter(...) %>%
     as.data.frame() %>%
-    mutate(across(where(is.numeric), .fns = ~signif(x = .x, digits = sig_digits)))
+    mutate(across(where(is.numeric), .fns = ~ signif(x = .x, digits = sig_digits)))
 
-  vars = paste0(outcome_var, "_.", c("mean", "lower", "upper"))
+  vars <- paste0(outcome_var, "_.", c("mean", "lower", "upper"))
 
-  paste0(r[,vars[1]], " (", r[,vars[2]], "-", r[,vars[3]],")")
+  paste0(r[, vars[1]], " (", r[, vars[2]], "-", r[, vars[3]], ")")
 }
 
 

### 2024-03-02 R/functions/theme_rand.R

@@ -1,5 +1,3 @@
-
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -35,9 +33,11 @@
 #'   gapminder %>%
 #'   filter(continent == "Oceania")
 #'
-#' ggplot(gdp, aes(x = year,
-#'                 y = gdpPercap,
-#'                 color = country)) +
+#' ggplot(gdp, aes(
+#'   x = year,
+#'   y = gdpPercap,
+#'   color = country
+#' )) +
 #'   geom_line() +
 #'   scale_x_continuous(expand = c(0, 0)) +
 #'   scale_y_continuous(
@@ -80,8 +80,8 @@ theme_rand_ppt <- function(font = "Arial") {
     plot.subtitle = ggplot2::element_text(
       face = "bold",
       size = 12, # 12
-      family = "Arimo"#,
-      #margin = ggplot2::margin(t = 4)
+      family = "Arimo" # ,
+      # margin = ggplot2::margin(t = 4)
     ),
 
     # Caption
@@ -90,9 +90,11 @@ theme_rand_ppt <- function(font = "Arial") {
 
     # Axes
     axis.line = ggplot2::element_line(color = light_gray, linewidth = 0.5),
-    axis.line.y = ggplot2::element_line(color = light_gray, linewidth = 0.5), #ggplot2::element_blank(),
-    axis.text = ggplot2::element_text(size = 12,
-                                      color = dark_gray),
+    axis.line.y = ggplot2::element_line(color = light_gray, linewidth = 0.5), # ggplot2::element_blank(),
+    axis.text = ggplot2::element_text(
+      size = 12,
+      color = dark_gray
+    ),
     axis.text.x = ggplot2::element_text(margin = ggplot2::margin(t = 4)),
     axis.ticks.y = ggplot2::element_line(color = light_gray), # ggplot2::element_blank(),
     axis.ticks = ggplot2::element_line(color = light_gray),
@@ -112,14 +114,18 @@ theme_rand_ppt <- function(font = "Arial") {
     legend.spacing = ggplot2::unit(0, "cm"),
     legend.position = "top",
     legend.title = ggplot2::element_blank(),
-    legend.text = ggplot2::element_text(size = 10,
-                                        margin = ggplot2::margin(r = 8)),
+    legend.text = ggplot2::element_text(
+      size = 10,
+      margin = ggplot2::margin(r = 8)
+    ),
 
     # Strip
[…]

### 2024-03-02 R/functions/var.labels.R

@@ -1,7 +1,3 @@
-
-
-
-
 # Variale labels for consistent naming and labeling in tables and within the model.
 # Names can be re-used across data.frames
 

### 2024-03-02 R/library.R

@@ -1,4 +1,3 @@
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -42,10 +41,16 @@ s <- yaml::read_yaml("settings.yml")
 
 
 # Source all functions
-invisible(sapply(setdiff(list.files(path = s$lib_path, pattern = "*.R",full.names = T),
-                         s$lib_file),
-                 source,
-                 echo = F)
-)
+invisible(sapply(
+  setdiff(
+    list.files(path = s$lib_path, pattern = "*.R", full.names = T),
+    s$lib_file
+  ),
+  source,
+  echo = F
+))
 
-ggplot <- function(...) ggplot2::ggplot(...) + theme_rand_ppt(font = "Helvetica")
+ggplot <- function(...) {
+  ggplot2::ggplot(...) +
+    theme_rand_ppt(font = "Helvetica")
+}

### 2024-03-02 R/odin_models/stochastic_metapopulation.R

@@ -1,4 +1,3 @@
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -24,19 +23,19 @@ n <- nr_patches # the number of regions is equal to the length of most inputs
 # See https://mrc-ide.github.io/odin/articles/functions.html
 
 ## parameters
-beta[,] <- user()     # effective contact rate (S to e)
-sigma   <- user()     # progression rate from Exposed to Pre-symptomatic
-delta   <- user()     # rate of progression from pre-symptomatic to Infected
-gamma   <- user()     # rate of progression from active disease to Removed
-rho   <- user()       # prop asymptomatic
-tau     <- user()     # policy marginal effectiveness
-c       <- user()     # policy stringency
-t_o <- user()         # Time of "opening" (i.e., where NPIs are no longer used)
-total_surv_lag <- user()    # case Observation lag (in days)
-a_up <- user()           # NPI *a*djustment time
-a_down <- user()           # a decrease
-L_max <- user()       # max intervention level
-C[,] <- user()        # *C*: NPI coordination matrix. 1 if jurisdiction i follows NPI of jurisdiction j if jurisdiction's j NPI is more stringent.
+beta[, ] <- user() # effective contact rate (S to e)
+sigma <- user() # progression rate from Exposed to Pre-symptomatic
+delta <- user() # rate of progression from pre-symptomatic to Infected
+gamma <- user() # rate of progression from active disease to Removed
+rho <- user() # prop asymptomatic
+tau <- user() # policy marginal effectiveness
+c <- user() # policy stringency
+t_o <- user() # Time of "opening" (i.e., where NPIs are no longer used)
+total_surv_lag <- user() # case Observation lag (in days)
+a_up <- user() # NPI *a*djustment time
+a_down <- user() # a decrease
+L_max <- user() # max intervention level
[…]

### 2024-03-02 R/scripts/calibration.R

@@ -1,4 +1,3 @@
-
 # Source all dependencies and model scripts
 source("./R/library.R")
 
@@ -36,12 +35,12 @@ deaths_target <- augm_inputs$locationtimeseries %>%
 
 median_days_near_max_intervention <- read.csv("./data/OxCGRT_USA_latest.txt") %>%
   dplyr::filter(Jurisdiction == "STATE_TOTAL") %>%
-  mutate(day = lubridate::as_date( as.character(Date))) %>%
+  mutate(day = lubridate::as_date(as.character(Date))) %>%
   filter(day <= as.Date("2021-03-01"), day >= as.Date("2020-03-01")) %>%
   select(RegionName, Date, StringencyIndex_Average) %>%
   group_by(RegionName) %>%
   mutate(max_stringency = max(StringencyIndex_Average)) %>%
-  mutate(max_intervention = StringencyIndex_Average >= (max_stringency * (1-(1/5)))) %>%
+  mutate(max_intervention = StringencyIndex_Average >= (max_stringency * (1 - (1 / 5)))) %>%
   summarise(days_max_intervention = sum(max_intervention), max_stringency = mean(max_stringency)) %>%
   .$days_max_intervention %>%
   median()
@@ -55,11 +54,10 @@ model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)
 
 # Create a calibration objective functio, with some notional defaults for testing
 calib_tau_c_obj_fn <- function(x = c(c = 15, tau = 0.15), target_deaths = 160, target_l5_days = 100, deaths_weight = 0.5) {
-
   set.seed(1234)
 
   model$set_input("c", x[1])$
-        set_input("tau", x[2])
+    set_input("tau", x[2])
 
   model$simulate(reps = 100)
 
@@ -67,16 +65,15 @@ calib_tau_c_obj_fn <- function(x = c(c = 15, tau = 0.15), target_deaths = 160, t
   l5days_sim <- as.numeric(model$summary_all$L5_days_.mean[1])
 
 
-  sq_distance <- deaths_weight * (target_deaths - deaths_sim)^2 + (1-deaths_weight) * (target_l5_days - l5days_sim)^2
-  deaths_ape <- round(100*(deaths_sim - target_deaths)/target_deaths, 2)
-  l5days_ape <- round(100*(l5days_sim - target_l5_days)/target_l5_days, 2)
[…]

### 2024-03-02 R/scripts/cluster_eval.R

@@ -1,5 +1,3 @@
-
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -19,7 +17,7 @@
 # Source necessary functions:
 source("./R/library.R")
 
-#set.seed(1234)
+# set.seed(1234)
 
 # Instantiate model -------------------------------------------------------
 model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file)

### 2024-03-02 archive/R/00_get_populations.R

@@ -7,13 +7,14 @@ library(dplyr)
 counties <- county_2019 %>%
   select(state, name, fips, pop) %>%
   filter(state %in% c("New York", "Connecticut", "New Jersey"))
-  #filter(state %in% c("California"))
-  #filter(state %in% c("Washington", "Oregon"))
-  group_by(state) %>%
-  summarise(n = n(),
-            pop = sum(pop))
+# filter(state %in% c("California"))
+# filter(state %in% c("Washington", "Oregon"))
+group_by(state) %>%
+  summarise(
+    n = n(),
+    pop = sum(pop)
+  )
 
 
 # I might be able to get data from tidycensus or from another dataset.
 library(tidycensus)
-

### 2024-03-02 archive/R/01_single_run_test.R

@@ -1,4 +1,3 @@
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -12,7 +11,7 @@ source("./R/library.R")
 
 # Single run examples -----------------------------------------------------
 model <- OdinMetapop$new("stochastic_metapopulation.R", s$data_file, "model1")
-#model2 <- OdinMetapop$new("stochastic_metapopulation2.R", "./data/metapopulation-inputs-heterogeneous.xlsx")
+# model2 <- OdinMetapop$new("stochastic_metapopulation2.R", "./data/metapopulation-inputs-heterogeneous.xlsx")
 
 model$simulate()
 
@@ -30,7 +29,7 @@ get_outcome(model$summary, outcome_var = "CH")
 # all default model inputs are in the inputs object
 model$inputs$beta
 # IT looks like the mixing is not being partitioned correctly!
-#. it should be 0.1, 0.1
+# . it should be 0.1, 0.1
 
 # We can run the model with default inputs by using the simulate function
 # i.e., run for 100 days, for 100 replications
@@ -52,7 +51,7 @@ res2 <- model2$res_long %>%
   summarise(I = sum(E + P + I)) %>%
   mutate(model = "heterogeneous")
 
-all_res <-rbind(res1, res2)
+all_res <- rbind(res1, res2)
 
 rep_model <- all_res %>%
   select(rep, model) %>%
@@ -60,10 +59,10 @@ rep_model <- all_res %>%
   mutate(rep_model_id = row_number())
 
 all_res %>%
-  #left_join(rep_model) %>%
+  # left_join(rep_model) %>%
   ggplot() +
   geom_line(mapping = aes(x = step, y = I, group = interaction(model, rep), color = model)) #+
-  #facet_wrap(~model)
+# facet_wrap(~model)
 
 
 
@@ -74,16 +73,16 @@ all_res %>%
 model$inputs$R
 model$inputs$beta
 
-View(model$res_long %>% mutate(E_perc = E/(E+P+I), P_perc = P/(E+P+I), I_perc = I/(E+P+I)))
+View(model$res_long %>% mutate(E_perc = E / (E + P + I), P_perc = P / (E + P + I), I_perc = I / (E + P + I)))
 
 # Compute R0 from the first 60 days:
[…]

### 2024-03-02 archive/R/02_sample_analysis.R

@@ -1,6 +1,3 @@
-
-
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -36,24 +33,23 @@ experiment <- R6Experiment$new(model)
 # Set experimental parameters
 experiment$
   set_parameter("total_surv_lag", "grid", seq.default(from = 0, to = 21, by = 3))$ # Surveillance lag
-  set_parameter("p", "grid", c(0.5,1))$ # case ascertainment proportion
-  set_parameter("tau", "grid", c(0.12,0.15,0.18))$ # intevention effectiveness per intervention level
-  set_parameter("c", "grid", c(0.01,1,5,10,100))$ # case intervention threshold per 100,000 people
-  set_parameter("R0", "grid", c(1.5,2,2.5))$ # Initial reproduction number
-  set_parameter("cost_max_npi", "grid", c(0.1,0.25,0.87))$ # Initial reproduction number
-  set_parameter("L_c", "grid", c(0,1)) # Turn intervention coordination on and off
+  set_parameter("p", "grid", c(0.5, 1))$ # case ascertainment proportion
+  set_parameter("tau", "grid", c(0.12, 0.15, 0.18))$ # intevention effectiveness per intervention level
+  set_parameter("c", "grid", c(0.01, 1, 5, 10, 100))$ # case intervention threshold per 100,000 people
+  set_parameter("R0", "grid", c(1.5, 2, 2.5))$ # Initial reproduction number
+  set_parameter("cost_max_npi", "grid", c(0.1, 0.25, 0.87))$ # Initial reproduction number
+  set_parameter("L_c", "grid", c(0, 1)) # Turn intervention coordination on and off
 
 
 # Set designs creates the experimental design data.frame
 experiment$set_design()
 
 # Run experiment (parallel option not available with odin at this time):
-exp_results <- experiment$run(parallel = T,cluster_eval_script = "./R/scripts/sample_analysis_cluster_eval.R", n_cores = parallel::detectCores() - 3, model_from_cluster_eval = T)
+exp_results <- experiment$run(parallel = T, cluster_eval_script = "./R/scripts/sample_analysis_cluster_eval.R", n_cores = parallel::detectCores() - 3, model_from_cluster_eval = T)
 
 
 # Save output (but do not push this file):
 
[…]

### 2024-03-02 archive/R/03_run_larger_exp_design.R

@@ -1,6 +1,3 @@
-
-
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -37,15 +34,15 @@ experiment <- R6Experiment$new(model)
 # Create an experimental design combining decision-maker characteristics and epi/social characteristics
 
 # Policy scenarios represent potential decision makers
-policy_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx",sheet = "policy_profile") %>%
+policy_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx", sheet = "policy_profile") %>%
   mutate(policy.id = row_number())
 
 
 # Other variables
-total_surv_lag <- c(13,11,8,3)
-R0 <- c(1.5,2.5,3)
-cost_max_npi <- c(0.1,0.25,0.87, 1.25)
-#L_c <- c(0,1)
+total_surv_lag <- c(13, 11, 8, 3)
+R0 <- c(1.5, 2.5, 3)
+cost_max_npi <- c(0.1, 0.25, 0.87, 1.25)
+# L_c <- c(0,1)
 r <- c(0.01, 0.001, 0.02)
 
 # Surveillance design
@@ -55,10 +52,10 @@ surv_design <- data.frame(total_surv_lag = total_surv_lag) %>%
 
 # Other variables design:
 other_vars_design <- expand_grid(
-  R0
-  ,cost_max_npi
-#  ,L_c
-  ,r
+  R0,
+  cost_max_npi
+  #  ,L_c
+  , r
 ) %>%
   mutate(other.vars.id = row_number())
 
@@ -82,7 +79,7 @@ full_design <- expand_grid(policy_scenarios, surv_design, other_vars_design)
 experiment$set_design(grid_design_df = full_design)
 
 # Run experiment (parallel option not available with odin at this time):
-exp_results <- experiment$run(parallel = T,cluster_eval_script = "./R/scripts/cluster_eval.R", n_cores = parallel::detectCores() - 3, model_from_cluster_eval = T)
+exp_results <- experiment$run(parallel = T, cluster_eval_script = "./R/scripts/cluster_eval.R", n_cores = parallel::detectCores() - 3, model_from_cluster_eval = T)
 
 
 # Compute NMB counterfactuals:
@@ -92,7 +89,10 @@ ref_costs <- exp_results %>%
   select(policy.id, other.vars.id, ref_C)
 
 # Do we have the right number of rows?
[…]

### 2024-03-02 archive/R/edge_cases_analysis.R

@@ -1,5 +1,3 @@
-
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -32,7 +30,7 @@ model$set_input("c", 10000)$
   simulate()
 
 
-#model$res_long %>% View()
+# model$res_long %>% View()
 
 # Strict and Stringent scenario
 # In this scenario disease elimination is achieved
@@ -61,4 +59,3 @@ fig_prevalence <- reps_data %>%
 
 
 fig_intlevel / fig_prevalence
-

### 2024-03-02 archive/R/jurisdiction-mixing.R

@@ -1,6 +1,3 @@
-
-
-
 # Approaches for inter-jurisdiction mixing matrices
 
 # 1 - Decomposed mixing matrix (similar to current PBM's approach, but without NPI weights)
@@ -16,20 +13,20 @@ n.modes <- length(names.modes)
 # k can be thought of the intensity of contacts across these mixing modes for each jurisdiction
 
 # For generality,
-k.mat <- matrix(data = c(0,0.5,0.5, 0.7,0.2,0.1), byrow = T, nrow = n.jurisdictions, ncol = n.modes, dimnames = list(names.jurisdictions, names.modes))
+k.mat <- matrix(data = c(0, 0.5, 0.5, 0.7, 0.2, 0.1), byrow = T, nrow = n.jurisdictions, ncol = n.modes, dimnames = list(names.jurisdictions, names.modes))
 
 # stop if rows are not normalized to 1
 stopifnot(!all((rowSums(k.mat) - 1)^2 > 1e-10))
 
 # now, for each mode, we have a mode-specific mixing matrix:
-home_mixing <-  diag(nrow = n.jurisdictions, ncol = n.jurisdictions, names = names.jurisdictions)
+home_mixing <- diag(nrow = n.jurisdictions, ncol = n.jurisdictions, names = names.jurisdictions)
 
 # Work mixing needs to be an input from the spreadsheet:
 perc_work_home_county <- 0.9
 
 work_mixing <- diag(x = perc_work_home_county, nrow = n.jurisdictions, ncol = n.jurisdictions)
 
-work_mixing[work_mixing != perc_work_home_county] <- (1-perc_work_home_county)/(n.jurisdictions-1)
+work_mixing[work_mixing != perc_work_home_county] <- (1 - perc_work_home_county) / (n.jurisdictions - 1)
 
 rownames(work_mixing) <- names.jurisdictions
 colnames(work_mixing) <- names.jurisdictions
@@ -40,21 +37,19 @@ perc_other_home_county <- 0.7
 
 other_mixing <- diag(x = perc_other_home_county, nrow = n.jurisdictions, ncol = n.jurisdictions)
 
-other_mixing[other_mixing != perc_other_home_county] <- (1-perc_other_home_county)/(n.jurisdictions-1)
+other_mixing[other_mixing != perc_other_home_county] <- (1 - perc_other_home_county) / (n.jurisdictions - 1)
 
 rownames(other_mixing) <- names.jurisdictions
[…]

### 2024-03-02 archive/R/policy_profile_analysis.R

@@ -1,5 +1,3 @@
-
-
 #------------------------------------------------------------------------------#
 # Code for "The value of environmental surveillance for pandemic response"
 #
@@ -13,7 +11,7 @@
 source("./R/library.R")
 
 
-scenarios <- readxl::read_xlsx("./data/scenarios.xlsx",sheet = "policy_profile")
+scenarios <- readxl::read_xlsx("./data/scenarios.xlsx", sheet = "policy_profile")
 
 
 # Instantiate model -------------------------------------------------------
@@ -25,13 +23,10 @@ all_long_res <- list()
 
 # Run the couple of scenarios and recover long results, under baseline assumptions
 
-for(i in 1:nrow(scenarios)) {
-
+for (i in 1:nrow(scenarios)) {
   # Set inputs:
-  for(var in colnames(scenarios)) {
-
-    model$set_input(name = var, value = scenarios[[i,var]])
-
+  for (var in colnames(scenarios)) {
+    model$set_input(name = var, value = scenarios[[i, var]])
   }
 
   model$simulate(reps = 200)
@@ -39,7 +34,6 @@ for(i in 1:nrow(scenarios)) {
   all_long_res[[i]] <- model$res_long
 
   all_long_res[[i]]$policy_profile <- scenarios$policy_profile[i]
-
 }
 
 

### 2024-03-02 archive/R/tests.R

@@ -1,27 +1,27 @@
 # test of coordination
 
-test_coordination <- function(model){
+test_coordination <- function(model) {
   # TEST FOR WHEN JURISDICTIONS 1 and 2 coordinate and jurisdiction 3 doesn't
   g <- select(model$res_long, rep, step, jurisdiction.id, L)
-  h <- g %>% pivot_wider(names_from = jurisdiction.id, id_cols=c(step, rep), values_from= L)
+  h <- g %>% pivot_wider(names_from = jurisdiction.id, id_cols = c(step, rep), values_from = L)
 
   # function to check if only has digits
   numbers_only <- function(x) !grepl("\\D", x)
 
 
-  numerical_col_names = sort(as.numeric(colnames(h)[numbers_only(colnames(h))]))
-  for (row in 1:nrow(model$inputs$npi_coord)){
-    for (col in 1:ncol(model$inputs$npi_coord)){
-      if (model$inputs$npi_coord[row,col]==1){
+  numerical_col_names <- sort(as.numeric(colnames(h)[numbers_only(colnames(h))]))
+  for (row in 1:nrow(model$inputs$npi_coord)) {
+    for (col in 1:ncol(model$inputs$npi_coord)) {
+      if (model$inputs$npi_coord[row, col] == 1) {
         # make sure coordinating jurisdictions have same NPI level
-        if (any(h[as.character(row)] != h[as.character(col)])){
+        if (any(h[as.character(row)] != h[as.character(col)])) {
           warning("Jurisdictions not coordinating when they should!")
         }
       }
       # Make sure non-coordinating ones have different NPI levels at some point
       # NOTE: This may not always need to be true (maybe not if they have full mixing
       # and the same responses to the same levels of COVID?)
-      else if (all(h[as.character(row)] == h[as.character(col)])){
+      else if (all(h[as.character(row)] == h[as.character(col)])) {
         warning("Jurisdictions MAY be coordinationg when they shouldn't")
       }
     }

### 2024-03-02 run_analysis.R

@@ -14,7 +14,7 @@
 source("./R/library.R")
 
 # results list contains the results we want to save
-if(!file.exists("./output/r.rds")) {
+if (!file.exists("./output/r.rds")) {
   r <- list()
 } else {
   r <- readRDS("./output/r.rds")
@@ -46,13 +46,13 @@ base_experiment$set_design(grid_design_df = all_scenarios)
 
 # tau vs R0 experimental design
 
-tau <- c(seq.default(from = 0, to = 1, by = 0.025)/5)
+tau <- c(seq.default(from = 0, to = 1, by = 0.025) / 5)
 R0 <- 2.5 * c(0.5, 1, 1.5)
 
-EWS <- c(T,F)
+EWS <- c(T, F)
 tau_design <- expand_grid(tau, EWS, R0) %>%
   mutate(p = ifelse(EWS, 1, 0.3)) %>%
-  mutate(total_surv_lag = ifelse(EWS, 8,13))
+  mutate(total_surv_lag = ifelse(EWS, 8, 13))
 
 tau_experiment <- R6Experiment$new(model)
 
@@ -64,10 +64,12 @@ tau_experiment$set_design(grid_design_df = tau_design)
 
 ## 2.1 Run main experiment -----------------------------------------------------
 
-base_results <- base_experiment$run(parallel = T,
-                                    cluster_eval_script = "./R/scripts/cluster_eval.R",
-                                    parallel::detectCores() - 4,
-                                    model_from_cluster_eval = T)
+base_results <- base_experiment$run(
+  parallel = T,
+  cluster_eval_script = "./R/scripts/cluster_eval.R",
+  parallel::detectCores() - 4,
+  model_from_cluster_eval = T
+)
 
 
 # Compute differences at the replication level:
@@ -83,7 +85,7 @@ base_res_diff <- base_results %>%
   left_join(comp_results, by = join_by(rep, counterfactual.id))
 
 base_res_diff <- base_res_diff %>%
-  mutate(across( any_of(outcomes), ~ pull(base_res_diff, paste0(cur_column(), "_no_ews") ) - ., .names = "{.col}_diff") ) %>%
+  mutate(across(any_of(outcomes), ~ pull(base_res_diff, paste0(cur_column(), "_no_ews")) - ., .names = "{.col}_diff")) %>%
   rename(NMB = C_diff)
 
 
@@ -99,7 +101,7 @@ r$base_results_long <- base_results_rep_summaries %>%
[…]

### 2024-03-02 archive/R/tests.R
fixing tests

@@ -27,4 +27,3 @@
 #     }
 #   }
 # }
->>>>>>> 6a5bef483c6ac267a193fb8b9a2f458ab047e9ec

### 2024-03-02 archive/reults-draft.Rmd
results draft from google docs

--- 2024-03-02 archive/reproduce-results.qmd
@@
 ---
-title: "The value of early warning systems for pandemic response"
-format:
-  html:
-    self-contained: true
-execute: 
-  echo: true
-  warning: false
-  error: false
+title: "results-draft"
 author: "Pedro Nascimento de Lima"
+date: "2024-03-02"
+output: html_document
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
+```
 
@@
 Individuals can travel and work across jurisdiction lines, and the six jurisdictions each have a population of 10^6 individuals.
-Mortality parameters were chosen to reflect mortality rates observed in the United States during the first year of the COVID-19 pandemic, hence our results can be interpreted as expected outcomes for a cluster of US counties.
+Mortality parameters were chosen to reflect mortality rates observed in the United States during the first year of the COVID-19 pandemic; hence our results can be interpreted as expected outcomes for a cluster of US counties.
 
@@
 We find that $tau = 0.137$.
-We explore the implication of a wider range of policy effectiveness in sensitivity analyses, and explore this particular parameter in more details.
-Surprisingly, the policy effectiveness
+We explore the implication of a wider range of policy effectiveness in sensitivity analyses and explore this particular parameter in more detail.
 
@@
 Points from this section for discussion:
-* **EWS provide meaningful value under base-case assumptions.** Net-monetary benefit from surveillance system is on the order of 1,000 dollars per person over a one-year time-frame.
+* **EWS provides meaningful value under base-case assumptions.** Net-monetary benefit from the surveillance system is on the order of 1,000 dollars per person over a one-year time-frame.
 This outweighs any available estimate of costs.
@@
 
-## Value of EWS system depends on epidemiological and societal parameters
+## Surveillance net benefit depends on epidemiological and societal parameters
 
@@
 Such a pandemic would cause a substantial societal cost of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "C", NMB_comparator == T)` dollars per person.
-In those conditions, the EWS would be particularly valuable, reducing the death toll of the pandemic to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "deaths_per_100k", NMB_comparator == F)` per 100,000 people, and would be projected to provide a NMB of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
+In those conditions, the EWS would be particularly valuable, reducing the death toll of the pandemic to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "deaths_per_100k", NMB_comparator == F)` per 100,000 people, and would be projected to provide an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
 Conversely, a pathogen that is 50% less transmissible than SARS-Cov-2 would result in a lower death toll of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "deaths_per_100k", NMB_comparator == T)` without the EWS system, reducing the net-monetary benefit of the system to `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
-The value of the EWS system also depends on the infection fatality rate of the pathogen, but the net-benefit of the EWS would be positive even for a pathogen that is half as deadly as SARS-Cov-2.
+The value of the EWS system also depends on the infection fatality rate of the pathogen, but the net benefit of the EWS would be positive even for a pathogen that is half as deadly as SARS-Cov-2.
 All else equal, the EWS would still yield an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in a pandemic caused by a pathogen that is 50% less deadly than SARS-Cov-2, and would yield `r pull_estimate(r$table_1_long_all, Scenario == "1.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in responding to a pathogen that is 50% more deadly than COVID-19.
@@
 Longer decision and implementation lags have an overall negative effect on the total costs of the pandemic, longer decision lags also would *increase* the net benefit of EWS systems for pandemic response.
-Our base-case scenario assume that decision-makers take one week to decide and implement non-pharmaceutical interventions.
-If instead policymakers took two weeks to introduce and implement interventions, the total pandemic cost increases from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "C", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == T)` dollars per person, assuming conventional surveillance.
-The early warning system effectively helps to roll-back decision lags, reducing total costs to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == F)` dollars per person, yielding a NMB of `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "NMB", NMB_comparator == F)` relative to conventional surveillance.
+Our base-case scenario assumes that decision-makers take one week to decide and implement non-pharmaceutical interventions.
+If policymakers took two weeks to introduce and implement interventions, the total pandemic cost increases from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "C", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == T)` dollars per person, assuming conventional surveillance.
+The early warning system effectively helps to roll-back decision lags, reducing total costs to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == F)` dollars per person, yielding an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "NMB", NMB_comparator == F)` relative to conventional surveillance.
 Conversely, in contexts where NPI implementation is swift (i.e, decisions are made within 3 days), the NMB of EWS would be lower but still positive at `r pull_estimate(r$table_1_long_all, Scenario == "3-day decision lag", variable == "NMB", NMB_comparator == F)` dollars per person.
@@
 In that scenario, the EWS yields `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "NMB", NMB_comparator == F)` dollars per person of net benefit since most of the deaths would already be prevented through more stringent NPIs.
-Conversely, if policy makers were more lenient policy choices would change the syst overall outcomes but also imply a higher value of the enhanced surveillance system.
+Conversely, if policymakers were more lenient policy choices would change the system overall outcomes but also imply a higher value of the enhanced surveillance system.
 
@@
 If NPI effectiveness is 0%, an improved surveillance system has no value, as expected.
-The value of the enhanced surveillance system increases with NPI effectiveness slowly for maximum NPI effectiveness below around 50%, and peaked when the maximum NPI effectiveness reaches 60%, enough to reduce the baseline R0 of 2.5 below 1 at pandemic outset.
+The value of the enhanced surveillance system increases with NPI effectiveness slowly for maximum NPI effectiveness below around 50%, and peaks when the maximum NPI effectiveness reaches 60%, enough to reduce the baseline R0 of 2.5 below 1 at the pandemic onset.
 In general, we find that the value of the EWS peaks at a maximum NPI effectiveness value at $1-1/R0$, for different values of R0.
 Beyond that point, increases in NPI effectiveness result in lower total pandemic costs, but the value of the surveillance system starts to decrease.
-
-<!--
-
-# If we have time to look at those other cases:
-
-## EWS and policy coordination
-EWS alleviate lack of policy coordination. 
-There is a concern that EWS data will need to be integrated to be useful. However, we find that policy coordination and EWS act as substitutes.
-
-## If we have time: Marginal Value of EWS
-EWS deployment also face decreasing marginal returns. The value of EWS decreases as coverage increases.
-
--->

### 2024-03-02 archive/reults-draft.Rmd
more results draft updates

@@
 Individuals can travel and work across jurisdiction lines, and the six jurisdictions each have a population of 10^6 individuals.
-Mortality parameters were chosen to reflect mortality rates observed in the United States during the first year of the COVID-19 pandemic; hence our results can be interpreted as expected outcomes for a cluster of US counties.
+Mortality parameters were chosen to reflect mortality rates observed in the United States during the first year of the COVID-19 pandemic; hence, our results can be interpreted as expected outcomes for a cluster of US counties.
 
@@
 We also account for economic costs caused by the nonpharmaceutical interventions introduced (discussed in the Supplementary Appendix).
-All outcomes are computed over a one-year time frame from the pandemic onset and are averages computed across the jurisdictions in our model.
-Since the model is stochastic, we simulate the model 1,000 times and present means and 95% prediction intervals (computed with the 2.5% and the 97.5% percentiles) of each outcome.
+All outcomes are computed over one year from the pandemic onset and are averages computed across the jurisdictions in our model.
+Since the model is stochastic, we simulate the model 1,000 times and present means and 95% prediction intervals (calculated with the 2.5% and the 97.5% percentiles) of each outcome.
 
-# Calibration: Add as a footnote to tau.
-We use parameter estimates from the literature whenever possible (Supplementary Table 1).
-Parameters that are unobserved in the data, notably tau and.
-R0 is computed based on disease progression rates.
+# Calibration
+We used parameter estimates from the literature whenever possible (Supplementary Table 1).
+However, two of our model parameters need calibration so that model outcomes in our base-case surveillance scenario match the number of deaths observed in the US during the first year of the COVID-19 pandemic.
+We calibrate two parameters that cannot be directly observed from data: $tau$ and $c$.
+Our calibration procedure uses two data targets: i) the average number of COVID-19 deaths per 100,000 population from March 1st 2020 to March 1st 2021 in US states, and ii) the median number of days spent at the highest intervention level adopted in US states, defined as a Policy Stringency of 95% or more relative to the maximum NPI stringency level, as measured by the Oxford COVID-19 NPI policy database.
+We compute a mean square error distance function, equally weighing the deaths and the days under maximum intervention level,  choosing $tau$ and $c$ to minimize that distance function.
+In this calibration exercise, $tau = 0.14$, and $c = 17.8$, which are plausible values.
+As a validation target, we observed taht the epidemic size at the end of the first year of the pandemic was on the order of 30% (cite specific numbers), which is above US-level seroprevalence survey estimates at that point in time (~20%).
+Since seroprevalence surveys only provide a lower bound to the epidemic size due to waning immunity, we found our calibration to appropriately capture the order of magnitude of reduction in mixing caused by NPI policies, under the assumption that the baseline R0 in the US was 2.5.
+
+ the optimization search    R0 is computed based on disease progression rates.
 We calibrate only the $tau$ parameter, which is highly heterogeneous and context-dependent (cite Laura), to match the number of deaths observed during the first year in the United States.
@@
 ## Sensitivity Analyses
-We explore the performance of the EWS under conditions that depart from COVID-19 by varying model parameters one at a time, aiming to represent a new emerging pathogen against which the population has no immunity.
-Since a probabilistic characterization of uncertainty to parameters is elusive, we refrain from making probabilistic statements
+We explore the performance of the EWS under conditions that depart from COVID-19 by varying model parameters one at a time, aiming to represent a new emerging pathogen for which the population has no immunity.
 
@@
 
-**MAybe this is useful for a briefing or slides, but maybe not the paper**:
-An early warning system that shifts the epidemiological time-series used to inform non-pharmaceutical decisions can change the course of disease when used with effective interventions (Figure 1).
-We consider a network of jurisdictions that are independently managed by their own public health departments.
-For simplicity, we consider two cases: One in which the jurisdictions base their decisions upon syndromic surveillance data, and another (counterfactual) scenario where they are able to use data from environmental sampling sources (i.e., without loss of generality, wastewater surveillance), which lead the conventional surveillance data by a few days.
-The two sets of jurisdictions are otherwise equal, and face the exact same pandemic.
+An early warning system that shifts the epidemiological time series used to inform non-pharmaceutical decisions can change the course of the pandemic with effective interventions (Figure 1).
+We consider a network of jurisdictions managed by independent public health departments.
+For simplicity, we consider two cases: One in which the jurisdictions base their decisions upon syndromic surveillance and another (counterfactual) scenario where they can use data from environmental sampling sources (i.e., without loss of generality, wastewater surveillance), which leads syndromic surveillance data by a few days.
+The two sets of jurisdictions are otherwise equal and face the same pandemic.
 
@@
 
-<!--EWS systems provide meaningful value under base-case assumptions representing a hypothetical new pandemic with parameters mirroring those of SARS-Cov-2. -->
-
 Assuming no NPIs interventions are put in place (No NPIs scenario), a new COVID-19-like pandemic would produce significant damage, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` deaths per 100,000 people [mean (95% prediction interval of stochastic replications)] over a one-year time-frame (Table 1).
-In the absence of any behavioral change, the unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to illness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
+In the absence of any behavioral change, the unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, costing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to illness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
 The total health cost would be projected at `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH")` dollars per person.
 
-If policymakers followed the base-case NPI policy simulated in this paper, the death toll of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` deaths per 100,000 population if no enhanced early-warning system (EWS) is adopted (NPIs w/o EWS scenario).
-As it was the case during the COVID-19 pandemic, such an achievement would involve a substantial public health effort.
+If policymakers followed our base-case NPI policy, the death toll of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` deaths per 100,000 population if a baseline syndromic surveillance system is used to trigger NPIs (NPIs w/o EWS scenario).
+Such an achievement would involve a substantial public health effort.
 Under baseline assumptions, this effort would require `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L1plus_days")` days of public health interventions, of which `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L5_days")` days are spent under the most-restrictive policy intervention level.

### 2024-03-02 install_dependencies.R

@@ -11,9 +11,11 @@
 # then run renv::restore() to install the same package versions we used or use renv::install().
 
 # CRAN packages
-cran_packages <- c("Hmisc", "R6", "tidyr", "odin", "dplyr", "ggplot2", "dde",
-                   "gt", "lemon", "lubridate", "parallel", "purrr", "readxl",
-                   "scales", "writexl", "yaml", "pkgbuild", "pkgload")
+cran_packages <- c(
+  "Hmisc", "R6", "tidyr", "odin", "dplyr", "ggplot2", "dde",
+  "gt", "lemon", "lubridate", "parallel", "purrr", "readxl",
+  "scales", "writexl", "yaml", "pkgbuild", "pkgload"
+)
 
 install.packages(cran_packages)
 
@@ -21,4 +23,3 @@ install.packages(cran_packages)
 remotes::install_github("https://github.com/RANDCorporation/randplot")
 
 remotes::install_github("https://github.com/RANDCorporation/R6Sim")
-

### 2024-04-02 R/scripts/run_main_experiment.R
separating out the number fo replications across the runs

@@ -74,7 +74,7 @@ base_results <- base_experiment$run(
   cluster_eval_script = s$cluster_eval_script,
   n_cores = s$n_cores,
   model_from_cluster_eval = s$model_from_cluster_eval,
-  reps = s$n_reps,
+  reps = s$n_reps_main,
   seed = s$seed
 )
 

### 2024-04-02 R/scripts/run_tau_experiment.R

@@ -62,7 +62,7 @@ tau_results <- tau_experiment$run(
   cluster_eval_script = s$cluster_eval_script,
   n_cores = s$n_cores,
   model_from_cluster_eval = s$model_from_cluster_eval,
-  reps = s$n_reps,
+  reps = s$n_reps_tau,
   seed = s$seed
 )
 

### 2024-04-02 archive/reults-draft.Rmd
saving n_reps for testing

@@
 
-This `quarto` document produces the results presented in *Nascimento de Lima et al.
+This document is an early draft of the results section in *Nascimento de Lima et al.
 (2024).
 The value of environmental surveillance for pandemic response*.
+We use it to pull results from the results object programmatically.
 For details on how to install dependencies, see the `README.md` file.
@@
 
-# Scenarios
-Supplementary Table 1 provides details on parameter values used under our baseline scenario.
-The baseline scenario with conventional surveillance assumes a syndromic surveillance system based upon confirmed case counts, and there is a thirteen-day time span from exposure to the pathogen to detection by the public health authority.
-This thirteen-day lag covers the disease incubation phase (6.6 days)(cite) and testing and reporting delays (6 days).(cite) Consistent with estimates from the COVID-19 pandemic, the conventional surveillance system has a 30% case ascertainment rate.(cite) We assume the environmental sampling system can provide absolute, unbiased disease incidence estimates that lead the case counts by five days in the base-case scenario.
-We explore longer and shorter lead times, as this lag depends on two to ten days.
-We use 5 days as the base-case scenario, as wastewater surveillance systems have been able to demonstrate.
-(cite) Moreover, this analysis assumed that all the populations within the jurisdictions are covered by the EWS (or equivalently, that the sampled population is a representative sample of all the population).
-
-# Population
-The model simulates disease percolating through six interconnected jurisdictions that independently make nonpharmaceutical decisions.
-Individuals can travel and work across jurisdiction lines, and the six jurisdictions each have a population of 10^6 individuals.
-Mortality parameters were chosen to reflect mortality rates observed in the United States during the first year of the COVID-19 pandemic; hence, our results can be interpreted as expected outcomes for a cluster of US counties.
-
-# Outcomes
-The average mortality per 100,000 population across the jurisdiction is the primary health outcome in this study.
-We also compute the cost of illness due to disease, using SARS-Cov-2 as the base-case scenario.
-We also account for economic costs caused by the nonpharmaceutical interventions introduced (discussed in the Supplementary Appendix).
-All outcomes are computed over one year from the pandemic onset and are averages computed across the jurisdictions in our model.
-Since the model is stochastic, we simulate the model 1,000 times and present means and 95% prediction intervals (calculated with the 2.5% and the 97.5% percentiles) of each outcome.
-
-# Calibration
-We used parameter estimates from the literature whenever possible (Supplementary Table 1).
-However, two of our model parameters need calibration so that model outcomes in our base-case surveillance scenario match the number of deaths observed in the US during the first year of the COVID-19 pandemic.
-We calibrate two parameters that cannot be directly observed from data: $tau$ and $c$.
-Our calibration procedure uses two data targets: i) the average number of COVID-19 deaths per 100,000 population from March 1st 2020 to March 1st 2021 in US states, and ii) the median number of days spent at the highest intervention level adopted in US states, defined as a Policy Stringency of 95% or more relative to the maximum NPI stringency level, as measured by the Oxford COVID-19 NPI policy database.
-We compute a mean square error distance function, equally weighing the deaths and the days under maximum intervention level,  choosing $tau$ and $c$ to minimize that distance function.
-In this calibration exercise, $tau = 0.14$, and $c = 17.8$, which are plausible values.
-As a validation target, we observed taht the epidemic size at the end of the first year of the pandemic was on the order of 30% (cite specific numbers), which is above US-level seroprevalence survey estimates at that point in time (~20%).
-Since seroprevalence surveys only provide a lower bound to the epidemic size due to waning immunity, we found our calibration to appropriately capture the order of magnitude of reduction in mixing caused by NPI policies, under the assumption that the baseline R0 in the US was 2.5.
-
- the optimization search    R0 is computed based on disease progression rates.
-We calibrate only the $tau$ parameter, which is highly heterogeneous and context-dependent (cite Laura), to match the number of deaths observed during the first year in the United States.
-We find that $tau = 0.137$.
-We explore the implication of a wider range of policy effectiveness in sensitivity analyses and explore this particular parameter in more detail.
-
-## Sensitivity Analyses
-We explore the performance of the EWS under conditions that depart from COVID-19 by varying model parameters one at a time, aiming to represent a new emerging pathogen for which the population has no immunity.
-
 # Results
-
-## Early warning systems can shape disease and policy dynamics
-
-An early warning system that shifts the epidemiological time series used to inform non-pharmaceutical decisions can change the course of the pandemic with effective interventions (Figure 1).
-We consider a network of jurisdictions managed by independent public health departments.
-For simplicity, we consider two cases: One in which the jurisdictions base their decisions upon syndromic surveillance and another (counterfactual) scenario where they can use data from environmental sampling sources (i.e., without loss of generality, wastewater surveillance), which leads syndromic surveillance data by a few days.
-The two sets of jurisdictions are otherwise equal and face the same pandemic.
 

### 2024-04-19 archive/reults-draft.Rmd
renaming EWS to ESS

@@
 
-## Value of EWS under baseline assumptions
+## Value of ESS under baseline assumptions
 
@@
 
-If policymakers followed our base-case NPI policy, the death toll of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` deaths per 100,000 population if a baseline syndromic surveillance system is used to trigger NPIs (NPIs w/o EWS scenario).
+If policymakers followed our base-case NPI policy, the death toll of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "deaths_per_100k")` deaths per 100,000 population if a baseline syndromic surveillance system is used to trigger NPIs (NPIs w/o ESS scenario).
 Such an achievement would involve a substantial public health effort.
-Under baseline assumptions, this effort would require `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L1plus_days")` days of public health interventions, of which `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L5_days")` days are spent under the most-restrictive policy intervention level.
-We project this effort to cost on the order of `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "CNPI")` per person.
-Based on those assumptions, the total cost of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "C")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "C")` per person.
+Under baseline assumptions, this effort would require `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "L1plus_days")` days of public health interventions, of which `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "L5_days")` days are spent under the most-restrictive policy intervention level.
+We project this effort to cost on the order of `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "CNPI")` per person.
+Based on those assumptions, the total cost of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "C")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "C")` per person.
 
-Under baseline assumptions, the EWS provides net-positive value to society because they enhance the effectiveness of NPIs (i.e., preventing illness and death) more than they increase the costs of additional interventions introduced due to the early warning.
-Even an early warning system that provides only a 2-day early warning relative to a conventional pandemic response system (NPI + 2-day EWS scenario) is projected to reduce the pandemic death toll from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 2-day EWS", variable == "deaths_per_100k")` deaths per 100,000 population.
+Under baseline assumptions, the ESS provides net-positive value to society because they enhance the effectiveness of NPIs (i.e., preventing illness and death) more than they increase the costs of additional interventions introduced due to the early warning.
+Even an early warning system that provides only a 2-day early warning relative to a conventional pandemic response system (NPI + 2-day ESS scenario) is projected to reduce the pandemic death toll from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 2-day ESS", variable == "deaths_per_100k")` deaths per 100,000 population.
 Health benefits increase if the system can provide earlier warning.
-A system that provides a 5-day early warning (NPIs + 5-day EWS scenario) would be projected to result in `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "deaths_per_100k")` deaths per 100,000 population, and a 10-day early warning would result in `r pull_estimate(r$table_1_long, Scenario == "NPIs + 10-day EWS", variable == "deaths_per_100k")` deaths per 100,000 people.
+A system that provides a 5-day early warning (NPIs + 5-day ESS scenario) would be projected to result in `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day ESS", variable == "deaths_per_100k")` deaths per 100,000 population, and a 10-day early warning would result in `r pull_estimate(r$table_1_long, Scenario == "NPIs + 10-day ESS", variable == "deaths_per_100k")` deaths per 100,000 people.
 
-The use of early warning would result in earlier public health intervention, increasing the number of days of under NPIs from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L1plus_days")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "L1plus_days")` days for a system providing a 5-day early warning signal.
+The use of early warning would result in earlier public health intervention, increasing the number of days of under NPIs from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "L1plus_days")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day ESS", variable == "L1plus_days")` days for a system providing a 5-day early warning signal.
 That said, the number of days at the *maximum* intervention level (i.e., a lockdown) was projected to *decrease* with the early warning system, as earlier action prevented surges from reaching alarming levels that prompted blunt interventions.
-For example, a system that provided a 5-day early warning reduced the number of days under the maximum intervention level from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "L5_days")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "L5_days")` days.
-Still, the net effect of the EWS was to marginally increase NPI intervention costs, from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "CNPI")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "CNPI")` dollars per person, assuming a 5-day early warning.
-Health and economic outcomes weighted, the early warning system reduced the total pandemic costs from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o EWS", variable == "C")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "C")` dollars per person.
-The difference between the two costs is the net-monetary benefit of the system, which is estimated at `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day EWS", variable == "NMB")` dollars per person in the first year of the pandemic.
+For example, a system that provided a 5-day early warning reduced the number of days under the maximum intervention level from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "L5_days")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day ESS", variable == "L5_days")` days.
+Still, the net effect of the ESS was to marginally increase NPI intervention costs, from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "CNPI")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day ESS", variable == "CNPI")` dollars per person, assuming a 5-day early warning.
+Health and economic outcomes weighted, the early warning system reduced the total pandemic costs from `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "C")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day ESS", variable == "C")` dollars per person.
+The difference between the two costs is the net-monetary benefit of the system, which is estimated at `r pull_estimate(r$table_1_long, Scenario == "NPIs + 5-day ESS", variable == "NMB")` dollars per person in the first year of the pandemic.
 
@@
 Points from this section for discussion:
-* **EWS provides meaningful value under base-case assumptions.** Net-monetary benefit from the surveillance system is on the order of 1,000 dollars per person over a one-year time-frame.
+* **ESS provides meaningful value under base-case assumptions.** Net-monetary benefit from the surveillance system is on the order of 1,000 dollars per person over a one-year time-frame.
 This outweighs any available estimate of costs.
@@
 Such a pandemic would cause a substantial societal cost of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "C", NMB_comparator == T)` dollars per person.
-In those conditions, the EWS would be particularly valuable, reducing the death toll of the pandemic to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "deaths_per_100k", NMB_comparator == F)` per 100,000 people, and would be projected to provide an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
-Conversely, a pathogen that is 50% less transmissible than SARS-Cov-2 would result in a lower death toll of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "deaths_per_100k", NMB_comparator == T)` without the EWS system, reducing the net-monetary benefit of the system to `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
-The value of the EWS system also depends on the infection fatality rate of the pathogen, but the net benefit of the EWS would be positive even for a pathogen that is half as deadly as SARS-Cov-2.
-All else equal, the EWS would still yield an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in a pandemic caused by a pathogen that is 50% less deadly than SARS-Cov-2, and would yield `r pull_estimate(r$table_1_long_all, Scenario == "1.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in responding to a pathogen that is 50% more deadly than COVID-19.
+In those conditions, the ESS would be particularly valuable, reducing the death toll of the pandemic to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "deaths_per_100k", NMB_comparator == F)` per 100,000 people, and would be projected to provide an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
+Conversely, a pathogen that is 50% less transmissible than SARS-Cov-2 would result in a lower death toll of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "deaths_per_100k", NMB_comparator == T)` without the ESS system, reducing the net-monetary benefit of the system to `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
+The value of the ESS system also depends on the infection fatality rate of the pathogen, but the net benefit of the ESS would be positive even for a pathogen that is half as deadly as SARS-Cov-2.
+All else equal, the ESS would still yield an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in a pandemic caused by a pathogen that is 50% less deadly than SARS-Cov-2, and would yield `r pull_estimate(r$table_1_long_all, Scenario == "1.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in responding to a pathogen that is 50% more deadly than COVID-19.
 
@@
 (Supplementary Figure 1, **Supplementary Table X**).
-Despite that increase, the NMB of the system would be only marginally reduced from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "NMB", NMB_comparator == F)` to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x NPI costs", variable == "NMB", NMB_comparator == F)` dollars per person, because the EWS system only marginally increased the number of days under interventions.
+Despite that increase, the NMB of the system would be only marginally reduced from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "NMB", NMB_comparator == F)` to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x NPI costs", variable == "NMB", NMB_comparator == F)` dollars per person, because the ESS system only marginally increased the number of days under interventions.
 
@@
 
-Longer decision and implementation lags have an overall negative effect on the total costs of the pandemic, longer decision lags also would *increase* the net benefit of EWS systems for pandemic response.
+Longer decision and implementation lags have an overall negative effect on the total costs of the pandemic, longer decision lags also would *increase* the net benefit of ESS systems for pandemic response.
 Our base-case scenario assumes that decision-makers take one week to decide and implement non-pharmaceutical interventions.
@@
 The early warning system effectively helps to roll-back decision lags, reducing total costs to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == F)` dollars per person, yielding an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "NMB", NMB_comparator == F)` relative to conventional surveillance.
-Conversely, in contexts where NPI implementation is swift (i.e, decisions are made within 3 days), the NMB of EWS would be lower but still positive at `r pull_estimate(r$table_1_long_all, Scenario == "3-day decision lag", variable == "NMB", NMB_comparator == F)` dollars per person.
+Conversely, in contexts where NPI implementation is swift (i.e, decisions are made within 3 days), the NMB of ESS would be lower but still positive at `r pull_estimate(r$table_1_long_all, Scenario == "3-day decision lag", variable == "NMB", NMB_comparator == F)` dollars per person.
 
-Policy stringency (i.e., how stringent the case thresholds used to introduce or remove interventions) has a similar effect on the value of EWS relative to decision lags - the EWS provides the most value under the least desirable conditions for disease prevention.
-In a context where decision-makers are 50% more stringent than in our base-case assumptions (i.e., their case thresholds are half of its base-case), society is able to avert a substantial number of deaths, reducing mortality from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "deaths_per_100k", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "deaths_per_100k", NMB_comparator == T)` deaths per 100,000 population without the EWS.
-In that scenario, the EWS yields `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "NMB", NMB_comparator == F)` dollars per person of net benefit since most of the deaths would already be prevented through more stringent NPIs.
+Policy stringency (i.e., how stringent the case thresholds used to introduce or remove interventions) has a similar effect on the value of ESS relative to decision lags - the ESS provides the most value under the least desirable conditions for disease prevention.
+In a context where decision-makers are 50% more stringent than in our base-case assumptions (i.e., their case thresholds are half of its base-case), society is able to avert a substantial number of deaths, reducing mortality from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "deaths_per_100k", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "deaths_per_100k", NMB_comparator == T)` deaths per 100,000 population without the ESS.
+In that scenario, the ESS yields `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "NMB", NMB_comparator == F)` dollars per person of net benefit since most of the deaths would already be prevented through more stringent NPIs.
 Conversely, if policymakers were more lenient policy choices would change the system overall outcomes but also imply a higher value of the enhanced surveillance system.
@@
 The value of the enhanced surveillance system increases with NPI effectiveness slowly for maximum NPI effectiveness below around 50%, and peaks when the maximum NPI effectiveness reaches 60%, enough to reduce the baseline R0 of 2.5 below 1 at the pandemic onset.
-In general, we find that the value of the EWS peaks at a maximum NPI effectiveness value at $1-1/R0$, for different values of R0.
+In general, we find that the value of the ESS peaks at a maximum NPI effectiveness value at $1-1/R0$, for different values of R0.
 Beyond that point, increases in NPI effectiveness result in lower total pandemic costs, but the value of the surveillance system starts to decrease.

### 2024-04-19 R/scripts/run_main_experiment.R

@@ -45,7 +45,7 @@ r$base_scenarios <- readxl::read_xlsx("./data/scenarios.xlsx", sheet = "one-way-
 
 r$outcomes <- c("deaths_per_100k", "CH_illness", "CH_deaths", "CH", "L5_days", "L1plus_days", "CNPI", "C", "epi_size")
 
-# counterfactual scenarios without enhanced EWS
+# counterfactual scenarios without enhanced ESS
 count_scenarios <- r$base_scenarios %>%
   filter(Section == "Scenarios") %>%
   mutate(total_surv_lag = 13, p = 0.3, NMB_comparator = T)
@@ -85,7 +85,7 @@ comp_results <- base_results %>%
   filter(NMB_comparator) %>%
   select(rep, counterfactual.id, any_of(r$outcomes)) %>%
   clear.labels() %>%
-  rename_at(vars(any_of(r$outcomes)), function(x) paste0(x, "_no_ews"))
+  rename_at(vars(any_of(r$outcomes)), function(x) paste0(x, "_no_ESS"))
 
 
 base_res_diff <- base_results %>%
@@ -93,7 +93,7 @@ base_res_diff <- base_results %>%
   left_join(comp_results, by = join_by(rep, counterfactual.id))
 
 base_res_diff <- base_res_diff %>%
-  mutate(across(any_of(r$outcomes), ~ pull(base_res_diff, paste0(cur_column(), "_no_ews")) - ., .names = "{.col}_diff")) %>%
+  mutate(across(any_of(r$outcomes), ~ pull(base_res_diff, paste0(cur_column(), "_no_ESS")) - ., .names = "{.col}_diff")) %>%
   rename(NMB = C_diff)
 
 

### 2024-04-19 R/scripts/run_tables_figures.R

@@ -78,7 +78,7 @@ r$table_1 <- r$table_1_long %>%
   relocate(labels) %>%
   rename(Outcome = labels) %>%
   filter(!is.na(Outcome)) %>%
-  relocate(Outcome, `No NPIs`, `NPIs w/o EWS`, `NPIs + 2-day EWS`, `NPIs + 5-day EWS`, `NPIs + 10-day EWS`)
+  relocate(Outcome, `No NPIs`, `NPIs w/o ESS`, `NPIs + 2-day ESS`, `NPIs + 5-day ESS`, `NPIs + 10-day ESS`)
 
 writexl::write_xlsx(r$table_1, path = "./output/table_1.xlsx")
 
@@ -86,21 +86,21 @@ writexl::write_xlsx(r$table_1, path = "./output/table_1.xlsx")
 
 r$sup_table_2 <- r$table_1_long_all %>%
   filter(Section == "Scenarios") %>%
-  mutate(EWS = ifelse(NMB_comparator, "N", "Y")) %>%
+  mutate(ESS = ifelse(NMB_comparator, "N", "Y")) %>%
   mutate(variable = factor(variable, levels = c("epi_size", "CH_illness", "deaths_per_100k", "deaths_per_100k_diff", "CH_deaths", "CH", "L1plus_days", "L5_days", "CNPI", "C", "NMB"), ordered = T)) %>%
   filter(variable %in% c("deaths_per_100k", "CH", "CNPI", "C", "NMB")) %>%
   left_join(var.labels.df, by = join_by(variable)) %>%
-  select(Scenario, EWS, labels, estimate) %>%
-  pivot_wider(id_cols = c(Scenario, EWS), names_from = labels, values_from = estimate) %>%
+  select(Scenario, ESS, labels, estimate) %>%
+  pivot_wider(id_cols = c(Scenario, ESS), names_from = labels, values_from = estimate) %>%
   mutate(Scenario = factor(Scenario, levels = unique(r$base_scenarios$Scenario), ordered = T)) %>%
-  arrange(Scenario, EWS)
+  arrange(Scenario, ESS)
 
 writexl::write_xlsx(r$sup_table_2, path = "./output/sup_table_2.xlsx")
 
 # 4. Figures -----------------------------------------------------------------
 
 r$fig_1_data <- r$table_1_long_numeric %>%
[…]

### 2024-04-19 R/scripts/run_tau_experiment.R

@@ -46,10 +46,10 @@ model$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param =
 tau <- c(seq.default(from = 0, to = 1, by = 0.025) / 5)
 R0 <- 2.5 * c(0.5, 1, 1.5)
 
-EWS <- c(T, F)
-tau_design <- expand_grid(tau, EWS, R0) %>%
-  mutate(p = ifelse(EWS, 1, 0.3)) %>%
-  mutate(total_surv_lag = ifelse(EWS, 8, 13))
+ESS <- c(T, F)
+tau_design <- expand_grid(tau, ESS, R0) %>%
+  mutate(p = ifelse(ESS, 1, 0.3)) %>%
+  mutate(total_surv_lag = ifelse(ESS, 8, 13))
 
 tau_experiment <- R6Experiment$new(model)
 
@@ -70,30 +70,30 @@ msg_time("Post-processing tau experiment.")
 
 r$fig_effectiveness_data <- tau_results %>%
   mutate(effectiveness = tau * 5) %>%
-  select(effectiveness, EWS, any_of(r$outcomes)) %>%
-  group_by(effectiveness, EWS) %>%
+  select(effectiveness, ESS, any_of(r$outcomes)) %>%
+  group_by(effectiveness, ESS) %>%
   summarise_all(summary_functions) %>%
-  mutate(EWS = ifelse(EWS, "NPIs w/ EWS", "NPIs w/o EWS"))
+  mutate(ESS = ifelse(ESS, "NPIs w/ ESS", "NPIs w/o ESS"))
 
-# Comparator scenarios without EWS
+# Comparator scenarios without ESS
 tau_comp <- tau_results %>%
-  filter(!EWS) %>%
+  filter(!ESS) %>%
   select(rep, tau, R0, C) %>%
-  rename(C_no_EWS = C)
+  rename(C_no_ESS = C)
 
 # Net monetary benefit
 tau_NMB <- tau_results %>%
-  filter(EWS) %>%
+  filter(ESS) %>%
   left_join(tau_comp, by = join_by(rep, tau, R0)) %>%
   mutate(effectiveness = tau * 5) %>%
-  mutate(NMB = C_no_EWS - C)
+  mutate(NMB = C_no_ESS - C)
 
 # Net monetary benefit summary stats
 r$tau_NMB_summary <- tau_NMB %>%
-  group_by(effectiveness, R0, EWS) %>%
[…]

### 2024-04-30 archive/results-archive/reproduce-results.qmd
update figures to pdf

--- 2024-04-19 archive/reults-draft.Rmd
@@
 ---
-title: "results-draft"
+title: "The value of early warning systems for pandemic response"
+format:
+  html:
+    self-contained: true
+execute: 
+  echo: true
+  warning: false
+  error: false
 author: "Pedro Nascimento de Lima"
-date: "2024-03-02"
-output: html_document
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
-```
 
@@
 
-This document is an early draft of the results section in *Nascimento de Lima et al.
+This `quarto` document produces the results presented in *Nascimento de Lima et al.
 (2024).
 The value of environmental surveillance for pandemic response*.
-We use it to pull results from the results object programmatically.
 For details on how to install dependencies, see the `README.md` file.
@@
 
+# Scenarios
+Supplementary Table 1 provides details on parameter values used under our baseline scenario.
+The baseline scenario with conventional surveillance assumes a syndromic surveillance system based upon confirmed case counts, and there is a thirteen-day time span from exposure to the pathogen to detection by the public health authority.
+This thirteen-day lag covers the disease incubation phase (6.6 days)(cite) and testing and reporting delays (6 days).(cite) Consistent with estimates from the COVID-19 pandemic, the conventional surveillance system has a 30% case ascertainment rate.(cite) We assume the environmental sampling system can provide absolute, unbiased disease incidence estimates that lead the case counts by five days in the base-case scenario.
+We explore longer and shorter lead times, as this lag depends on  two to ten days.
+We use 5 days as the base-case scenario, as wastewater surveillance systems have been able to demonstrate.
+(cite) Moreover, this analysis assumed that all the populations within the jurisdictions are covered by the ESS (or equivalently, that the sampled population is a representative sample of all the population).
+
+# Population
+The model simulates disease percolating through six interconnected jurisdictions that independently make nonpharmaceutical decisions.
+Individuals can travel and work across jurisdiction lines, and the six jurisdictions each have a population of 10^6 individuals.
+Mortality parameters were chosen to reflect mortality rates observed in the United States during the first year of the COVID-19 pandemic, hence our results can be interpreted as expected outcomes for a cluster of US counties.
+
+# Outcomes
+The average mortality per 100,000 population across the jurisdiction is the primary health outcome in this study.
+We also compute the cost of illness due to disease, using SARS-Cov-2 as the base-case scenario.
+We also account for economic costs caused by the nonpharmaceutical interventions introduced (discussed in the Supplementary Appendix).
+All outcomes are computed over a one-year time frame from the pandemic onset and are averages computed across the jurisdictions in our model.
+Since the model is stochastic, we simulate the model 1,000 times and present means and 95% prediction intervals (computed with the 2.5% and the 97.5% percentiles) of each outcome.
+
+# Calibration: Add as a footnote to tau.
+We use parameter estimates from the literature whenever possible (Supplementary Table 1).
+Parameters that are unobserved in the data, notably tau and.
+R0 is computed based on disease progression rates.
+We calibrate only the $tau$ parameter, which is highly heterogeneous and context-dependent (cite Laura), to match the number of deaths observed during the first year in the United States.
+We find that $tau = 0.137$.
+We explore the implication of a wider range of policy effectiveness in sensitivity analyses, and explore this particular parameter in more details.
+Surprisingly, the policy effectiveness
+
+## Sensitivity Analyses
+We explore the performance of the ESS under conditions that depart from COVID-19 by varying model parameters one at a time, aiming to represent a new emerging pathogen against which the population has no immunity.
+Since a probabilistic characterization of uncertainty to parameters is elusive, we refrain from making probabilistic statements
+
 # Results
+
+## Early warning systems can shape disease and policy dynamics
+
+**MAybe this is useful for a briefing or slides, but maybe not the paper**:
+An early warning system that shifts the epidemiological time-series used to inform non-pharmaceutical decisions can change the course of disease when used with effective interventions (Figure 1).
+We consider a network of jurisdictions that are independently managed by their own public health departments.
+For simplicity, we consider two cases: One in which the jurisdictions base their decisions upon syndromic surveillance data, and another (counterfactual) scenario where they are able to use data from environmental sampling sources (i.e., without loss of generality, wastewater surveillance), which lead the conventional surveillance data by a few days.
+The two sets of jurisdictions are otherwise equal, and face the exact same pandemic.
 
@@
 
+<!--ESS systems provide meaningful value under base-case assumptions representing a hypothetical new pandemic with parameters mirroring those of SARS-Cov-2. -->
+
 Assuming no NPIs interventions are put in place (No NPIs scenario), a new COVID-19-like pandemic would produce significant damage, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` deaths per 100,000 people [mean (95% prediction interval of stochastic replications)] over a one-year time-frame (Table 1).
-In the absence of any behavioral change, the unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, costing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to illness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
+In the absence of any behavioral change, the unmitigated pandemic would infect `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "epi_size")` percent of the population, causing `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_illness")` dollars per person due to illness, and `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH_deaths")` dollars due to deaths.
 The total health cost would be projected at `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "CH")` dollars per person.
 
-If policymakers followed our base-case NPI policy, the death toll of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "deaths_per_100k")` deaths per 100,000 population if a baseline syndromic surveillance system is used to trigger NPIs (NPIs w/o ESS scenario).
-Such an achievement would involve a substantial public health effort.
+If policymakers followed the base-case NPI policy simulated in this paper, the death toll of the pandemic would be reduced from `r pull_estimate(r$table_1_long, Scenario == "No NPIs", variable == "deaths_per_100k")` to `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "deaths_per_100k")` deaths per 100,000 population if no enhanced early-warning system (ESS) is adopted (NPIs w/o ESS scenario).
+As it was the case during the COVID-19 pandemic, such an achievement would involve a substantial public health effort.
 Under baseline assumptions, this effort would require `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "L1plus_days")` days of public health interventions, of which `r pull_estimate(r$table_1_long, Scenario == "NPIs w/o ESS", variable == "L5_days")` days are spent under the most-restrictive policy intervention level.
@@
 Points from this section for discussion:
-* **ESS provides meaningful value under base-case assumptions.** Net-monetary benefit from the surveillance system is on the order of 1,000 dollars per person over a one-year time-frame.
+* **ESS provide meaningful value under base-case assumptions.** Net-monetary benefit from surveillance system is on the order of 1,000 dollars per person over a one-year time-frame.
 This outweighs any available estimate of costs.
@@
 
-## Surveillance net benefit depends on epidemiological and societal parameters
+## Value of ESS system depends on epidemiological and societal parameters
 
@@
 Such a pandemic would cause a substantial societal cost of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "C", NMB_comparator == T)` dollars per person.
-In those conditions, the ESS would be particularly valuable, reducing the death toll of the pandemic to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "deaths_per_100k", NMB_comparator == F)` per 100,000 people, and would be projected to provide an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
+In those conditions, the ESS would be particularly valuable, reducing the death toll of the pandemic to `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "deaths_per_100k", NMB_comparator == F)` per 100,000 people, and would be projected to provide a NMB of `r pull_estimate(r$table_1_long_all, Scenario == "1.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
 Conversely, a pathogen that is 50% less transmissible than SARS-Cov-2 would result in a lower death toll of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "deaths_per_100k", NMB_comparator == T)` without the ESS system, reducing the net-monetary benefit of the system to `r pull_estimate(r$table_1_long_all, Scenario == "0.5x transmissible", variable == "NMB", NMB_comparator == F)` dollars per person.
-The value of the ESS system also depends on the infection fatality rate of the pathogen, but the net benefit of the ESS would be positive even for a pathogen that is half as deadly as SARS-Cov-2.
+The value of the ESS system also depends on the infection fatality rate of the pathogen, but the net-benefit of the ESS would be positive even for a pathogen that is half as deadly as SARS-Cov-2.
 All else equal, the ESS would still yield an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "0.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in a pandemic caused by a pathogen that is 50% less deadly than SARS-Cov-2, and would yield `r pull_estimate(r$table_1_long_all, Scenario == "1.5x deadly", variable == "NMB", NMB_comparator == F)` dollars per person in responding to a pathogen that is 50% more deadly than COVID-19.
@@
 Longer decision and implementation lags have an overall negative effect on the total costs of the pandemic, longer decision lags also would *increase* the net benefit of ESS systems for pandemic response.
-Our base-case scenario assumes that decision-makers take one week to decide and implement non-pharmaceutical interventions.
-If policymakers took two weeks to introduce and implement interventions, the total pandemic cost increases from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "C", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == T)` dollars per person, assuming conventional surveillance.
-The early warning system effectively helps to roll-back decision lags, reducing total costs to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == F)` dollars per person, yielding an NMB of `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "NMB", NMB_comparator == F)` relative to conventional surveillance.
+Our base-case scenario assume that decision-makers take one week to decide and implement non-pharmaceutical interventions.
+If instead policymakers took two weeks to introduce and implement interventions, the total pandemic cost increases from `r pull_estimate(r$table_1_long_all, Scenario == "Base-case", variable == "C", NMB_comparator == T)` to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == T)` dollars per person, assuming conventional surveillance.
+The early warning system effectively helps to roll-back decision lags, reducing total costs to `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "C", NMB_comparator == F)` dollars per person, yielding a NMB of `r pull_estimate(r$table_1_long_all, Scenario == "2-week decision lag", variable == "NMB", NMB_comparator == F)` relative to conventional surveillance.
 Conversely, in contexts where NPI implementation is swift (i.e, decisions are made within 3 days), the NMB of ESS would be lower but still positive at `r pull_estimate(r$table_1_long_all, Scenario == "3-day decision lag", variable == "NMB", NMB_comparator == F)` dollars per person.
@@
 In that scenario, the ESS yields `r pull_estimate(r$table_1_long_all, Scenario == "1.5x stringent", variable == "NMB", NMB_comparator == F)` dollars per person of net benefit since most of the deaths would already be prevented through more stringent NPIs.
-Conversely, if policymakers were more lenient policy choices would change the system overall outcomes but also imply a higher value of the enhanced surveillance system.
+Conversely, if policy makers were more lenient policy choices would change the syst overall outcomes but also imply a higher value of the enhanced surveillance system.
 
@@
 If NPI effectiveness is 0%, an improved surveillance system has no value, as expected.
-The value of the enhanced surveillance system increases with NPI effectiveness slowly for maximum NPI effectiveness below around 50%, and peaks when the maximum NPI effectiveness reaches 60%, enough to reduce the baseline R0 of 2.5 below 1 at the pandemic onset.
+The value of the enhanced surveillance system increases with NPI effectiveness slowly for maximum NPI effectiveness below around 50%, and peaked when the maximum NPI effectiveness reaches 60%, enough to reduce the baseline R0 of 2.5 below 1 at pandemic outset.
 In general, we find that the value of the ESS peaks at a maximum NPI effectiveness value at $1-1/R0$, for different values of R0.
 Beyond that point, increases in NPI effectiveness result in lower total pandemic costs, but the value of the surveillance system starts to decrease.
+
+<!--
+
+# If we have time to look at those other cases:
+
+## ESS and policy coordination
+ESS alleviate lack of policy coordination. 
+There is a concern that ESS data will need to be integrated to be useful. However, we find that policy coordination and ESS act as substitutes.
+
+## If we have time: Marginal Value of ESS
+ESS deployment also face decreasing marginal returns. The value of ESS decreases as coverage increases.
+
+-->

### 2024-04-30 R/odin_models/stochastic_metapopulation.R

@@ -131,7 +131,7 @@ eff_c <- if (step <= t_o) c else 100000
 # Note that L_star can be > L_max so that the target intervention can remain at the lockdown level.
 L_star_ind[] <- min(100000 * (lagged_incidence[i] / N[i]) / (eff_c * p), L_max + 0.01)
 
-# Target NPI considering other jurisdictions:
+# Target NPI considering other jurisdictions (in case they coordinate policies)
 L_star_matrix[, ] <- C[i, j] * L_star_ind[j]
 
 # Stores the maximum target level at the last column for each row:

### 2024-04-30 R/scripts/run_tables_figures.R

@@ -129,17 +129,11 @@ r$fig_1 <- r$fig_1_data %>%
     color = "ESS system"
   ) +
   scale_color_manual(values = colors) +
-  scale_y_continuous(labels = scales::dollar_format(scale = 1e-3, prefix = "", accuracy = 0.5)) +
-  theme(plot.margin = margin(
-    t = 0,
-    r = 0,
-    b = 0,
-    l = 0
-  ))
+  scale_y_continuous(labels = scales::dollar_format(scale = 1e-3, prefix = "", accuracy = 0.1))
 
 r$fig_1
 
-ggsave(plot = r$fig_1, filename = "./output/fig_1.svg", units = "in", width = 6.5, height = 5, scale = 1.4, bg = "white", dpi = 300)
+ggsave(plot = r$fig_1, filename = "./output/fig_1.pdf", units = "in", width = 6.5, height = 5, scale = 1.4, bg = "white", dpi = 300)
 
 ## Fig 2 ----------------------------------------------------------------
 
@@ -163,7 +157,7 @@ r$fig_2 <- r$tau_NMB_summary %>%
 
 r$fig_2
 
-ggsave(plot = r$fig_2, filename = "./output/fig_2.svg", units = "in", width = 6.5, height = 5, scale = 1.4, bg = "white", dpi = 300)
+ggsave(plot = r$fig_2, filename = "./output/fig_2.pdf", units = "in", width = 6.5, height = 5, scale = 1.4, bg = "white", dpi = 300)
 
 
 ## Sup Figure 1 ------------------------------------------------------------
@@ -183,25 +177,14 @@ r$sup_fig_1 <- r$fig_1_data %>%
     color = "ESS system"
   ) +
   scale_color_manual(values = colors) +
-  scale_y_continuous(labels = scales::dollar_format(scale = 1e-3, prefix = "", accuracy = 1)) +
-  theme(plot.margin = margin(
-    t = 0,
-    r = 12,
-    b = 0,
-    l = 0
-  ))
+  scale_y_continuous(labels = scales::dollar_format(scale = 1e-3, prefix = "", accuracy = 1))
 
 r$sup_fig_1
 
-ggsave(plot = r$sup_fig_1, filename = "./output/sup_fig_1.svg", units = "in", width = 6.5, height = 6.5, scale = 1.4, bg = "white", dpi = 300)
[…]
